### 2014-11-07 paper/main.tex
added paper

\documentclass[11pt,twocolumn]{article}
\usepackage{multicol}

\begin{document}

%\title{Reverse Engineering Valid Input via Side Channels}
%\maketitle
\twocolumn[{%
  \centering
  \LARGE Reverse Engineering Valid Input via Side Channels \\[1em]
  \large Bogdan Copos \\
  	    Praveen Murthy \\[1.5em]
}]

\section{Abstract}
Automatic identification of bugs remains one of the biggest challenges in security.
Despite widespread research of various approaches, both static and dynamic, the problem remains (REWRITE).
One of the biggest obstacles in automatic testing is thorough exercising of the program.
While fuzzing and symbolic execution can be successful in finding bugs [cite], both approaches have limitations.
To date, there is no alternative for generating input for unknown binaries.
In this paper we present NAME, an architecture and source code independent, instrumentation-free technique for generating valid commands/input for unknown stripped binaries.
NAME exploits side channels to incrementally build valid commands which allow for more extensive testing.

\section{Introduction}

\section{Background}

fuzzers 
symbolic execution limitations (some require source, others require instrumentation, cannot handle floating point instructions, fundamental limitations such as non-linear constraints solving)
lack of architecture independent, source independent, instrumentation free automatic valid input determination.

\section{Related Work}

\section{Methodology}

\section{Evaluation}

\section{Results}

\section{Conclusions}
\end{document}

### 2014-11-13 paper/main.tex
updated paper

@@
 
+% paper for USENIX has to be 13 pages
+
 \begin{document}
 
-%\title{Reverse Engineering Valid Input via Side Channels}
-%\maketitle
 \twocolumn[{%
   \centering
-  \LARGE Reverse Engineering Valid Input via Side Channels \\[1em]
-  \large Bogdan Copos \\
-  	    Praveen Murthy \\[1.5em]
+  \LARGE Reverse Engineering Binary's Input via Side Channels \\[1em]
+  \large \textbf{Bogdan Copos}, UC Davis\\
+  	    \textbf{Praveen Murthy}, Fujitsu Laboratories of America \\[1.5em]
 }]
@@
 Despite widespread research of various approaches, both static and dynamic, the problem remains (REWRITE).
-One of the biggest obstacles in automatic testing is thorough exercising of the program.
+One of the biggest obstacles in automatic testing is extensive exercising of the program.
 While fuzzing and symbolic execution can be successful in finding bugs [cite], both approaches have limitations.
-To date, there is no alternative for generating input for unknown binaries.
+To date, there is no alternate method for generating input for unknown binaries.
 In this paper we present NAME, an architecture and source code independent, instrumentation-free technique for generating valid commands/input for unknown stripped binaries.
@@
 
+Software testing is an expensive and labor intensive process.
+Researchers have developed tools to help identify bugs and speed up the testing process.
+While many bugs have been identified, no approach is perfect and many challenges in autonomous software testing remain.
+One of the biggest obstacles is thorough exploration of the program.
+Without high code coverage, bugs may be omitted.
+At the same time, high code coverage cannot be accomplished without sufficient exercising of the program using valid input/commands.
+Valid input can be learned from documentation or specifications.
+When such information is not available, other methods must be used.
+To date, there are two methods which may be used for input generation: fuzzing and symbolic execution.
+In this paper, we present a third method for valid input generation for unknown stripped binaries.
+Our method takes advantage of instruction counters to incrementally build valid inputs which are guaranteed to expand code coverage.
+
+The paper is organized as follows [MAYBE]
+
+\section{Previous Work}
+
+Fuzzers are a popular choice amongst software testers.
+They are easy to use and have produced good results in the past.
+There are two classes of fuzzers: black box fuzzers and white box fuzzers.
+In black box fuzzing, testing is based on randomly generated input and observing output to identify problems.
+No program specifications or knowledge of the input format is available in such cases.
+On the other hand, in white box fuzzing, inputs are used to gather symbolic constraints which are in turn systematically negated and solved to produce more inputs.
+White box fuzzing differs from black box fuzzing in that it uses the program's feedback to make inferences about valid input and use constraint solvers to generate more input.
+Both fuzzing approach have some limitations.
+Since black box fuzzing is completely random, its effectiveness is limited, especially with respect to exercising programs.
+White box fuzzing can reach better code coverage but can be slow.
+[MAYBE TALK ABOUT FUZZING RESULTS, CITATIONS]
+
+Symbolic execution is another method used for generating valid input.
+In symbolic execution, an interpreter gathers constraints related to user input throughout the program's execution, rather than using actual user input.
+At the end of the execution, a constraint solver takes the constraints gathered and outputs valid program input.
+Symbolic execution has been studied extensively and has been proved effective [CITE].
+[TALK ABOUT ONE OF THE CITATIONS]
+Despite positive results, symbolic execution has a fundamental limitation with respect to constraint solving.
+While it performs well when constraints are linear, non-linear constraints complicate things greatly.
+Furthermore, there is no autonomous symbolic execution engine.
+All symbolic executioners today require some instrumentation of the program to determine the program's input buffer.
+Some symbolic execution engines even require source code, which is not always available to the tester.
+
+The method presented in this paper takes advantage of instruction hardware counters to observe program's behavior and incrementally build valid input.
+As opposed to fuzzers, where input is randomly generated, this approach uses side channels to intelligently compose valid program input.
+Also, unlike symbolic execution, this approach does not require any instrumentation or source code.
+
 \section{Background}
 
-fuzzers 
-symbolic execution limitations (some require source, others require instrumentation, cannot handle floating point instructions, fundamental limitations such as non-linear constraints solving)
-lack of architecture independent, source independent, instrumentation free automatic valid input determination.
+Side channels are streams of useful information that can be retrieved from the hardware of a device running a given program.
+This information can be then used to learn more about the program's internals.
+Side channels have been widely explored by security specialist in order to exploit weaknesses in programs.
+In cryptography, side channel attacks exploit such side channels to break crypto systems rather than attempting brute force or finding weaknesses in the underlying algorithms.
+Recently, side channel attacks, such as OpenSSL's CREAM cache timing attack have made the headlines. [CITE]
+Similarly to CREAM and other side channel attacks, the technique discussed in this paper exploits side channels to generate valid input for binaries.
 
-\section{Related Work}
+The majority of software programs accept user input, transform the input and provide the result as useful output. [REWORD]
+However, most programs do not accept completely random input.
+User input usually passes through an input validation filter.
+Such filters are snippets of code responsible for distinguishing between valid input (i.e.
+input the program was constructed to understand) and input which is not useful or does not follow the desired format.
+Although they can be very sophisticated, input validation filters are often a combination of `strcmp()` function calls and conditional statements.
+These additional instructions, designed to verify the input, can be measured to make inferences about the program's validation mechanism.
+
+Most modern microprocessors have a set of special-purpose registers built to store the counts of hardware events.
+Each counter can be programmed to measure specific events such as cache misses, floating point operations, and even instructions executed.
+Utilities, such as Linux/Unix `perf', take advantage of such hardware counters in order to provide developers with useful information about a program's performance.
+`perf' is a kernel module which ... [HOW DOES PERF WORK]
+In this work, we use `perf' to count the number of instructions executed as a program is given various input.
 
@@
 
+[some short introduction]
+For the purpose of this paper, good or valid input is defined as input which successfully passes the input validation mechanism of a given program.
+Contrarily, bad or invalid input is input which does not pass validation and is rejected or ignored by a given program.
+Building valid input for a given program using our method is comprised of two components.
+One component is responsible for finding characters of valid input strings at a given index.
+The second component coordinates with the first to incrementally build valid input strings.
+In order to build valid input for a given program, we start by measuring the number of instructions retired during the program's execution using the `perf' utility.
+Specifically, we want to observe differences in the number of instructions executed as a result of different input.
+
+\subsection{Determining valid characters at a given index}
+
+The process of finding valid characters for a given index is composed of two stages, the testing stage and the filtering stage.
+
+\subsubsection{Testing stage}
+
+During the testing stage, the program is executed with `perf' and given all of the printable characters, one at a time, as input.
+Each printable character results in a different program execution.
+At the end of each execution, the number of instructions retired for that printable character is recorded in a dictionary.
+The difference in the number of instructions between various printable characters is a side effect of the input validation mechanism.
+The dictionary data structure is then used in the filtering stage to determine which characters are likely to be part of a valid input string for the given program.
+
+\subsubsection{Filtering stage}
+
+The filtering stage is responsible for identifying characters which are highly likely to be part of a valid input string at a given index.
+Using the dictionary obtained from the testing stage, the number of instructions retired are analyzed.
+First, the mode of the number of instructions executed is calculated.
+The filtering of possible valid characters for a given index \textit{i}, of an input string \textit{s}, is based on the assumption that the majority of characters do not belong at index \textit{i} of string \textit{s} (i.e.
+\textit{s[i]}).
+Therefore, valid characters are defined as any printable character for which the number of instructions retired is outside the mode.
+
+While the initial reaction is to expect a valid character to result in more instructions executed than a non-valid character, this is not always the case.
+As discussed earlier, the input validation mechanism can be complex and can vary in behavior.
+The validation mechanism may verify input against all accepted input strings before denying the provided input.
+Such a scenario may cause bad input to result in more instructions executed than valid input.
+
+\subsection{Building valid input string}
+
+In order to build valid input strings, ...
+The two stages are repeated for every index of the input string.
+
 \section{Evaluation}
+
+- talk about interview problem [MAYBE?]
 
@@
 
+\section{Limitations}
+
+Here are some limitations:
+
+- programs that have random number of instructions executing to defeat side channels.
+
+- programs that randomly generate input at each execution (this can be defeated by sampling number of instructions retired during one execution).
+
+- programs that do not have any particular rules for acceptable input, such as an echo program.
+
+- programs where input fields depend on each other.
+
+- programs that validate input fields out of order.
+
 \section{Conclusions}

### 2014-11-16 paper/main.tex
wrote more of the paper...

--- 2014-11-07 paper/main.tex
@@
-\documentclass[11pt,twocolumn]{article}
+\documentclass[10pt,twocolumn]{article}
 \usepackage{multicol}
+\usepackage{graphicx}
+
+% definitions
+\def \tool {InputFinder}
 
@@
 
-%\title{Reverse Engineering Valid Input via Side Channels}
-%\maketitle
 \twocolumn[{%
   \centering
-  \LARGE Reverse Engineering Valid Input via Side Channels \\[1em]
-  \large Bogdan Copos \\
-  	    Praveen Murthy \\[1.5em]
+  \large \textbf{InputFinder: Reverse Engineering Closed Binaries via Side Channels} \\[2em]
+  Bogdan Copos, University of California, Davis\\
+  Praveen Murthy, Fujitsu Laboratories of America \\[4em]
 }]
 
-\section{Abstract}
-Automatic identification of bugs remains one of the biggest challenges in security.
-Despite widespread research of various approaches, both static and dynamic, the problem remains (REWRITE).
-One of the biggest obstacles in automatic testing is thorough exercising of the program.
-While fuzzing and symbolic execution can be successful in finding bugs [cite], both approaches have limitations.
-To date, there is no alternative for generating input for unknown binaries.
-In this paper we present NAME, an architecture and source code independent, instrumentation-free technique for generating valid commands/input for unknown stripped binaries.
-NAME exploits side channels to incrementally build valid commands which allow for more extensive testing.
+\begin{abstract}
+User input and test suites are crucial to dynamic program analysis techniques.
+Without extensive testing of the program, bugs may be omitted.
+While fuzzing and symbolic execution can be successful in finding input, both approaches have limitations.
+To date, there is no alternative method for generating input for closed (source) binaries.
+In this paper we present an architecture and source code independent, instrumentation-free technique for generating valid input for closed stripped binaries and its implementation \tool{}.
+\tool{} exploits side channels to automatically build valid user-input allowing for more extensive testing.
+Additionally, \tool{} can determine input buffer size and make inferences about input types (e.g. alphabetic, numeric, alphanumeric, etc.).
+\end{abstract}
 
@@
 
+Software testing is an expensive and labor intensive process.
+For many years, researchers have explored both static and dynamic methods for analyzing programs.
+While static analysis is the more thorough approach, dynamic testing is often preferred due to its time efficiency.
+Dynamic testing heavily depends on availability of user input or test suites.
+Without input, the program cannot be properly tested and bugs may be omitted.
+While valid input can be learned from software documentation or source code, when such information is not available, other methods must be used.
+Automated dynamic analysis techniques such as fuzzing and symbolic execution attempt to generate input for a given program.
+Currently these are the only two options available and both approaches present discouraging limitations.
+
+[DESCRIBE OUR WORK]
+
+[THE CONTRIBUTIONS OF THIS PAPER ARE AS FOLLOWS:]
+
+In this paper, we introduce a third method for generating valid user-input for closed stripped binaries. 
+Our method exploits side channels, specifically retired instructions counters, to build valid input for closed binaries.
+We also present an implementation of our method, \tool{}.
+
+The paper is organized as follows.
+Section 2 covers background information.
+Section 3 discusses related works and differences with respect to the presented method.
+In Section 4 our method is explained.
+Section 5 describes evaluation of the method presented and the results.
+In Section 6 we discuss future work and limitations of the presented work.
+
 \section{Background}
 
-fuzzers 
-symbolic execution limitations (some require source, others require instrumentation, cannot handle floating point instructions, fundamental limitations such as non-linear constraints solving)
-lack of architecture independent, source independent, instrumentation free automatic valid input determination.
+\subsection{Side Channels}
 
-\section{Related Work}
+Side channels are streams of information that can be retrieved from the hardware of a device running a given program.
+This information can be used to learn more about the program's internals.
+Side channels have been widely explored by security researchers in order to exploit weaknesses in programs.
+In cryptography, side channel attacks exploit such information channels to break crypto systems rather than applying traditional brute force or reverse engineering techniques.
+Recently, side channel attacks, such as OpenSSL's CREAM cache timing attack have made headlines. [CITE]
+Similar to CREAM and other side channel attacks, the technique discussed in this paper exploits such information channels to generate valid input for binaries.
+
+\subsection{Hardware counter registers}
+
+Most modern microprocessors have a set of special-purpose registers built to store the counts of hardware events.
+Each counter can be programmed to measure specific events such as cache misses, floating point operations, and even instructions executed.
+Utilities, such as Linux/Unix `perf', take advantage of such hardware counters in order to provide developers with useful information about a program's performance.
+`perf' is a tool which allows developers to specify the desired hardware event they wish to measure.
+Once events have been selected, `perf' interacts with kernel modules to program the hardware counters accordingly.
+In this work, we use `perf' to count the number of \textbf{user-land} instructions executed as a program is tested with various input.
+The number of total instructions retired oscillates significantly for a binary from execution to execution.
+This phenomena reflects fluctuations in instructions executed by the kernel depending on the state of the machine due to cache-misses and other events.
+Since we rely on this count, such variations are problematic.
+However, the number of user-land instructions only varies by one instruction on average, giving us a reliable counter.
+
+\section{Previous Work}
+
+Fuzzers are a popular choice amongst software testers.
+They are easy to use and have produced good results in the past.
+There are two classes of fuzzers: black box fuzzers and white box fuzzers.
+In black box fuzzing, testing is based on randomly generating input and observing output to identify problems.
+No program specifications or knowledge of the input format is available in such cases.
+On the other hand, in white box fuzzing, inputs are used to gather symbolic constraints which are in turn systematically negated and solved to produce more inputs.
+White box fuzzing differs from black box fuzzing in that it uses the program's feedback to make inferences about valid input and use constraint solvers to generate more input.
+Both fuzzing approach have limitations.
+Since black box fuzzing is completely random, its effectiveness is limited, especially with respect to exercising programs.
+White box fuzzing can attain better code coverage but can be slow.
+[MAYBE TALK ABOUT FUZZING RESULTS, CITATIONS]
+
+Symbolic execution is another method used for generating valid input.
+In symbolic execution, an interpreter gathers constraints related to user input throughout the program's execution, rather than using actual user input.
+At the end of the execution, a constraint solver takes the constraints gathered and outputs valid program input.
+Symbolic execution has been studied extensively and has been proved effective [CITE].
+[TALK ABOUT ONE OF THE CITATIONS]
+Despite positive results, symbolic execution has a fundamental limitation with respect to constraint solving.
+While symbolic execution engines perform well when constraints are linear, non-linear constraints greatly impact efficiency.
+Furthermore, there is no autonomous symbolic execution engine.
+All symbolic executioners today require some instrumentation of the program to determine the program's input buffer.
+Some symbolic execution engines even require source code, which is not always available to the tester.
+
+[TALK ABOUT SIDE CHANNEL ATTACKS AND CITATIONS]
+
+The method presented in this paper takes advantage of hardware counters to observe program's behavior and automatically build valid input.
+As opposed to fuzzers, where input is randomly generated, this approach uses side channels to intelligently compose valid program input.
+Also, unlike symbolic execution, this approach is not hindered by constraint solver limitations and does not require any instrumentation or source code.
 
@@
 
-\section{Evaluation}
+The majority of software programs accept user input, perform transformations on the input and output the result.
+However, most programs do not accept completely random input.
+User input usually passes through an input validation filter.
+Such filters are snippets of code responsible for distinguishing between valid or good input (i.e.
+input the program was constructed to understand) and bad input (i.e.
+input which is not useful or does not follow the desired format).
+Although validation mechanisms can be very sophisticated, they are often a combination of string comparisons and conditional statements.
+Our method exploits the presence (or lack) of these additional instructions during an execution to make inferences about the program's validation mechanism.
+Specifically, we test the program with different input and compare the number of retired user-land instructions executed.
+Assuming most input is bad, we observe differences in the number of instructions executed and detect input which passes the validation mechanism.
 
-\section{Results}
+We present \tool{} an implementation of the method described above.
+Given a closed binary, \tool{} can build valid input, inform the tester about the type of acceptable input, and determine input buffer size.
+
+\subsection{Finding Valid Input}
+
+\begin{figure*}[t]
+\centering
+\includegraphics[height=2.5in,width=6.5in]{architecture.png}
+\label{fig:architecture}
+\caption{}
+\end{figure*}
+
+In order to build valid input for a given program, \tool{} measures the number of user-land instructions retired during the program's execution using the `perf' utility.
+Specifically, \tool{} observes differences in the number of instructions executed as a result of different input.
+Generating valid input strings is a two-step process.
+\tool{} works by identifying valid characters for each index \textit{i}, starting with \textit{i=0} and incrementally building valid input strings using the discovered characters.
+Figure~\ref{fig:architecture} shows the overall architecture of \tool{}.
+
+\subsubsection{Finding Valid Characters}
+
+The process of finding valid characters for a given index is composed of two stages, the testing stage and the filtering stage.
+During the testing stage, the program is executed with `perf' and given all of the printable characters, one at a time, as input.
+The automatic interaction with the program is performed using `Expect', an extension to the `Tcl' scripting language [CITE]
+In our implementation, each printable character results in a different program execution.
+At the end of each execution, the number of instructions retired for that printable character is recorded in a dictionary.
+The difference in the number of instructions between various printable characters is a side effect of the input validation mechanism.
+The dictionary data structure is used in the filtering stage to determine which characters are likely to be part of a valid input string for the given program.
+
+The filtering stage is responsible for identifying characters which are highly likely to be part of a valid input string at a given index.
+Using the dictionary obtained from the testing stage, the number of instructions retired are analyzed.
+First, the mode of the number of instructions executed is calculated.
+The filtering of possible valid characters for a given index \textit{i}, of an input string \textit{s}, is based on the assumption that the majority of characters do not belong at index \textit{i} of string \textit{s} (i.e.
+\textit{s[i]}).
+Therefore, valid characters are defined as any printable character for which the number of instructions retired is outside the mode.
+
+While the initial reaction is to expect a valid character to result in more instructions executed than a non-valid character, this is not always the case.
+As discussed earlier, the input validation mechanism can be complex and can vary in behavior.
+The validation mechanism may verify input against all accepted input strings before denying the provided input.
+Such a scenario may cause bad input to result in more instructions executed than valid input.
+
+When no valid characters are identified as part of the filtering stage and the current index is greater than 0, the end of a valid input string has been reached.
+If the index is 0, the given binary has no predefined input commands.
+
+\subsubsection{Building Input Strings}
+
+\begin{figure*}[t]
+\centering
+\includegraphics[height=2in,width=6.5in]{string_builder.png}
+\label{fig:string_builder}
+\caption{}
+\end{figure*}
+
+Building valid input strings is a recursive process composed of multiple calls to the character finding component.
+Starting with an empty string and index 0, \tool{} tries to find valid characters at each index and incrementally builds valid input strings.
+More specifically, the process starts at index 0.
+Using the character finding technique, relevant characters for index 0 are identified.
+The task is forked \textit{n} times, where \textit{n} is the number of characters found at the current index (i.e. index 0).
+Each child task is assigned one of the \textit{n} characters and is responsible for finding valid characters at the next index.
+These steps are repeated until no valid characters are found, marking the end of the input string.
+Figure~\ref{fig:string_builder} depicts the process.
+
+\subsection{Finding Desired Input Size}
+
+The same side channels can be used to detect the expected size of the program's input or even the program's input buffer.
+Most programs are designed to read a predefined number of bytes from the standard input interface and store them in a buffer.
+When the program is given more input than the size of the buffer, additional instructions are executed.
+Our method observes such differences in the number of user-instructions retired as the program is given increasingly larger input.
+
+To start, \tool{} finds all the invalid characters for index 0 of all input strings and picks one at random.
+The character selected is used to build invalid input strings of various sizes.
+It is crucial to use invalid input to avoid noisy fluctuations in the number of instructions executed due to valid input processing.
+Initially, the program is executed with short length input strings (i.e. two and three characters).
+The number of user-land instructions executed during these two executions are recorded.
+These two values may be equal or differ, depending on the program's logic.
+If the two values are equal [EQUAL]
+The two values can also differ because... [DIFFERENT]
+
+\subsection{Classifying Input}
+
+Our method is also capable of classifying the input discovered.
+Generated input strings are classified in three categories: alphabetic, numeric, alphanumeric.
+Similarly, the classification algorithm detects special characters.
+Once enough input strings have been identified, \tool{} uses the categorization data to generalize about the type of input the program expects.
+Such information is valuable especially in cases where the inputs discovered are all numeric or contain (one or more) special characters.
+Having learned such information about the program, the testing platform (e.g. fuzzer) can be adjusted to proceed accordingly.
+
+\section{Evaluation and Results}
+
+\tool{} was tested on [NUMBER] closed-source binaries to measure its effectiveness.
+The results are summaried in Figure [FIGURE].
+Out of the total [NUMBER] binaries, [NUMBER2] had predefined accepted input strings.
+\tool{} found 100\% of the input strings for [NUMBER3]...
+[CONTINUE].
+
+Additionally, our method was tested on a binary used for technical interviews.
+The binary was designed as a reverse engineering challenge for interviewers.
+To pass the challenge, the interviewer has to crack the secret passcode to the program.
+A blog post describes the manual process used to crack the password [CITE].
+The manual process is tedious.
+As described in the blog post, the cracking of the binary can take several hours and requires advanced skills and years of background knowledge.
+\tool{} cracked the binary in a matter of minutes automatically, proving itself as a useful tool not only finding valid input but also cracking passwords.
+
+\section{Future Work and Limitations}
+
+The current implementation use `perf' to count instructions, which outputs only when the program executed exits.
+As a result, the tested program must be executed once with every input.
+This is not only a slow and inefficient process but also hinders the ability to crack programs which randomly generate passwords at execution time.
+In the future, we'd like to explore methods which allow sampling of instruction count during runtime.
+
+While the method presented in this paper has many advantages, it also has a few limitations.
+For certain programs, such as an echo server, our method is not any more efficient than a fuzzer.
+Additionally, our method cannot handle cases in which user input is composed of multiple inter-dependent fields or when programs validate fields of the input out of order.
 
 \section{Conclusions}
+
 \end{document}

### 2015-01-02 paper/main.tex
state changing stuff

@@
 \begin{abstract}
-User input and test suites are crucial to dynamic program analysis techniques.
+Valid input and test suites are crucial to dynamic program analysis techniques.
 Without extensive testing of the program, bugs may be omitted.
@@
 To date, there is no alternative method for generating input for closed (source) binaries.
-In this paper we present an architecture and source code independent, instrumentation-free technique for generating valid input for closed stripped binaries and its implementation \tool{}.
+In this paper we present an architecture and source code independent, instrumentation-free technique and its implementation, \tool{}, for generating valid input for closed stripped binaries.
 \tool{} exploits side channels to automatically build valid user-input allowing for more extensive testing.
-Additionally, \tool{} can determine input buffer size and make inferences about input types (e.g. alphabetic, numeric, alphanumeric, etc.).
+Additionally, \tool{} can determine expected input size and make inferences about input types (e.g. alphabetic, numeric, alphanumeric, etc.).
 \end{abstract}
@@
 While static analysis is the more thorough approach, dynamic testing is often preferred due to its time efficiency.
-Dynamic testing heavily depends on availability of user input or test suites.
+Dynamic testing heavily depends on availability of input or test suites.
 Without input, the program cannot be properly tested and bugs may be omitted.
@@
 
-[DESCRIBE OUR WORK]
+In this paper, we introduce a third method for generating valid user-input for closed stripped binaries. 
+Our method exploits side channels to build valid input for closed binaries.
+Specifically, our method relies on hardware instruction counter registers.
+We show that by measuring the number of user-land instructions retired and comparing them between different executions of the program with various input, it is possible to learn the program's expected input.
+We also present an implementation of our method, \tool{}.
 
-[THE CONTRIBUTIONS OF THIS PAPER ARE AS FOLLOWS:]
-
-In this paper, we introduce a third method for generating valid user-input for closed stripped binaries. 
-Our method exploits side channels, specifically retired instructions counters, to build valid input for closed binaries.
-We also present an implementation of our method, \tool{}.
+The contributions of this paper are as follows:
+\begin{itemize}
+	\item We introduce a new approach to automatically generate input valid strings for unknown, closed-source binaries by leveraging hardware instruction counters.
+Our method relies on observing additional instructions executed by programs during their input validation process to learn valid input strings.
+	\item We implement the method presented and evaluate it using example binaries published by DARPA as part of the 2014 Cyber Grand Challenge competition.
+Additionally, we compare our approach to the manual process used by a security specialist to crack a binary as part of a job interview.
+\end{itemize}
 
@@
 `perf' is a tool which allows developers to specify the desired hardware event they wish to measure.
-Once events have been selected, `perf' interacts with kernel modules to program the hardware counters accordingly.
+Once events have been s§elected, `perf' interacts with kernel modules to program the hardware counters accordingly.
 In this work, we use `perf' to count the number of \textbf{user-land} instructions executed as a program is tested with various input.
@@
 Symbolic execution is another method used for generating valid input.
-In symbolic execution, an interpreter gathers constraints related to user input throughout the program's execution, rather than using actual user input.
+In symbolic execution, an interpreter gathers constraints related to input throughout the program's execution, rather than using the provided input.
 At the end of the execution, a constraint solver takes the constraints gathered and outputs valid program input.
@@
 
-[TALK ABOUT SIDE CHANNEL ATTACKS AND CITATIONS]
+Side channel attacks and citations\ldots
 
@@
 
-The majority of software programs accept user input, perform transformations on the input and output the result.
+The majority of software programs accept input, perform transformations on the input and output the result.
 However, most programs do not accept completely random input.
-User input usually passes through an input validation filter.
+Input usually passes through an input validation filter.
 Such filters are snippets of code responsible for distinguishing between valid or good input (i.e.
@@
 Specifically, we test the program with different input and compare the number of retired user-land instructions executed.
-Assuming most input is bad, we observe differences in the number of instructions executed and detect input which passes the validation mechanism.
+Assuming most input is invalid, our method observes differences in the number of instructions executed (from the norm) to detect input which does pass the validation mechanism.
 
@@
 Specifically, \tool{} observes differences in the number of instructions executed as a result of different input.
-Generating valid input strings is a two-step process.
-\tool{} works by identifying valid characters for each index \textit{i}, starting with \textit{i=0} and incrementally building valid input strings using the discovered characters.
+Generating valid input strings is a recursive two-step process.
+\tool{} works by identifying valid characters for each index \textit{i}, starting with \textit{i}=0 and incrementally building valid input strings using the discovered characters.
 Figure~\ref{fig:architecture} shows the overall architecture of \tool{}.
@@
 
-The process of finding valid characters for a given index is composed of two stages, the testing stage and the filtering stage.
-During the testing stage, the program is executed with `perf' and given all of the printable characters, one at a time, as input.
+The process of finding valid characters for a given index is composed of two stages: the testing stage and the filtering stage.
+During the testing stage, the program is executed with `perf' and being provided as input, all of the printable characters, one at a time.
+Once the program has processed the input and is waiting for more input, the program is stopped.
 The automatic interaction with the program is performed using `Expect', an extension to the `Tcl' scripting language [CITE]
 In our implementation, each printable character results in a different program execution.
-At the end of each execution, the number of instructions retired for that printable character is recorded in a dictionary.
-The difference in the number of instructions between various printable characters is a side effect of the input validation mechanism.
-The dictionary data structure is used in the filtering stage to determine which characters are likely to be part of a valid input string for the given program.
+At the end of each execution, the number of instructions retired is outputted by `perf'.
+This information is saved in a dictionary as a tuple of the format ( <printable character> , <number of instructions> ).
+As mentioned earlier, the differences in the number of instructions between various printable characters reflect different execution paths taken during the validation process.
 
 The filtering stage is responsible for identifying characters which are highly likely to be part of a valid input string at a given index.
-Using the dictionary obtained from the testing stage, the number of instructions retired are analyzed.
+Using the dictionary data obtained from the testing stage, the number of instructions retired are analyzed.
 First, the mode of the number of instructions executed is calculated.
-The filtering of possible valid characters for a given index \textit{i}, of an input string \textit{s}, is based on the assumption that the majority of characters do not belong at index \textit{i} of string \textit{s} (i.e.
-\textit{s[i]}).
+The filtering of possible valid characters for a given index \textit{i}, of an input string \textit{s}, is based on the assumption that the majority of printable characters \textit{c} do not belong at index \textit{i} of string \textit{s} (i.e.
+\textit{c} $\notin$ \textit{s[i]}).
 Therefore, valid characters are defined as any printable character for which the number of instructions retired is outside the mode.
@@
 The validation mechanism may verify input against all accepted input strings before denying the provided input.
-Such a scenario may cause bad input to result in more instructions executed than valid input.
+Such a scenario results in more instructions executed for invalid input than valid input.
 
 When no valid characters are identified as part of the filtering stage and the current index is greater than 0, the end of a valid input string has been reached.
-If the index is 0, the given binary has no predefined input commands.
+If the index is 0 and no valid characters have been found, the given binary has no predefined input commands.
 
@@
 
-Building valid input strings is a recursive process composed of multiple calls to the character finding component.
+Building valid input strings is a recursive process.
 Starting with an empty string and index 0, \tool{} tries to find valid characters at each index and incrementally builds valid input strings.
-More specifically, the process starts at index 0.
-Using the character finding technique, relevant characters for index 0 are identified.
+More specifically, using the character finding technique described previously, relevant characters for index 0 are identified.
 The task is forked \textit{n} times, where \textit{n} is the number of characters found at the current index (i.e. index 0).
@@
 The same side channels can be used to detect the expected size of the program's input or even the program's input buffer.
-Most programs are designed to read a predefined number of bytes from the standard input interface and store them in a buffer.
-When the program is given more input than the size of the buffer, additional instructions are executed.
+Most programs are designed to read a predefined number of bytes from the source of input and store them in a buffer.
+Depending on the implementation, when a program is given more input than the size of the buffer, additional instructions may be executed.
+Regardless of the number of instructions executed, the behavior of the program changes.
+A program may exit when the number of provided bytes is larger than expected.
+Similarly, once the input buffer size limit has been reach, a program may check to see if the next byte is a newline or EOF.
+Some methods, such as C's `fgets()' function, read character by character [EXPLAIN]
 Our method observes such differences in the number of user-instructions retired as the program is given increasingly larger input.
 
-To start, \tool{} finds all the invalid characters for index 0 of all input strings and picks one at random.
+To start, \tool{} finds the set of all invalid characters for index 0 of all input strings and picks one at random.
 The character selected is used to build invalid input strings of various sizes.
@@
 Similarly, the classification algorithm detects special characters.
-Once enough input strings have been identified, \tool{} uses the categorization data to generalize about the type of input the program expects.
+Once enough input strings have been identified, \tool{} uses the categorization data to generalize about the type of input the program accepts.
 Such information is valuable especially in cases where the inputs discovered are all numeric or contain (one or more) special characters.
@@
 \tool{} was tested on [NUMBER] closed-source binaries to measure its effectiveness.
-The results are summaried in Figure [FIGURE].
+The results are summarized in Figure [FIGURE].
 Out of the total [NUMBER] binaries, [NUMBER2] had predefined accepted input strings.
@@
 
-The current implementation use `perf' to count instructions, which outputs only when the program executed exits.
-As a result, the tested program must be executed once with every input.
+The current implementation uses `perf' to count instructions, which outputs only when the program executed exits.
+As a result, the tested program must be executed once with every new input string.
 This is not only a slow and inefficient process but also hinders the ability to crack programs which randomly generate passwords at execution time.
@@
 While the method presented in this paper has many advantages, it also has a few limitations.
-For certain programs, such as an echo server, our method is not any more efficient than a fuzzer.
-Additionally, our method cannot handle cases in which user input is composed of multiple inter-dependent fields or when programs validate fields of the input out of order.
+For certain programs that allow any input, such as an echo server, our method is not any more efficient than a fuzzer.
+Additionally, our method cannot handle cases in which input is composed of multiple inter-dependent fields or when programs validate input fields out of order.
 

### 2015-04-21 paper/main.tex
updated paper

@@
 \begin{abstract}
-Valid input and test suites are crucial to dynamic program analysis techniques.
-Without extensive testing of the program, bugs may be omitted.
-While fuzzing and symbolic execution can be successful in finding input, both approaches have limitations.
-To date, there is no alternative method for generating input for closed (source) binaries.
-In this paper we present an architecture and source code independent, instrumentation-free technique and its implementation, \tool{}, for generating valid input for closed stripped binaries.
-\tool{} exploits side channels to automatically build valid user-input allowing for more extensive testing.
-Additionally, \tool{} can determine expected input size and make inferences about input types (e.g. alphabetic, numeric, alphanumeric, etc.).
+
+We propose a new approach for the autonomous generation of test suites for dynamic analysis of closed binaries and build a general tool called \tool{} for determining input for a closed binary and identifying the appropriate protocol (if any).
+We show that (a) our technique can autonomously detect input, (b) one can easily leverage the results from our tool to build a sophisticated dynamic analysis tool, (c) the evaluation of \tool{} are on par and often exceed prior works designed for autonomous testing of programs, which we demonstrate by comparing \tool{}'s results with those of prior works on a set of binaries from XXX.
+Since \tool{}'s analysis 
+%
+
+The effectiveness of dynamic analysis techniques depends heavily on the completeness of the test suite applied during the analysis process.
+Test suites are often composed by developers and aim at covering (and testing) all of the functionality of a software system.
+However, test suites may still not be complete, if they exist at all.
+To date, only two methods exist for automatically generating test input for closed binaries: black-box fuzzing and symbolic execution.
+Despite previous success methods in identifying bugs, both techniques have discouraging limitations.
+In this paper, we present a (novel?) method for automatically generating test suites for closed binaries.
+To assess its effectiveness, we have developed \tool{} and tested it against XXXX binaries from XXXXX.
+With a success rate of XXXX, \tool{}
+
 \end{abstract}
@@
 
+%introduction
 Software testing is an expensive and labor intensive process.
-For many years, researchers have explored both static and dynamic methods for analyzing programs.
-While static analysis is the more thorough approach, dynamic testing is often preferred due to its time efficiency.
-Dynamic testing heavily depends on availability of input or test suites.
-Without input, the program cannot be properly tested and bugs may be omitted.
-While valid input can be learned from software documentation or source code, when such information is not available, other methods must be used.
+For many years, researchers have explored both static and dynamic methods for analyzing programs \cite{}.
+While static analysis is thought to be more thorough, dynamic testing is often preferred due to its time efficiency.
+However, dynamic testing heavily depends on availability of input or test suites.
+
+%motivation
+Without complete test suites, the program cannot be properly tested and bugs may be omitted.
+While valid input can be learned from software documentation or source code, when such information is not available or complete, other methods must be used.
 Automated dynamic analysis techniques such as fuzzing and symbolic execution attempt to generate input for a given program.
@@
 
-In this paper, we introduce a third method for generating valid user-input for closed stripped binaries. 
-Our method exploits side channels to build valid input for closed binaries.
-Specifically, our method relies on hardware instruction counter registers.
-We show that by measuring the number of user-land instructions retired and comparing them between different executions of the program with various input, it is possible to learn the program's expected input.
-We also present an implementation of our method, \tool{}.
+In this paper, we introduce a new method for generating valid user-input for closed binaries. 
+Our method is composed of two main components.
+The first component exploits side channels to build valid input for closed binaries.
+Specifically, as the program is provided various input, \tool{} records the number of instructions retired by the program for each input and compares the recorded value to learn the program's expected input.
+The second component uses the discovered input to build a protocol state machine associated with the tested program.
+This enables the generation of a more thorough test suite.
 
-The contributions of this paper are as follows:
+%We show that by measuring the number of user-land instructions retired and comparing them between different executions of the program with various input, it is possible to learn the program's expected input.
+%We also present an implementation of our method, \tool{}.
+
+The main contributions of this paper are as follows:
 \begin{itemize}
 	\item We introduce a new approach to automatically generate input valid strings for unknown, closed-source binaries by leveraging hardware instruction counters.
-Our method relies on observing additional instructions executed by programs during their input validation process to learn valid input strings.
-	\item We implement the method presented and evaluate it using example binaries published by DARPA as part of the 2014 Cyber Grand Challenge competition.
+Our method relies on observing changes in the number of instructions executed by programs during their input validation process to learn valid input strings.
+	\item We develop a method for generating protocol state machines using the valid input previously found.
+	\item We implement the techniques presented and evaluate the tool using binaries published by DARPA as part of the 2014 Cyber Grand Challenge competition.
 Additionally, we compare our approach to the manual process used by a security specialist to crack a binary as part of a job interview.
@@
 Section 3 discusses related works and differences with respect to the presented method.
-In Section 4 our method is explained.
-Section 5 describes evaluation of the method presented and the results.
+In Section 4 our work is explained in detail. First, the method for finding valid input is described. Then we introduce our method for building protocol state machines using the discovered input.
+Section 5 describes evaluation of the work presented and the results.
 In Section 6 we discuss future work and limitations of the presented work.
 
-\section{Background}
+\section{Background} \label{background}
 
-\subsection{Side Channels}
+\subsection{Side Channels} \label{bg_sidechannels}
 
 Side channels are streams of information that can be retrieved from the hardware of a device running a given program.
-This information can be used to learn more about the program's internals.
+This information can be used to learn about the program's internal components.
 Side channels have been widely explored by security researchers in order to exploit weaknesses in programs.
-In cryptography, side channel attacks exploit such information channels to break crypto systems rather than applying traditional brute force or reverse engineering techniques.
+In cryptography, side channel attacks exploit such information channels to defeat crypto systems.
 Recently, side channel attacks, such as OpenSSL's CREAM cache timing attack have made headlines. [CITE]
@@
 
-\subsection{Hardware counter registers}
+\subsection{Hardware counter registers} \label{bg_hardwarereg}
 
-Most modern microprocessors have a set of special-purpose registers built to store the counts of hardware events.
-Each counter can be programmed to measure specific events such as cache misses, floating point operations, and even instructions executed.
+Most modern microprocessors have a set of special-purpose registers used to count hardware events.
+Each register, or counter, can be programmed to measure specific events such as cache misses, floating point operations, and even instructions executed.
 Utilities, such as Linux/Unix `perf', take advantage of such hardware counters in order to provide developers with useful information about a program's performance.
-`perf' is a tool which allows developers to specify the desired hardware event they wish to measure.
-Once events have been s§elected, `perf' interacts with kernel modules to program the hardware counters accordingly.
+`perf' is a Unix tool which allows developers to specify the desired hardware event they wish to measure.
+Once events have been selected, `perf' interacts with kernel modules to program the hardware counters accordingly.
 In this work, we use `perf' to count the number of \textbf{user-land} instructions executed as a program is tested with various input.
 The number of total instructions retired oscillates significantly for a binary from execution to execution.
-This phenomena reflects fluctuations in instructions executed by the kernel depending on the state of the machine due to cache-misses and other events.
+This phenomena reflects fluctuations in instructions executed by the kernel depending on the state of the machine due to cache-misses and other factors.
 Since we rely on this count, such variations are problematic.
-However, the number of user-land instructions only varies by one instruction on average, giving us a reliable counter.
+However, experimentally, the number of user-land instructions only varies, on average, by one instruction, giving us a reliable counter.
 
-\section{Previous Work}
+\subsection{Execution/Dynamic Similarity} \label{bg_dynamicsim}
+
+\section{Previous Work} \label{prevwork}
 
 Fuzzers are a popular choice amongst software testers.
-They are easy to use and have produced good results in the past.
+They are easy to use, implement, and have produced surprising results in the past.
 There are two classes of fuzzers: black box fuzzers and white box fuzzers.
-In black box fuzzing, testing is based on randomly generating input and observing output to identify problems.
+In black box fuzzing, the program is watched as it is tested with randomly generated input.
+If the program crashes, the input(s) are recorded and reported.
 No program specifications or knowledge of the input format is available in such cases.
@@
 White box fuzzing differs from black box fuzzing in that it uses the program's feedback to make inferences about valid input and use constraint solvers to generate more input.
-Both fuzzing approach have limitations.
-Since black box fuzzing is completely random, its effectiveness is limited, especially with respect to exercising programs.
-White box fuzzing can attain better code coverage but can be slow.
+Both fuzzing approaches have limitations.
+Since black box fuzzing is completely random, its effectiveness is limited, especially with respect to fully exercising programs.
+White box fuzzing can attain better code coverage but can be slow and can be crippled by innate limitations of constraint solvers.
 [MAYBE TALK ABOUT FUZZING RESULTS, CITATIONS]
@@
 At the end of the execution, a constraint solver takes the constraints gathered and outputs valid program input.
-Symbolic execution has been studied extensively and has been proved effective [CITE].
+Symbolic execution is studied extensively and can be effective [CITE].
 [TALK ABOUT ONE OF THE CITATIONS]
-Despite positive results, symbolic execution has a fundamental limitation with respect to constraint solving.
-While symbolic execution engines perform well when constraints are linear, non-linear constraints greatly impact efficiency.
+Similar to white box fuzzing, symbolic execution has a fundamental limitation with respect to constraint solving.
+While symbolic execution engines perform well with linear constraints, non-linear constraints greatly impact efficiency.
 Furthermore, there is no autonomous symbolic execution engine.
@@
 As opposed to fuzzers, where input is randomly generated, this approach uses side channels to intelligently compose valid program input.
-Also, unlike symbolic execution, this approach is not hindered by constraint solver limitations and does not require any instrumentation or source code.
+Also, unlike symbolic execution, this approach is not hindered by constraint solvers' limitations and does not require any instrumentation or source code.
 
-\section{Methodology}
+\section{Methodology} \label{methodology}
 
-The majority of software programs accept input, perform transformations on the input and output the result.
+% add some high level intro for both finding input and psm stuff
+The majority of software programs accept input, perform transformations on the input and output results.
 However, most programs do not accept completely random input.
@@
 Such filters are snippets of code responsible for distinguishing between valid or good input (i.e.
-input the program was constructed to understand) and bad input (i.e.
+input the program was constructed to understand and accept) and bad input (i.e.
 input which is not useful or does not follow the desired format).
 Although validation mechanisms can be very sophisticated, they are often a combination of string comparisons and conditional statements.
-Our method exploits the presence (or lack) of these additional instructions during an execution to make inferences about the program's validation mechanism.
-Specifically, we test the program with different input and compare the number of retired user-land instructions executed.
-Assuming most input is invalid, our method observes differences in the number of instructions executed (from the norm) to detect input which does pass the validation mechanism.
-
-We present \tool{} an implementation of the method described above.
-Given a closed binary, \tool{} can build valid input, inform the tester about the type of acceptable input, and determine input buffer size.
+Our method exploits changes in the number of user-land instructions retired during multiple executions with varying inputs to make inferences about the program's validation mechanism.
+These changes reflect different execution paths of the program as a response to both valid and invalid input.
+Assuming most input is invalid, our method observes differences in the number of instructions executed to detect input which passes the validation mechanism.
 
@@
 
-In order to build valid input for a given program, \tool{} measures the number of user-land instructions retired during the program's execution using the `perf' utility.
-Specifically, \tool{} observes differences in the number of instructions executed as a result of different input.
-Generating valid input strings is a recursive two-step process.
-\tool{} works by identifying valid characters for each index \textit{i}, starting with \textit{i}=0 and incrementally building valid input strings using the discovered characters.
-Figure~\ref{fig:architecture} shows the overall architecture of \tool{}.
+Figure~\ref{fig:architecture} shows the overall architecture of the finding input component of \tool{}.
+The process is recursive and is composed of two steps.
 
-\subsubsection{Finding Valid Characters}
-
-The process of finding valid characters for a given index is composed of two stages: the testing stage and the filtering stage.
-During the testing stage, the program is executed with `perf' and being provided as input, all of the printable characters, one at a time.
-Once the program has processed the input and is waiting for more input, the program is stopped.
-The automatic interaction with the program is performed using `Expect', an extension to the `Tcl' scripting language [CITE]
-In our implementation, each printable character results in a different program execution.
-At the end of each execution, the number of instructions retired is outputted by `perf'.
-This information is saved in a dictionary as a tuple of the format ( <printable character> , <number of instructions> ).
-As mentioned earlier, the differences in the number of instructions between various printable characters reflect different execution paths taken during the validation process.
-
-The filtering stage is responsible for identifying characters which are highly likely to be part of a valid input string at a given index.
-Using the dictionary data obtained from the testing stage, the number of instructions retired are analyzed.
-First, the mode of the number of instructions executed is calculated.
-The filtering of possible valid characters for a given index \textit{i}, of an input string \textit{s}, is based on the assumption that the majority of printable characters \textit{c} do not belong at index \textit{i} of string \textit{s} (i.e.
-\textit{c} $\notin$ \textit{s[i]}).
-Therefore, valid characters are defined as any printable character for which the number of instructions retired is outside the mode.
-
+% printable character or all charactersls?
+The steps of the process for finding valid input are outlined below. Starting with an empty string, \textit{s}:
+\begin{enumerate}
+\item \tool{} executes the program once for every printable character and records the number of user-land instructions executed by the program with the given input.
+The input is a string concatenation of the string \textit{s} and the current printable character.
+The Unix `perf' utility is used for monitoring the number of instructions executed.
+Note that \tool{} does not record the total number of instructions but rather user-land instructions.
+The reason for this, as explained in Section \ref{bg_hardwarereg}, is that the total number of instructions executed can vary greatly from execution to execution due to operating system factors (e.g.
+branch prediction, cache).
+The Expect tool is used for automating the program interaction.
+\item Once all executions have been completed (one for each character), the mode number of instructions retired is computed across all executions.
+The mode is defined as the value that appears most often in the set of recordings.
+\item Next is the filtering stage which is responsible for identifying characters likely to be part of a valid input string at the current index.
+Under the assumption that most characters are not valid for the current index, \tool{} gathers all characters with the number of instructions executed outside the range of the mode +/- an epsilon value.
+These characters represent valid characters for the current index of the input string.
+The epsilon value accounts for small variations in the hardware counters.
 While the initial reaction is to expect a valid character to result in more instructions executed than a non-valid character, this is not always the case.
@@
 Such a scenario results in more instructions executed for invalid input than valid input.
+\item For each valid character \textit{c}, \tool{} forks with every child process repeating steps 1-4, appending \textit{c} to the end of the string \textit{s}.
+If no valid characters are found, the end of a valid input string has been reached and the string is recorded.
+If the current index is 0 and no valid characters have been found, the given binary has no predefined input commands.
+\end{enumerate}
 
-When no valid characters are identified as part of the filtering stage and the current index is greater than 0, the end of a valid input string has been reached.
-If the index is 0 and no valid characters have been found, the given binary has no predefined input commands.
+Figure~\ref{fig:string_builder} depicts the results of the above process for 5 time stages.
+As seen in the figure, at time \textit{T0}, \tool{} begins with an empty string.
+At time \textit{T1}, \tool{} finds valid characters for index 0 of the input and is forked into \textit{n} child processes where \textit{n} is the number of valid characters found.
+The process continues until all child processes are done.
+
+%\subsubsection{Finding Valid Characters}
+%The process of finding valid characters for a given index is composed of two stages: the testing stage and the filtering stage.
+%During the testing stage, the program is executed with `perf' and being provided as input, all of the printable characters, one at a time.
+%Once the program has processed the input and is waiting for more input, the program is stopped.
+%The automatic interaction with the program is performed using `Expect', an extension to the `Tcl' scripting language [CITE]
+%In our implementation, each printable character results in a different program execution.
+%At the end of each execution, the number of instructions retired is outputted by `perf'.
+%This information is saved in a dictionary as a tuple of the format ( <printable character> , <number of instructions> ).
+%As mentioned earlier, the differences in the number of instructions between various printable characters reflect different execution paths taken during the validation process.
+
+%The filtering stage is responsible for identifying characters which are highly likely to be part of a valid input string at a given index.
+%Using the dictionary data obtained from the testing stage, the number of instructions retired are analyzed.
+%First, the mode of the number of instructions executed is calculated.
+%The filtering of possible valid characters for a given index \textit{i}, of an input string \textit{s}, is based on the assumption that the majority of printable characters \textit{c} do not belong at index \textit{i} of string \textit{s} (i.e.
+\textit{c} $\notin$ \textit{s[i]}).
+%Therefore, valid characters are defined as any printable character for which the number of instructions retired is outside the mode.
+
+%Building valid input strings is a recursive process.
+%Starting with an empty string and index 0, \tool{} tries to find valid characters at each index and incrementally builds valid input strings.
+%More specifically, using the character finding technique described previously, relevant characters for index 0 are identified.
+%The task is forked \textit{n} times, where \textit{n} is the number of characters found at the current index (i.e. index 0).
+%Each child task is assigned one of the \textit{n} characters and is responsible for finding valid characters at the next index.
+%These steps are repeated until no valid characters are found, marking the end of the input string.
+%Figure~\ref{fig:string_builder} depicts the process.
 
@@
 \end{figure*}
-
-Building valid input strings is a recursive process.
-Starting with an empty string and index 0, \tool{} tries to find valid characters at each index and incrementally builds valid input strings.
-More specifically, using the character finding technique described previously, relevant characters for index 0 are identified.
-The task is forked \textit{n} times, where \textit{n} is the number of characters found at the current index (i.e. index 0).
-Each child task is assigned one of the \textit{n} characters and is responsible for finding valid characters at the next index.
-These steps are repeated until no valid characters are found, marking the end of the input string.
-Figure~\ref{fig:string_builder} depicts the process.
 
@@
 
+\subsection{Generating Protocol State Machine}
+
 \section{Evaluation and Results}
@@
 
-The current implementation uses `perf' to count instructions, which outputs only when the program executed exits.
-As a result, the tested program must be executed once with every new input string.
-This is not only a slow and inefficient process but also hinders the ability to crack programs which randomly generate passwords at execution time.
-In the future, we'd like to explore methods which allow sampling of instruction count during runtime.
-
 While the method presented in this paper has many advantages, it also has a few limitations.

### 2015-04-21 paper/main.tex
wrote more of the paper

@@
-\documentclass[10pt,twocolumn]{article}
-\usepackage{multicol}
+\documentclass{acm_proc_article-sp}
 \usepackage{graphicx}
@@
 
-\twocolumn[{%
-  \centering
-  \large \textbf{InputFinder: Reverse Engineering Closed Binaries via Side Channels} \\[2em]
-  Bogdan Copos, University of California, Davis\\
-  Praveen Murthy, Fujitsu Laboratories of America \\[4em]
-}]
+\title{InputFinder: Reverse Engineering Closed Binaries via Side Channels}
+\numberofauthors{2}
+\author{
+\alignauthor
+Bogdan Copos\\
+	\affaddr{University of California, Davis}\\
+	\email{bcopos@ucdavis.edu}
+\alignauthor
+Praveen Murthy\\
+	\affaddr{Fujitsu Laboratories of America}\\
+	\email{praveen.murthy@us.fujitsu.com}
+}
 
+\maketitle
 \begin{abstract}
 
-We propose a new approach for the autonomous generation of test suites for dynamic analysis of closed binaries and build a general tool called \tool{} for determining input for a closed binary and identifying the appropriate protocol (if any).
+We propose a new approach for the autonomous generation of test suites for dynamic analysis of closed x86 binaries and build a tool called \tool{} for determining input for closed binarie and identifying the appropriate protocol (if any).
 We show that (a) our technique can autonomously detect input, (b) one can easily leverage the results from our tool to build a sophisticated dynamic analysis tool, (c) the evaluation of \tool{} are on par and often exceed prior works designed for autonomous testing of programs, which we demonstrate by comparing \tool{}'s results with those of prior works on a set of binaries from XXX.
 Since \tool{}'s analysis
-%
+
+% OR
 
@@
 Despite previous success methods in identifying bugs, both techniques have discouraging limitations.
-In this paper, we present a (novel?) method for automatically generating test suites for closed binaries.
+In this paper, we present a (novel?) method for automatically generating test suites for closed x86 binaries.
 To assess its effectiveness, we have developed \tool{} and tested it against XXXX binaries from XXXXX.
@@
 Software testing is an expensive and labor intensive process.
-For many years, researchers have explored both static and dynamic methods for analyzing programs \cite{}.
+For many years, researchers have explored both static and dynamic methods for analyzing programs \cite{smartfuzzer}.
 While static analysis is thought to be more thorough, dynamic testing is often preferred due to its time efficiency.
@@
 This information can be used to learn about the program's internal components.
-Side channels have been widely explored by security researchers in order to exploit weaknesses in programs.
-In cryptography, side channel attacks exploit such information channels to defeat crypto systems.
-Recently, side channel attacks, such as OpenSSL's CREAM cache timing attack have made headlines. [CITE]
+Side channels have been widely explored by security researchers in order to exploit weaknesses in programs \cite{weinbergside, schindler2002combined, genkin2014rsa, genkinstealing}.
+In cryptography, side channel attacks exploit such information channels to defeat crypto systems \cite{zhou2005side, black2002black, okeya2006side}.
+Recently, side channel attacks, such as OpenSSL's CREAM cache timing attack have made headlines \cite{}.
 Similar to CREAM and other side channel attacks, the technique discussed in this paper exploits such information channels to generate valid input for binaries.
@@
 Each register, or counter, can be programmed to measure specific events such as cache misses, floating point operations, and even instructions executed.
+These hardware performance counters are often used by software developers to analyze their programs and improve on running time and efficiency.
+However, as we will show, such performance counters can also be exploited as side channels.
 Utilities, such as Linux/Unix `perf', take advantage of such hardware counters in order to provide developers with useful information about a program's performance.
-`perf' is a Unix tool which allows developers to specify the desired hardware event they wish to measure.
-Once events have been selected, `perf' interacts with kernel modules to program the hardware counters accordingly.
+`perf' allows developers to specify the desired hardware event they wish to measure.
+Once events have been selected, `perf' interacts with the appropriate kernel modules to program the hardware counters accordingly.
 In this work, we use `perf' to count the number of \textbf{user-land} instructions executed as a program is tested with various input.
@@
 
-\subsection{Execution/Dynamic Similarity} \label{bg_dynamicsim}
+\subsection{Execution/Dynamic Matching} \label{bg_dynamicsim}
+
+In many application areas, including clone detection and software debugging, there is a need for comparing executions of a program (or multiple versions of a program).
+One of the proposed approaches aimed at tackling this issue is dynamic matching.
+Dynamic matching works by first collecting execution traces of different executions.
+Such traces can collected through a number of ways.
+One option includes the use of Pin \cite{pintool}.
+Pin is a framework for building customized binary analysis tools and can be used to design a tool which creates a transcript of the instructions executed by a program.
+Once the traces are collected, they are analyze and mappings between instructions of different executions are produced.
+These mappings can be used to understand the relationship between two executions (or two versions of a program).
+In this work, we rely on the concept of dynamic matching and the Pin framework to develop a protocol state machine generator for closed binaries.
 
@@
 
+Dynamic analysis has been extensively studied in academia and is used widely in the industry by security specialists.
+Two of the main approaches to finding bugs are fuzzing and symbolic execution.
+
 Fuzzers are a popular choice amongst software testers.
-They are easy to use, implement, and have produced surprising results in the past.
+They are easy to use, easy to implement, and have produced surprising results in the past \cite{millerfuzz, millerfuzzrevisited}.
 There are two classes of fuzzers: black box fuzzers and white box fuzzers.
@@
 No program specifications or knowledge of the input format is available in such cases.
+Since black box fuzzing is completely random, its effectiveness is limited, especially with respect to coverage.
 On the other hand, in white box fuzzing, inputs are used to gather symbolic constraints which are in turn systematically negated and solved to produce more inputs.
+The approach introduced in this paper is similar to white box fuzzing in that information from the program's execution is used to craft valid input.
 White box fuzzing differs from black box fuzzing in that it uses the program's feedback to make inferences about valid input and use constraint solvers to generate more input.
-Both fuzzing approaches have limitations.
-Since black box fuzzing is completely random, its effectiveness is limited, especially with respect to fully exercising programs.
-White box fuzzing can attain better code coverage but can be slow and can be crippled by innate limitations of constraint solvers.
-[MAYBE TALK ABOUT FUZZING RESULTS, CITATIONS]
+White box fuzzing can attain better code coverage but can be slow and is crippled by innate limitations of constraint solvers.
+One of the works utilizing white box fuzzing is \cite{molnar2009dynamic} which introduces SmartFuzz, a dynamic test generation tool.
+SmartFuzz works by using symbolic execution to learn about the program's input and passing that information to a fuzzer to generate test cases.
+The main difference between our work and SmartFuzz is that our work does not use symbolic execution to gain insight into the format of the input.
+Similarly to SmartFuzz, the output of our work can be used by a fuzzer to generate test cases.
+Another related work is presented in \cite{smartfuzzer}, where the authors use a combination of static and dynamic analysis to build a smart fuzzer.
+Their work uses static analysis to collect preliminary information such as an approximation of the inter procedural control flow graph.
+Having gathered static information, their dynamic engine executes the program while monitoring dependencies between the input and the dynamic control flow path experienced by the program.
+Their work also takes advantage of a constraint solver that given a path of the program generates input which forces the program to follow the chosen path.
+Our work \tool{} differs from \cite{smartfuzzer} in several ways.
+In our proposed methods, we do not use neither static analysis nor constraint solver.
+Instead we rely only on information extracted from the execution of the program.
 
@@
 At the end of the execution, a constraint solver takes the constraints gathered and outputs valid program input.
-Symbolic execution is studied extensively and can be effective [CITE].
-[TALK ABOUT ONE OF THE CITATIONS]
+Symbolic execution is studied extensively and can be effective \cite{symbolicexecution, exesymex}.
 Similar to white box fuzzing, symbolic execution has a fundamental limitation with respect to constraint solving.
 While symbolic execution engines perform well with linear constraints, non-linear constraints greatly impact efficiency.
-Furthermore, there is no autonomous symbolic execution engine.
-All symbolic executioners today require some instrumentation of the program to determine the program's input buffer.
-Some symbolic execution engines even require source code, which is not always available to the tester.
+One of the most important works in this area is described in \cite{exesymex} which introduces EXE, a dynamic analysis tool capable of generating inputs that crash programs by running the programs with symbolic input.
+Unlike symbolic execution, the approach presented in this paper is not hindered by constraint solvers' limitations and does not require any instrumentation or source code.
+\tool{} does not execute the program with symbolic input but rather it uses concolic input and information about the execution of the program to craft valid input.
 
-Side channel attacks and citations\ldots
+There have also been numerous works studying and applying dynamic similarity.
+In \cite{blanketexec}, the authors propose blanket execution, a novel dynamic equivalence testing technique.
+Blanket execution is designed to collect information about the execution of functions, including side effects, which is then used for similarity testing.
+Both our approach and blanket execution use a dynamic approach to gather information about the program.
+However, our approach aims at finding similarities in executions of entire programs (not just functions) with various inputs.
+This allows us to discover differences in inputs which experience different control flow paths.
 
-The method presented in this paper takes advantage of hardware counters to observe program's behavior and automatically build valid input.
-As opposed to fuzzers, where input is randomly generated, this approach uses side channels to intelligently compose valid program input.
-Also, unlike symbolic execution, this approach is not hindered by constraint solvers' limitations and does not require any instrumentation or source code.
+Another area of research relevant to this paper is protocol reverse engineering.
+An important result in this area is \cite{prospex} which introduces a technique for extracting protocol specifications from packet captures and automatically generating input which can be used by a stateful fuzzer.
+There are two notable difference between Prospex and our work.
+While the end result is similar, \tool{} assumes no prior knowledge about the tested program and hence it does not utilize packet captures for protocol specification extraction.
+Instead, \tool{} infers the protocol through dynamic analysis by automatically finding input and using that to generate a protocol state machine.
+To date, there are no protocol reverse engineering techniques which assume no prior knowledge about the program's input.
+
+\section{Motivating Example???} \label{example}
 
@@
 % add some high level intro for both finding input and psm stuff
+\tool{} is capable of find input strings for a closed binary and generating a protocol state machine using the crafted input.
+
 The majority of software programs accept input, perform transformations on the input and output results.
@@
 
-Figure~\ref{fig:string_builder} depicts the results of the above process for 5 time stages.
-As seen in the figure, at time \textit{T0}, \tool{} begins with an empty string.
-At time \textit{T1}, \tool{} finds valid characters for index 0 of the input and is forked into \textit{n} child processes where \textit{n} is the number of valid characters found.
-The process continues until all child processes are done.
-
-%\subsubsection{Finding Valid Characters}
-%The process of finding valid characters for a given index is composed of two stages: the testing stage and the filtering stage.
-%During the testing stage, the program is executed with `perf' and being provided as input, all of the printable characters, one at a time.
-%Once the program has processed the input and is waiting for more input, the program is stopped.
-%The automatic interaction with the program is performed using `Expect', an extension to the `Tcl' scripting language [CITE]
-%In our implementation, each printable character results in a different program execution.
-%At the end of each execution, the number of instructions retired is outputted by `perf'.
-%This information is saved in a dictionary as a tuple of the format ( <printable character> , <number of instructions> ).
-%As mentioned earlier, the differences in the number of instructions between various printable characters reflect different execution paths taken during the validation process.
-
-%The filtering stage is responsible for identifying characters which are highly likely to be part of a valid input string at a given index.
-%Using the dictionary data obtained from the testing stage, the number of instructions retired are analyzed.
-%First, the mode of the number of instructions executed is calculated.
-%The filtering of possible valid characters for a given index \textit{i}, of an input string \textit{s}, is based on the assumption that the majority of printable characters \textit{c} do not belong at index \textit{i} of string \textit{s} (i.e.
-\textit{c} $\notin$ \textit{s[i]}).
-%Therefore, valid characters are defined as any printable character for which the number of instructions retired is outside the mode.
-
-%Building valid input strings is a recursive process.
-%Starting with an empty string and index 0, \tool{} tries to find valid characters at each index and incrementally builds valid input strings.
-%More specifically, using the character finding technique described previously, relevant characters for index 0 are identified.
-%The task is forked \textit{n} times, where \textit{n} is the number of characters found at the current index (i.e. index 0).
-%Each child task is assigned one of the \textit{n} characters and is responsible for finding valid characters at the next index.
-%These steps are repeated until no valid characters are found, marking the end of the input string.
-%Figure~\ref{fig:string_builder} depicts the process.
-
-\subsubsection{Building Input Strings}
-
 \begin{figure*}[t]
@@
 \end{figure*}
+
+Figure~\ref{fig:string_builder} depicts the results of the above process over the span of 5 time stages.
+As seen in the figure, at time \textit{T0}, \tool{} begins with an empty string.
+At time \textit{T1}, \tool{} finds valid characters for index 0 of the input and is forked into \textit{n} child processes where \textit{n} is the number of valid characters found.
+The process continues until all child processes are done.
 
@@
 
+% TODO
+% add diagram depicting process
+% add explanation
+
 \section{Evaluation and Results}
 
+\subsection{Finding Input, Input Size and Classifying Input}
 \tool{} was tested on [NUMBER] closed-source binaries to measure its effectiveness.
@@
 As described in the blog post, the cracking of the binary can take several hours and requires advanced skills and years of background knowledge.
-\tool{} cracked the binary in a matter of minutes automatically, proving itself as a useful tool not only finding valid input but also cracking passwords.
+\tool{} cracked the binary in a matter of minutes autonomously, proving itself as a useful tool not only finding valid input but also cracking passwords.
+
+\subsection{Protocol State Machine Generation}
+
+% TODO
 
@@
 
+% TODO
+% talk about finding arguments
+
 \section{Conclusions}
 
+\bibliographystyle{abbrv}
+\bibliography{inputfinder}  % sigproc.bib is the name of the Bibliography in this case
+
 \end{document}

### 2015-05-01 paper/main.tex
updated paper

@@
 \usepackage{graphicx}
+\usepackage{listings}
+\usepackage{hyperref}
 
@@
 }
+\maketitle
 
-\maketitle
 \begin{abstract}
-
-We propose a new approach for the autonomous generation of test suites for dynamic analysis of closed x86 binaries and build a tool called \tool{} for determining input for closed binarie and identifying the appropriate protocol (if any).
-We show that (a) our technique can autonomously detect input, (b) one can easily leverage the results from our tool to build a sophisticated dynamic analysis tool, (c) the evaluation of \tool{} are on par and often exceed prior works designed for autonomous testing of programs, which we demonstrate by comparing \tool{}'s results with those of prior works on a set of binaries from XXX.
-Since \tool{}'s analysis
-
-% OR
-
-The effectiveness of dynamic analysis techniques depends heavily on the completeness of the test suite applied during the analysis process.
+The effectiveness of many dynamic analysis techniques depends heavily on the completeness of the test suite applied during the analysis process.
 Test suites are often composed by developers and aim at covering (and testing) all of the functionality of a software system.
-However, test suites may still not be complete, if they exist at all.
-To date, only two methods exist for automatically generating test input for closed binaries: black-box fuzzing and symbolic execution.
-Despite previous success methods in identifying bugs, both techniques have discouraging limitations.
-In this paper, we present a (novel?) method for automatically generating test suites for closed x86 binaries.
-To assess its effectiveness, we have developed \tool{} and tested it against XXXX binaries from XXXXX.
-With a success rate of XXXX, \tool{}
-
+However, test suites may not be complete, if they exist at all.
+To date, only two methods exist for automatically generating test input for closed binaries: fuzzing and symbolic execution.
+Despite previous success methods in identifying bugs, both techniques have limitations.
+In this paper, we propose a new method for autonomously generating valid input and identifying protocols for closed x86 binaries.
+To assess its effectiveness, we have developed \tool{} and tested it against binaries from the DARPA Cyber Grand Challenge example set.
+Our evaluations show that ...
 \end{abstract}
@@
 \section{Introduction}
-
 %introduction
-Software testing is an expensive and labor intensive process.
+Software testing, particularly vulnerability testing, is an expensive and labor intensive process.
 For many years, researchers have explored both static and dynamic methods for analyzing programs \cite{smartfuzzer}.
@@
 The first component exploits side channels to build valid input for closed binaries.
-Specifically, as the program is provided various input, \tool{} records the number of instructions retired by the program for each input and compares the recorded value to learn the program's expected input.
+Specifically, as the program is given various input, \tool{} records the number of instructions retired during the program's execution for each input and compares the recorded value to learn the program's expected input.
+The same side channel is used by \tool{} to also find the expected input size and to categorize the expect input based on type.
 The second component uses the discovered input to build a protocol state machine associated with the tested program.
 This enables the generation of a more thorough test suite.
-
-%We show that by measuring the number of user-land instructions retired and comparing them between different executions of the program with various input, it is possible to learn the program's expected input.
-%We also present an implementation of our method, \tool{}.
 
@@
 Our method relies on observing changes in the number of instructions executed by programs during their input validation process to learn valid input strings.
+	\item We describe how our method can be used to not only find valid input strings, but also to determine the expect input size and classify input based on type (e.g.
+alphabetic, alphanumeric, numeric, etc.).
 	\item We develop a method for generating protocol state machines using the valid input previously found.
-	\item We implement the techniques presented and evaluate the tool using binaries published by DARPA as part of the 2014 Cyber Grand Challenge competition.
-Additionally, we compare our approach to the manual process used by a security specialist to crack a binary as part of a job interview.
+	\item We implement the techniques presented and evaluate the tool using binaries published by DARPA as part of the 2014 Cyber Grand Challenge competition \cite{darpacgc}.
+Additionally, we compare our approach to the manual process used by a security specialist to crack a binary as part of a job interview \cite{interviewbinary}.
 \end{itemize}
@@
 The paper is organized as follows.
-Section 2 covers background information.
-Section 3 discusses related works and differences with respect to the presented method.
-In Section 4 our work is explained in detail. First, the method for finding valid input is described. Then we introduce our method for building protocol state machines using the discovered input.
-Section 5 describes evaluation of the work presented and the results.
-In Section 6 we discuss future work and limitations of the presented work.
+Section \ref{background} covers background information relating to concepts utilized in this work.
+Section \ref{prevwork} discusses related works and highlights differences between the related works and the work presented here.
+The details of our approach are explained in Section \ref{methodology}.
+First, the method for finding valid input is described.
+Then we introduce our method for building protocol state machines using the discovered input.
+Section \ref{results} describes evaluation of the work presented and the results.
+In Section \ref{futurework}, we share future plans for our work and explain cases which the proposed method does not handle.
 
 \section{Background} \label{background}
-
 \subsection{Side Channels} \label{bg_sidechannels}
-
 Side channels are streams of information that can be retrieved from the hardware of a device running a given program.
-This information can be used to learn about the program's internal components.
-Side channels have been widely explored by security researchers in order to exploit weaknesses in programs \cite{weinbergside, schindler2002combined, genkin2014rsa, genkinstealing}.
+This information can be used to learn gather knowledge regarding the program's internal components.
+Side channels have been widely explored by security researchers in order to discover weaknesses in programs \cite{weinbergside, schindler2002combined, genkin2014rsa, genkinstealing}.
 In cryptography, side channel attacks exploit such information channels to defeat crypto systems \cite{zhou2005side, black2002black, okeya2006side}.
-Recently, side channel attacks, such as OpenSSL's CREAM cache timing attack have made headlines \cite{}.
+Recently, side channel attacks, such as OpenSSL's CREAM cache timing attack have made headlines \cite{creamssl}.
 Similar to CREAM and other side channel attacks, the technique discussed in this paper exploits such information channels to generate valid input for binaries.
-
 \subsection{Hardware counter registers} \label{bg_hardwarereg}
-
 Most modern microprocessors have a set of special-purpose registers used to count hardware events.
 Each register, or counter, can be programmed to measure specific events such as cache misses, floating point operations, and even instructions executed.
-These hardware performance counters are often used by software developers to analyze their programs and improve on running time and efficiency.
-However, as we will show, such performance counters can also be exploited as side channels.
+Such hardware performance counters are often used by software developers to analyze their programs and improve on running time and efficiency.
+However, in this paper we show how such performance counters can also be exploited as side channels.
 Utilities, such as Linux/Unix `perf', take advantage of such hardware counters in order to provide developers with useful information about a program's performance.
@@
 In this work, we use `perf' to count the number of \textbf{user-land} instructions executed as a program is tested with various input.
-The number of total instructions retired oscillates significantly for a binary from execution to execution.
+While other options are available for counting instructions executed, we chose `perf' for its efficiency and minimal overhead.
+The number of \textit{total} instructions retired oscillates significantly for a binary from execution to execution.
 This phenomena reflects fluctuations in instructions executed by the kernel depending on the state of the machine due to cache-misses and other factors.
 Since we rely on this count, such variations are problematic.
-However, experimentally, the number of user-land instructions only varies, on average, by one instruction, giving us a reliable counter.
-
+However, experimentally, the number of \textit{user-land} instructions only varies, on average, by one instruction, giving us a reliable counter.
 \subsection{Execution/Dynamic Matching} \label{bg_dynamicsim}
-
 In many application areas, including clone detection and software debugging, there is a need for comparing executions of a program (or multiple versions of a program).
@@
 Dynamic matching works by first collecting execution traces of different executions.
-Such traces can collected through a number of ways.
+Such traces can be collected through a number of ways.
 One option includes the use of Pin \cite{pintool}.
@@
 \section{Previous Work} \label{prevwork}
-
 Dynamic analysis has been extensively studied in academia and is used widely in the industry by security specialists.
-Two of the main approaches to finding bugs are fuzzing and symbolic execution.
+Two of the main approaches to vulnerability testing are fuzzing and symbolic execution.
 
@@
 If the program crashes, the input(s) are recorded and reported.
-No program specifications or knowledge of the input format is available in such cases.
+No prior program specifications or knowledge of the input format is available in such cases.
 Since black box fuzzing is completely random, its effectiveness is limited, especially with respect to coverage.
@@
 White box fuzzing can attain better code coverage but can be slow and is crippled by innate limitations of constraint solvers.
+
 One of the works utilizing white box fuzzing is \cite{molnar2009dynamic} which introduces SmartFuzz, a dynamic test generation tool.
-SmartFuzz works by using symbolic execution to learn about the program's input and passing that information to a fuzzer to generate test cases.
+SmartFuzz uses symbolic execution to learn about the program's input and passes that information to a fuzzer for test case generation.
 The main difference between our work and SmartFuzz is that our work does not use symbolic execution to gain insight into the format of the input.
-Similarly to SmartFuzz, the output of our work can be used by a fuzzer to generate test cases.
+Similar to SmartFuzz, the output of our work can be used by a fuzzer to generate test cases.
+
 Another related work is presented in \cite{smartfuzzer}, where the authors use a combination of static and dynamic analysis to build a smart fuzzer.
 Their work uses static analysis to collect preliminary information such as an approximation of the inter procedural control flow graph.
-Having gathered static information, their dynamic engine executes the program while monitoring dependencies between the input and the dynamic control flow path experienced by the program.
+Having gathered static information, their dynamic analysis engine executes the program while monitoring dependencies between the input and the dynamic control flow path experienced by the program.
 Their work also takes advantage of a constraint solver that given a path of the program generates input which forces the program to follow the chosen path.
 Our work \tool{} differs from \cite{smartfuzzer} in several ways.
-In our proposed methods, we do not use neither static analysis nor constraint solver.
+One difference is that our work uses neither static analysis nor constraint solvers.
 Instead we rely only on information extracted from the execution of the program.
@@
 At the end of the execution, a constraint solver takes the constraints gathered and outputs valid program input.
-Symbolic execution is studied extensively and can be effective \cite{symbolicexecution, exesymex}.
+Symbolic execution has been proven effective by previous works \cite{symbolicexecution, exesymex}.
 Similar to white box fuzzing, symbolic execution has a fundamental limitation with respect to constraint solving.
@@
 One of the most important works in this area is described in \cite{exesymex} which introduces EXE, a dynamic analysis tool capable of generating inputs that crash programs by running the programs with symbolic input.
-Unlike symbolic execution, the approach presented in this paper is not hindered by constraint solvers' limitations and does not require any instrumentation or source code.
+Unlike symbolic execution, the approach presented in our work is not hindered by constraint solvers and does not require any instrumentation or source code.
 \tool{} does not execute the program with symbolic input but rather it uses concolic input and information about the execution of the program to craft valid input.
@@
 
-\section{Motivating Example???} \label{example}
-
 \section{Methodology} \label{methodology}
-
 % add some high level intro for both finding input and psm stuff
-\tool{} is capable of find input strings for a closed binary and generating a protocol state machine using the crafted input.
-
+\tool{} is composed of two main modules.
+The first module is responsible for discovering input and properties of the expected input, including input size and the type of the input.
+The input can then be used by the second component to generate a protocol state machine, information which may then be used by a stateful fuzzer.
+\subsection{Finding Valid Input}
 The majority of software programs accept input, perform transformations on the input and output results.
@@
 
-\subsection{Finding Valid Input}
-
 \begin{figure*}[t]
@@
 \label{fig:architecture}
-\caption{}
+\caption{This figure depicts the architecture of the finding input process. There are three main parts: the execution engine, the valid character filter and the input string builder.}
 \end{figure*}
 
-Figure~\ref{fig:architecture} shows the overall architecture of the finding input component of \tool{}.
+Figure \ref{fig:architecture} shows the overall architecture of the finding input component of \tool{}.
 The process is recursive and is composed of two steps.
 
-% printable character or all charactersls?
-The steps of the process for finding valid input are outlined below. Starting with an empty string, \textit{s}:
+\begin{lstlisting}[caption={Pseudocode implementing the input finding process},label={findinputcode},escapeinside={@}{@}]
+string s = ""
+find_input(s, results) {
+    for char in printableChars:
+    #record number of instr executed per input
+    @\label{findinputcode_exec}@instrPerChar[char] = exec_bin_wperf(s+char)
+
+    #calculate mode for all instr counts	
+    @\label{findinputcode_mode}@mode = calculate_mode(instrPerChar)
+
+    #filtering stage: identify valid chars
+    for char in instrPerChar:
+      count = instrPerChar[char]
+      min = mode - epsilon
+      max = mode - epsilon
+      @\label{findinputcode_filter}@if not count in range(min, max):
+	validCharsList.append(char)
+
+    #base case
+    if len(validCharsList) == 0:
+       # reached end of input string
+       results.put(s)
+       return 0
+
+    #continue crafting input strings	
+    @\label{findinputcode_fork}@for valid_char in validCharsList:
+      find_input(s+validChar, results)
+}
+
+\end{lstlisting}
+
+The steps of the process for finding valid input are described in \autoref{findinputcode} and outlined below. Starting with an empty string, \textit{s}:
 \begin{enumerate}
@@
 The Expect tool is used for automating the program interaction.
+This step is shown in line \autoref{findinputcode_exec} of \autoref{findinputcode}.
 \item Once all executions have been completed (one for each character), the mode number of instructions retired is computed across all executions.
 The mode is defined as the value that appears most often in the set of recordings.
+This is done in line \autoref{findinputcode_mode} of \autoref{findinputcode}
 \item Next is the filtering stage which is responsible for identifying characters likely to be part of a valid input string at the current index.
-Under the assumption that most characters are not valid for the current index, \tool{} gathers all characters with the number of instructions executed outside the range of the mode +/- an epsilon value.
+Under the assumption that most characters are not valid for the current index, \tool{} gathers all characters with the number of instructions executed outside the range of the mode +/- an epsilon value, as shown in line \autoref{findinputcode_filter} of \autoref{findinputcode} .
 These characters represent valid characters for the current index of the input string.
@@
 Such a scenario results in more instructions executed for invalid input than valid input.
-\item For each valid character \textit{c}, \tool{} forks with every child process repeating steps 1-4, appending \textit{c} to the end of the string \textit{s}.
+\item For each valid character \textit{c}, \tool{} forks with every child process repeating steps 1-4, appending \textit{c} to the end of the string \textit{s} (starting at line \autoref{findinputcode_fork} of \autoref{findinputcode}).
 If no valid characters are found, the end of a valid input string has been reached and the string is recorded.
@@
 \label{fig:string_builder}
-\caption{}
+\caption{This diagram visualizes the input string being constructed throughout five consecutive time stages of the execution of \tool{}.
+Multiple outgoing arrows represent a process forking.
+This occurs when multiple valid characters are detected for a given index of the input string and, as a result, the finding input process forks to concurrently handle all of the discovered characters}
 \end{figure*}
 
-Figure~\ref{fig:string_builder} depicts the results of the above process over the span of 5 time stages.
+Figure~\ref{fig:string_builder} depicts the results of the above process over the span of five consecutive time stages.
 As seen in the figure, at time \textit{T0}, \tool{} begins with an empty string.
@@
 
-\subsection{Finding Desired Input Size}
-
+\subsubsection{Finding Desired Input Size}
 The same side channels can be used to detect the expected size of the program's input or even the program's input buffer.
@@
 
-\subsection{Classifying Input}
-
+\subsubsection{Categorizing Input}
 Our method is also capable of classifying the input discovered.
@@
 \subsection{Generating Protocol State Machine}
-
 % TODO
 % add diagram depicting process
-% add explanation
+Thus far, we have described how \tool{} discovers valid input for a given binary autonomously.
+However, some binaries have an implicit protocol.
+Consider, for example, a database management program.
+Without first selecting a database and perhaps even a table, some commands such as `insert' or `select' are useless and have no effect on any data.
+This is where the second component of \tool{} steps in.
+Specifically, this component identifies expected permutations of the discovered input strings.
 
-\section{Evaluation and Results}
+\begin{lstlisting}[caption={Pseudocode implementing the protocol state machine generator},label={lst:psmcode}]
+void main(inputs) {
+  for len in len(inputs):
+    permutations = get_permutation(len)
+    results = []
+    for p in permutations:
+      exec(p, 0, results)
+    compare_inputs(results, inputs)
+}
 
+void exec(cmds, i, results) {  
+  i = 0
+
+  # base case
+  if i == len(cmds):
+    results.append(cmds)
+    return 0
+
+  #finds args for input string at index i
+  initCmds = cmds[0:i]
+  cmd = cmds[i]
+  args = find_args(initCmds, cmd, i)
+
+  for arg in args:
+    cmds[i] = cmds[i] + arg  #concatenation
+    exec(cmds, i+1, results)
+}
+
+void compare_inputs(results, inputs) {
+  #organize results
+  for cmds in results:
+    lastCmd = cmds[len(cmds)]
+
+    trace = get_pin_trace(cmds)
+
+    traceForLastCmd = split_trace(cmds)
+    traceVector = dfs_cfg(trace)
+
+    traces[lastCmd][cmds] = traceVector
+
+  #compare traces for same input/cmd
+  # from different permutations
+  for cmd in cmds:
+    for cmdTraces in traces[cmd]:
+      clusters = compare(cmdTraces)
+
+   #find outliers in clusters
+   filter(clusters)
+}
+
+\end{lstlisting}
+
+Once input has been discovered, \tool{} constructs a protocol state machine.
+The architecture of this component is depicted in Figure \ref{fig:XXXXpsm_arch}.
+The process is describe below and shown as pseudocode in Listing \ref{lst:psmcode}.
+First, all possible permutations of all sizes from 1 to the number of input strings discovered are computed.
+Every permutation is passed through the dynamic analysis engine.
+The dynamic analysis execution engine first splits the permutation into individual input strings.
+Maintaining the original order of the inputs, the engine tests each string for the presence of expected arguments.
+For example, if the input string is `SET' and the program accepts input of the format `SET data', the analysis tries to determine that `data' is a valid argument.
+The arguments found are saved for later use.
+Once every input pertaining to a permutation has been tested for arguments, the inputs are passed through a instrumented binary for the collection of execution trace.
+The binary is instrumented with a Pin tool, developed by us, which logs the instructions executed as a function of time.
+Once all permutations of a given size have been executed, the dynamic similarity test begins.
+This test splits each execution trace into parts, each part being associated with an input string.
+%Next, the execution trace part associated with the last input string in the permutation is transformed into the depth-first search (DFS) string representation of its control flow graph (CFG).
+We refer to the execution trace of the input string `SET' as $\alpha(SET)$.
+For all permutations of the same size with the same last input string, the $ [last_cmd] $ are clustered.
+Given a permutation $p$ of size $s$ let $\alpha(p_s)$ represent the execution trace of the command at index $s$ of the permutation (i.e. last command of the permutation).
+It is natural to assume that given the correct (or expected) order the of input strings $1$ through $(i-1)$ of the permutation, the last input of the permutation experiences a different execution trace (or control flow path) compared to that of the same command in an incorrect permutation (assuming it is the same index in the permutation).
+This information is then used to determine, for every input string discovered, if there exist a permutations which affects its control flow path.
+If such a permutation is discovered, it it added to the protocol state machine.
+
+\begin{figure}[h!]
+\caption{Example of protocol state machine for one of the DARPA Cyber Grand Challenge example binaries.}
+\centering
+\includegraphics[width=0.5\textwidth]{protocoldiagram}
+\label{fig:psm_example}
+\end{figure}
+
+An example of a protocol state machine can be seen in Figure \ref{fig:psm_example}.
+Since there is no information about the states and their meaning, the states are represented by circles with generic names (e.g. $S0$).
+The arrows represent input strings which allow for transition between states.
+
+\section{Evaluation and Results} \label{results}
 \subsection{Finding Input, Input Size and Classifying Input}
@@
 [CONTINUE].
+%To further show the advantage of this method, we built a smart fuzzer which uses our method to determine input and protocol for a given binary.
+%We use the smart fuzzer and compare it to black box fuzzers symbolic execution engines with respect to code coverage over units of time.
+%Using such a comparison allows the reader to determine which method is more efficient while also performing well.
+
+% 7f... tic tac toe, asks for player names first so without skipping those steps it cannot find any valid input
 
@@
 To pass the challenge, the interviewer has to crack the secret passcode to the program.
-A blog post describes the manual process used to crack the password [CITE].
-The manual process is tedious.
+The blog post describes the manual process used to crack the password \cite{interviewbinary}.
+The manual process is tedious and it includes methods such as disassembling the binary, stepping through the instructions using a debugger, and reverse engineering the logic of the program.
 As described in the blog post, the cracking of the binary can take several hours and requires advanced skills and years of background knowledge.
@@
 \subsection{Protocol State Machine Generation}
+% TODO
+The evaluation of the protocol state machine generator was also done using the DARPA Cyber Grand Challenge binaries.
+However, out of the total of [NUMBER] binaries, only [NUMBER] had an implicit protocol.
+Our method was able to generate a correct protocol state machine for [NUMBER]\% of these binaries.
 
-% TODO
+The reason for failure include ...
 
-\section{Future Work and Limitations}
+\subsection{Case Study}
+One interesting and noteworthy case study is a binary which implements a simple protocol that lets the user call functions applying root64 and parcour schemes.
+The binary is built such that after sending the ``HELLO'' command, the binary will generate an authentication token which is echoed to the user and must be used for the ``AUTH" command.
+Without proper authentication, the ``SET" and ``CALL" commands have no effect.
+Before generating the protocol state machine, \tool{} analyzed the binary and discovered input strings.
+The inputs discovered are: ``HELLO", ``AUTH", ``SET", ``CALL" and ``BYE".
+These inputs are the only inputs implemented by the binary's protocol.
+The protocol state machine generator not only determined that authentication is necessary but also discovered that a user can properly authenticate themselves with a value of 0 or any special character if they omit the ``HELLO" command (Figure \ref{fig:psm_example}).
+This shows that the input finding approach in combination with the protocol state machine generator can help identify potential backdoors or unexpected control flow paths for a given program.
 
+\section{Future Work and Limitations} \label{futurework}
 While the method presented in this paper has many advantages, it also has a few limitations.
-For certain programs that allow any input, such as an echo server, our method is not any more efficient than a fuzzer.
-Additionally, our method cannot handle cases in which input is composed of multiple inter-dependent fields or when programs validate input fields out of order.
+For certain programs that allow any input, such as an echo server, our method is not any more efficient than a black box fuzzer.
+Additionally, our method cannot handle cases in which input is composed of multiple inter-dependent fields, such as when the value of a field determines the length of another field.
+Also, \tool{} cannot handle programs which validate input fields out of order or use any safe hashing algorithm for their input validation technique. 
+In the future, we would like to apply the input finding technique recursively.
+It is easy to imagine programs for which the second input varies based on the value of the first input.
+Also, as we've experienced during our evaluation, there exist programs which first request for a username, input which will not have any predefined valid values.
+It would be helpful to determine such cases, provide some input and continue the execution of the program, applying the input finding technique once again.
+However, there are challenges with this, such as determining when to stop.
 
-% TODO
-% talk about finding arguments
+With respect to the protocol state machine, currently our method uses the input finding technique to determine valid arguments.
+However, the protocol state machine does not currently handle situations in which there are no predefined arguments (e.g.
+given input ``CREATE'' the arguments could be predefined such as ``dog'' or random such as ``x'').
+Ultimately, we would like to generate some random arguments that remain constant for every input string in a given permutation.
+Moreover, our current approach is not capable of identifying equivalent states.
+As seen in Figure \ref{fig:psm_example}, the different authentication paths (use of ``HELLO" vs.
+omission of ``HELLO") lead to different states when in reality the two states are identical.
+Also, our method is based on the assumption that given a number of input strings, the majority of input string permutations are not valid and do not reflect the expected protocol.
+There may be cases where this assumption does not hold.
+In the future, we would like build onto our current approach for generating protocol state machines. 
+First, we want to be able to determine if two states are identical for a more accurate protocol state machine.
+Furthermore, we want to determine a better approach for identifying the valid sequence of input strings.
+One option would be to monitor other aspects of the program, particularly the \textit{.data}, \text{.bss}, and the heap during the executions.
+Changes in these may reflect changes of the state of the program.
 
 \section{Conclusions}
+This paper introduces two novel techniques for autonomously generating valid input strings for closed x86 binaries and for generating protocol state machine using the discovered input.
+For generating valid input strings, our method exploits the number of instructions executed to infer if a given character is valid for a given index of an input string.
+This process allows us to recursively and incrementally craft valid inputs.
+To further gain knowledge about the user interaction with the program, we test the program with the inputs generated and construct a protocol state machine which depicts the expected order of the input strings.
+
+We show that our proposed methods are efficient and successful by testing them on [NUMBER] DARPA binaries from the Cyber Grand Challenge \cite{darpacgc}. 
+Unlike previous works, our methods do not have any fundamental limitations (such as constraint solvers) and are robust enough to handle even binaries of other formats.
+
+We believe that the our techniques could be applied and used for a number of reasons including building test suites, discovering unwanted or unexpected control flow paths and even finding bugs (when used in combination with a fuzzer).
 
 \bibliographystyle{abbrv}
-\bibliography{inputfinder}  % sigproc.bib is the name of the Bibliography in this case
+\bibliography{inputfinder}
 

### 2015-05-04 paper/main.tex
updated paper

@@
 \def \tool {InputFinder}  
+\def \numbinaries {24}
 
@@
 \section{Methodology} \label{methodology}
-% add some high level intro for both finding input and psm stuff
+
 \tool{} is composed of two main modules.
@@
 \subsubsection{Finding Desired Input Size}
-The same side channels can be used to detect the expected size of the program's input or even the program's input buffer.
-Most programs are designed to read a predefined number of bytes from the source of input and store them in a buffer.
-Depending on the implementation, when a program is given more input than the size of the buffer, additional instructions may be executed.
-Regardless of the number of instructions executed, the behavior of the program changes.
-A program may exit when the number of provided bytes is larger than expected.
-Similarly, once the input buffer size limit has been reach, a program may check to see if the next byte is a newline or EOF.
-Some methods, such as C's `fgets()' function, read character by character [EXPLAIN]
-Our method observes such differences in the number of user-instructions retired as the program is given increasingly larger input.
+The same side channels can be used to detect the expected size of the program's input or even the size of the program's input buffer.
+Specifically, there is a change in the behavior of the program with respect to the number of instructions executed when the size of the input is larger than expected.
+Most programs are designed to read a predefined number of bytes from the source of input and store the read bytes in a buffer.
+A popular way of doing that is by using the \textit{fgets} or \textit{scanf} functions of the GNU C stdio library.
+As the input size grows, the number of instructions executed also increases linearly.
+This happens because additional instructions are executed to read more characters and store them into the buffer.
+However, when the input string is larger than the expected size the number of instructions executed no longer increases as a function of the input size.
+This behavior can be observed and exploited to find the expected input size, which may often also be the size of the input buffer.
+Our method observes such differences in the number of user-land instructions retired as the program is given increasingly larger input.
 
@@
 These two values may be equal or differ, depending on the program's logic.
-If the two values are equal [EQUAL]
-The two values can also differ because... [DIFFERENT]
+If the two values are equal, the program is executed with increasingly large input until a change in the number of executions is recorded.
+%TODO
+The two values may be equal if the function responsible for reading input ..
+If the values differ (as discussed earlier), the difference in the number of instructions executed is recorded and the program is executed, increasing the size of the input, until a change is detected.
+Instead of increasing the input size linearly, the input size is increased exponentially until a change is detected.
+Once the change is detected, the input size is decreased to find the exact size at which the change happens.
 
@@
 Once enough input strings have been identified, \tool{} uses the categorization data to generalize about the type of input the program accepts.
+This is done by iterating through the characters of the input strings discovered and determining which category they fall into.
 Such information is valuable especially in cases where the inputs discovered are all numeric or contain (one or more) special characters.
@@
 Once input has been discovered, \tool{} constructs a protocol state machine.
-The architecture of this component is depicted in Figure \ref{fig:XXXXpsm_arch}.
+The architecture of this component is depicted in Figure \ref{fig:psm_arch}.
 The process is describe below and shown as pseudocode in Listing \ref{lst:psmcode}.
@@
 This test splits each execution trace into parts, each part being associated with an input string.
-%Next, the execution trace part associated with the last input string in the permutation is transformed into the depth-first search (DFS) string representation of its control flow graph (CFG).
 We refer to the execution trace of the input string `SET' as $\alpha(SET)$.
-For all permutations of the same size with the same last input string, the $ [last_cmd] $ are clustered.
+For all permutations of the same size with the same last input string, the $\alpha(last_cmd)$ are clustered.
 Given a permutation $p$ of size $s$ let $\alpha(p_s)$ represent the execution trace of the command at index $s$ of the permutation (i.e. last command of the permutation).
-It is natural to assume that given the correct (or expected) order the of input strings $1$ through $(i-1)$ of the permutation, the last input of the permutation experiences a different execution trace (or control flow path) compared to that of the same command in an incorrect permutation (assuming it is the same index in the permutation).
-This information is then used to determine, for every input string discovered, if there exist a permutations which affects its control flow path.
-If such a permutation is discovered, it it added to the protocol state machine.
+It is natural to assume that given the correct (or expected) order the of input strings $1$ through $(i-1)$ of the permutation, the input $p_s$ experiences a different execution trace (or control flow path) compared to an incorrect permutation (assuming it is the same index in the permutation).
+This information is used to determine, for every input string discovered, if there exist a permutations which affects its control flow path.
+If such a permutation is discovered, it is added to the protocol state machine.
 
@@
 \subsection{Finding Input, Input Size and Classifying Input}
-\tool{} was tested on [NUMBER] closed-source binaries to measure its effectiveness.
-The results are summarized in Figure [FIGURE].
-Out of the total [NUMBER] binaries, [NUMBER2] had predefined accepted input strings.
-\tool{} found 100\% of the input strings for [NUMBER3]...
-[CONTINUE].
+\tool{} was tested on \numbinaries{} x86 binaries to measure its effectiveness from the DARPA Cyber Grand Challenge Example set \cite{darpacgc}.
+The results are summarized in Figure \ref{fig:findinputresults}.
+Analyzing the source code showed that out of the total \numbinaries{} binaries, 21 had predefined valid input strings.
+\tool{} found input strings for 14 of the 21 binaries and categorize the input correctly for 100\% of the 14 binaries.
+%TODO
+Our method was able to determine the input length for XXX of the binaries.
+
+\begin{figure}[h!]
+\caption{The results of our input finding method. `N/A' signifies that the binary did not have expect any predefined input strings but on the contrary accepted any input.}
+\centering
+\includegraphics[width=0.5\textwidth]{findinputresults}
+\label{fig:findinputresults}
+\end{figure}
 %To further show the advantage of this method, we built a smart fuzzer which uses our method to determine input and protocol for a given binary.
@@
 
-Additionally, our method was tested on a binary used for technical interviews.
+Additionally, our input finding method was tested on a binary used for technical interviews.
 The binary was designed as a reverse engineering challenge for interviewers.
@@
 The evaluation of the protocol state machine generator was also done using the DARPA Cyber Grand Challenge binaries.
-However, out of the total of [NUMBER] binaries, only [NUMBER] had an implicit protocol.
+However, out of the total of \numbinaries{} binaries, only [NUMBER] had an implicit protocol.
 Our method was able to generate a correct protocol state machine for [NUMBER]\% of these binaries.
@@
 For certain programs that allow any input, such as an echo server, our method is not any more efficient than a black box fuzzer.
-Additionally, our method cannot handle cases in which input is composed of multiple inter-dependent fields, such as when the value of a field determines the length of another field.
-Also, \tool{} cannot handle programs which validate input fields out of order or use any safe hashing algorithm for their input validation technique. 
-In the future, we would like to apply the input finding technique recursively.
+Additionally, our method cannot handle cases in which input is composed of multiple inter-dependent fields, such as when the value of a field determines the length of another field and when binaries validate input fields out of order.
+This is one reason why our method failed to detect input for some of the binaries tested.
+Another reason of failure is binaries which accept non printable input (e.g. hex characters, binary, etc.).
+
+In the future, we wish to improve our current method.
+First, we would like to extend it to handle any type of input.
+Furthermore, we would like to apply the input finding technique recursively.
 It is easy to imagine programs for which the second input varies based on the value of the first input.
@@
 It would be helpful to determine such cases, provide some input and continue the execution of the program, applying the input finding technique once again.
-However, there are challenges with this, such as determining when to stop.
+However, there are challenges with this, such as determining when to not apply the finding input method because the program has reached an already observed and tested state.
 
@@
 Furthermore, we want to determine a better approach for identifying the valid sequence of input strings.
-One option would be to monitor other aspects of the program, particularly the \textit{.data}, \text{.bss}, and the heap during the executions.
+One option would be to monitor other aspects of the program, particularly the \textit{.data}, \textit{.bss}, and the heap during the executions.
 Changes in these may reflect changes of the state of the program.
@@
 
-We show that our proposed methods are efficient and successful by testing them on [NUMBER] DARPA binaries from the Cyber Grand Challenge \cite{darpacgc}. 
+We show that our proposed methods are efficient and successful by testing them on \numbinaries{} DARPA binaries from the Cyber Grand Challenge \cite{darpacgc}. 
 Unlike previous works, our methods do not have any fundamental limitations (such as constraint solvers) and are robust enough to handle even binaries of other formats.

### 2015-05-04 paper/piechart.py

@@ -1,0 +1,20 @@
+from pylab import *
+
+# make a square figure and axes
+figure(1, figsize=(5,5))
+#ax = axes([0.1, 0.1, 0.8, 0.8])
+
+# The slices will be ordered and plotted counter-clockwise.
+labels = 'YES', 'N/A', 'NO'
+fracs = [58.34,  12.5, 29.16]
+explode=(0, 0, 0)
+
+pie(fracs, explode=explode, labels=labels,
+                autopct='%1.1f%%', shadow=False, startangle=90)
+                # The default startangle is 0, which would start
+                # the Frogs slice on the x-axis.  With startangle=90,
+                # everything is rotated counter-clockwise by 90 degrees,
+                # so the plotting starts on the positive y-axis.
+
+
+show()

### 2015-05-06 paper/main.tex
small changes to paper

@@
 The effectiveness of many dynamic analysis techniques depends heavily on the completeness of the test suite applied during the analysis process.
-Test suites are often composed by developers and aim at covering (and testing) all of the functionality of a software system.
+Test suites are often composed by developers and aim at testing all of the functionality of a software system.
 However, test suites may not be complete, if they exist at all.
@@
 %introduction
-Software testing, particularly vulnerability testing, is an expensive and labor intensive process.
+Vulnerability testing is an expensive and labor intensive process.
 For many years, researchers have explored both static and dynamic methods for analyzing programs \cite{smartfuzzer}.
@@
 Automated dynamic analysis techniques such as fuzzing and symbolic execution attempt to generate input for a given program.
-Currently these are the only two options available and both approaches present discouraging limitations.
+Currently these are the only two options available and both approaches present limitations.
 
@@
 Specifically, as the program is given various input, \tool{} records the number of instructions retired during the program's execution for each input and compares the recorded value to learn the program's expected input.
-The same side channel is used by \tool{} to also find the expected input size and to categorize the expect input based on type.
+The same side channel is used by \tool{} to also find the expected input size and to categorize the expected input based on type.
 The second component uses the discovered input to build a protocol state machine associated with the tested program.
@@
 	\item We develop a method for generating protocol state machines using the valid input previously found.
-	\item We implement the techniques presented and evaluate the tool using binaries published by DARPA as part of the 2014 Cyber Grand Challenge competition \cite{darpacgc}.
+	\item We implement the techniques presented as a proof of concept tool and evaluate it using binaries published by DARPA as part of the 2014 Cyber Grand Challenge competition \cite{darpacgc}.
 Additionally, we compare our approach to the manual process used by a security specialist to crack a binary as part of a job interview \cite{interviewbinary}.
@@
 Side channels are streams of information that can be retrieved from the hardware of a device running a given program.
-This information can be used to learn gather knowledge regarding the program's internal components.
+This information can be used to gather knowledge regarding the program's internal components.
 Side channels have been widely explored by security researchers in order to discover weaknesses in programs \cite{weinbergside, schindler2002combined, genkin2014rsa, genkinstealing}.
@@
 \subsection{Hardware counter registers} \label{bg_hardwarereg}
-Most modern microprocessors have a set of special-purpose registers used to count hardware events.
+Most modern microprocessors are equipped with a set of special-purpose registers used to count hardware events.
 Each register, or counter, can be programmed to measure specific events such as cache misses, floating point operations, and even instructions executed.
@@
 However, in this paper we show how such performance counters can also be exploited as side channels.
-Utilities, such as Linux/Unix `perf', take advantage of such hardware counters in order to provide developers with useful information about a program's performance.
-`perf' allows developers to specify the desired hardware event they wish to measure.
-Once events have been selected, `perf' interacts with the appropriate kernel modules to program the hardware counters accordingly.
-In this work, we use `perf' to count the number of \textbf{user-land} instructions executed as a program is tested with various input.
-While other options are available for counting instructions executed, we chose `perf' for its efficiency and minimal overhead.
+Utilities, such as Linux/Unix \textit{perf}, take advantage of such hardware counters in order to provide developers with useful information about a program's performance.
+\textit{perf} allows developers to specify the desired hardware event they wish to measure.
+Once events have been selected, \textit{perf} interacts with the appropriate kernel modules to program the hardware counters accordingly.
+In this work, we use \textit{perf} to count the number of \textbf{user-land} instructions executed as a program is tested with various input.
+While other options are available for counting instructions executed, we choose \textit{perf} for its efficiency and minimal overhead.
 The number of \textit{total} instructions retired oscillates significantly for a binary from execution to execution.
@@
 Such traces can be collected through a number of ways.
-One option includes the use of Pin \cite{pintool}.
-Pin is a framework for building customized binary analysis tools and can be used to design a tool which creates a transcript of the instructions executed by a program.
+One option includes the use of \textit{Pin} \cite{pintool}.
+\textit{Pin} is a framework for building customized binary analysis tools and can be used to design a tool which creates a transcript of the instructions executed by a program.
 Once the traces are collected, they are analyze and mappings between instructions of different executions are produced.
 These mappings can be used to understand the relationship between two executions (or two versions of a program).
-In this work, we rely on the concept of dynamic matching and the Pin framework to develop a protocol state machine generator for closed binaries.
+In this work, we rely on the concept of dynamic matching and the \textit{Pin} framework to develop a protocol state machine generator for closed binaries.
 
 \section{Previous Work} \label{prevwork}
-Dynamic analysis has been extensively studied in academia and is used widely in the industry by security specialists.
+Dynamic analysis is extensively studied in academia and is used widely in the industry by security specialists.
 Two of the main approaches to vulnerability testing are fuzzing and symbolic execution.
@@
 Fuzzers are a popular choice amongst software testers.
-They are easy to use, easy to implement, and have produced surprising results in the past \cite{millerfuzz, millerfuzzrevisited}.
+They are easy to use, easy to implement, and can produce surprising results \cite{millerfuzz, millerfuzzrevisited}.
 There are two classes of fuzzers: black box fuzzers and white box fuzzers.
@@
 Unlike symbolic execution, the approach presented in our work is not hindered by constraint solvers and does not require any instrumentation or source code.
-\tool{} does not execute the program with symbolic input but rather it uses concolic input and information about the execution of the program to craft valid input.
+\tool{} does not execute the program with symbolic input but rather it uses concrete input and information about the execution of the program to craft valid input.
 
@@
 Blanket execution is designed to collect information about the execution of functions, including side effects, which is then used for similarity testing.
-Both our approach and blanket execution use a dynamic approach to gather information about the program.
-However, our approach aims at finding similarities in executions of entire programs (not just functions) with various inputs.
+Another important work is \cite{differential} which introduces \textit{differential slicing} for the purpose of identifying flow and value differences across multiple executions of the same program for security applications.
+Both our approach and the previous approaches mentioned above use a dynamic approach to gather information about the program.
+However, our approach aims at finding differences in executions of a given program in order to make inferences about the underlying protocol expected by the program.
 This allows us to discover differences in inputs which experience different control flow paths.
@@
 The first module is responsible for discovering input and properties of the expected input, including input size and the type of the input.
-The input can then be used by the second component to generate a protocol state machine, information which may then be used by a stateful fuzzer.
+The input is then be used by the second component to generate a protocol state machine.
+The protocol state machine can be used by third party tools such as stateful fuzzers for even better code coverage.
 \subsection{Finding Valid Input}
@@
 The input is a string concatenation of the string \textit{s} and the current printable character.
-The Unix `perf' utility is used for monitoring the number of instructions executed.
+The Unix \textit{perf} utility is used for monitoring the number of instructions executed.
 Note that \tool{} does not record the total number of instructions but rather user-land instructions.
@@
 The mode is defined as the value that appears most often in the set of recordings.
-This is done in line \autoref{findinputcode_mode} of \autoref{findinputcode}
+This is done in line \autoref{findinputcode_mode} of \autoref{findinputcode}.
 \item Next is the filtering stage which is responsible for identifying characters likely to be part of a valid input string at the current index.
-Under the assumption that most characters are not valid for the current index, \tool{} gathers all characters with the number of instructions executed outside the range of the mode +/- an epsilon value, as shown in line \autoref{findinputcode_filter} of \autoref{findinputcode} .
+Under the assumption that most characters are not valid for the current index, \tool{} gathers all characters with the number of instructions executed outside the range of the mode +/- an epsilon value, as shown in line \autoref{findinputcode_filter} of \autoref{findinputcode}.
 These characters represent valid characters for the current index of the input string.
@@
 Such a scenario results in more instructions executed for invalid input than valid input.
-\item For each valid character \textit{c}, \tool{} forks with every child process repeating steps 1-4, appending \textit{c} to the end of the string \textit{s} (starting at line \autoref{findinputcode_fork} of \autoref{findinputcode}).
+\item For each valid character \textit{c}, \tool{} forks, with every child process repeating steps 1-4, starting with $ c + s $ as the initial input string, where $+$ signifies concatenation (starting at line \autoref{findinputcode_fork} of \autoref{findinputcode}).
 If no valid characters are found, the end of a valid input string has been reached and the string is recorded.
@@
 At time \textit{T1}, \tool{} finds valid characters for index 0 of the input and is forked into \textit{n} child processes where \textit{n} is the number of valid characters found.
-The process continues until all child processes are done.
+This continues until all child processes are done.
 
@@
 Most programs are designed to read a predefined number of bytes from the source of input and store the read bytes in a buffer.
-A popular way of doing that is by using the \textit{fgets} or \textit{scanf} functions of the GNU C stdio library.
+A popular way of doing that is by using the \textit{fgets()} or \textit{scanf()} functions of the GNU C stdio library.
 As the input size grows, the number of instructions executed also increases linearly.
@@
 These two values may be equal or differ, depending on the program's logic.
-If the two values are equal, the program is executed with increasingly large input until a change in the number of executions is recorded.
-%TODO
-The two values may be equal if the function responsible for reading input ..
+The two values may be equal if, for example, the program uses the system call \textit{read()} directly to read data from the input file descriptor.
+Since we are measuring the number of user-land instructions retired (not also instructions executed by the kernel), the instructions executed by the \textit{read} functions are not included in our counter so there will be no linear increase in the number of instructions executed as the input size is increased.
+If the two values are equal, the program is executed with increasingly large input until a change in the number of executions is recorded due to the program logic.
+There may not always be a change (depending on the program logic when the input size exceeds the expected size), however, we note that this scenario is not very interesting since the expected size of the input may also be obtained by executing the program with \textit{strace}.
 If the values differ (as discussed earlier), the difference in the number of instructions executed is recorded and the program is executed, increasing the size of the input, until a change is detected.
@@
 Once the change is detected, the input size is decreased to find the exact size at which the change happens.
+It is also possible that the number of instructions executed oscillates between two (or more values).
+One example is a palindrome finder.
+The same underlying concept can be applied to such cases.
 
@@
 Consider, for example, a database management program.
-Without first selecting a database and perhaps even a table, some commands such as `insert' or `select' are useless and have no effect on any data.
+Without first selecting a database and perhaps even a table, some commands such as \textit{INSERT} or \textit{SELECT} are useless and have no effect on any data.
 This is where the second component of \tool{} steps in.
-Specifically, this component identifies expected permutations of the discovered input strings.
+Specifically, this component identifies expected order (or permutation) of the discovered input strings.
 
@@
 
+\begin{figure}[h!]
+\caption{Architecture of the protocol state machine generator}
+\centering
+\includegraphics[width=0.5\textwidth]{psmarch}
+\label{fig:psm_arch}
+\end{figure}
+
 Once input has been discovered, \tool{} constructs a protocol state machine.
@@
 Maintaining the original order of the inputs, the engine tests each string for the presence of expected arguments.
-For example, if the input string is `SET' and the program accepts input of the format `SET data', the analysis tries to determine that `data' is a valid argument.
+For example, if the input string is \textit{SET} and the program accepts input of the format \textit{SET data}, the analysis tries to determine that `data' is a valid argument.
 The arguments found are saved for later use.
-Once every input pertaining to a permutation has been tested for arguments, the inputs are passed through a instrumented binary for the collection of execution trace.
-The binary is instrumented with a Pin tool, developed by us, which logs the instructions executed as a function of time.
+Once every input string of a given permutation has been tested for arguments, the inputs are passed through a instrumented binary for the collection of execution trace.
+The binary is instrumented with a \textit{Pin} tool, developed by us, which logs the instructions executed as a function of time.
 Once all permutations of a given size have been executed, the dynamic similarity test begins.
 This test splits each execution trace into parts, each part being associated with an input string.
-We refer to the execution trace of the input string `SET' as $\alpha(SET)$.
-For all permutations of the same size with the same last input string, the $\alpha(last_cmd)$ are clustered.
+We refer to the execution trace of the input string \textit{SET} as $\alpha(SET)$.
+For all permutations of the same size with the same last input string, the $\alpha(lastCmd)$ are clustered.
 Given a permutation $p$ of size $s$ let $\alpha(p_s)$ represent the execution trace of the command at index $s$ of the permutation (i.e. last command of the permutation).
-It is natural to assume that given the correct (or expected) order the of input strings $1$ through $(i-1)$ of the permutation, the input $p_s$ experiences a different execution trace (or control flow path) compared to an incorrect permutation (assuming it is the same index in the permutation).
+It is natural to assume that given the correct (or expected) order the of input strings $1$ through $(i-1)$ of the permutation, the input $p_s$ experiences a different execution trace (or control flow path) compared to an incorrect permutation (assuming $p_s$ is at the same index in the permutation).
 This information is used to determine, for every input string discovered, if there exist a permutations which affects its control flow path.
@@
 One interesting and noteworthy case study is a binary which implements a simple protocol that lets the user call functions applying root64 and parcour schemes.
-The binary is built such that after sending the ``HELLO'' command, the binary will generate an authentication token which is echoed to the user and must be used for the ``AUTH" command.
-Without proper authentication, the ``SET" and ``CALL" commands have no effect.
+The binary is built such that after sending the \textit{HELLO} command, the binary will generate an authentication token which is echoed to the user and must be used for the \textit{AUTH} command.
+Without proper authentication, the \textit{SET} and \textit{CALL} commands have no effect.
 Before generating the protocol state machine, \tool{} analyzed the binary and discovered input strings.
-The inputs discovered are: ``HELLO", ``AUTH", ``SET", ``CALL" and ``BYE".
+The inputs discovered are: $HELLO, AUTH, SET, CALL$ and $BYE$.
 These inputs are the only inputs implemented by the binary's protocol.
-The protocol state machine generator not only determined that authentication is necessary but also discovered that a user can properly authenticate themselves with a value of 0 or any special character if they omit the ``HELLO" command (Figure \ref{fig:psm_example}).
+The protocol state machine generator not only determined that authentication is necessary but also discovered that a user can properly authenticate themselves with a value of 0 or any special character if they omit the \textit{HELLO} command (Figure \ref{fig:psm_example}).
 This shows that the input finding approach in combination with the protocol state machine generator can help identify potential backdoors or unexpected control flow paths for a given program.
@@
 However, the protocol state machine does not currently handle situations in which there are no predefined arguments (e.g.
-given input ``CREATE'' the arguments could be predefined such as ``dog'' or random such as ``x'').
+given input \textit{CREATE} the arguments could be predefined such as \textit{dog} or random such as ``x'').
 Ultimately, we would like to generate some random arguments that remain constant for every input string in a given permutation.
 Moreover, our current approach is not capable of identifying equivalent states.
-As seen in Figure \ref{fig:psm_example}, the different authentication paths (use of ``HELLO" vs.
-omission of ``HELLO") lead to different states when in reality the two states are identical.
+As seen in Figure \ref{fig:psm_example}, the different authentication paths (use of \textit{HELLO} vs.
+omission of \textit{HELLO}) lead to different states when in reality the two states are identical.
 Also, our method is based on the assumption that given a number of input strings, the majority of input string permutations are not valid and do not reflect the expected protocol.

### 2015-05-07 paper/main.tex
updated paper after comments from Praveen

@@
 \usepackage{hyperref}
+\usepackage{algorithmicx}
+\usepackage{algpseudocode}
+\usepackage{algorithm}
 
@@
 \begin{abstract}
-The effectiveness of many dynamic analysis techniques depends heavily on the completeness of the test suite applied during the analysis process.
+The effectiveness of many dynamic program analysis techniques depends heavily on the completeness of the test suite applied during the analysis process.
 Test suites are often composed by developers and aim at testing all of the functionality of a software system.
@@
 To date, only two methods exist for automatically generating test input for closed binaries: fuzzing and symbolic execution.
-Despite previous success methods in identifying bugs, both techniques have limitations.
+Despite previous successes of these methods in identifying bugs, both techniques have limitations.
 In this paper, we propose a new method for autonomously generating valid input and identifying protocols for closed x86 binaries.
@@
 For many years, researchers have explored both static and dynamic methods for analyzing programs \cite{smartfuzzer}.
-While static analysis is thought to be more thorough, dynamic testing is often preferred due to its time efficiency.
-However, dynamic testing heavily depends on availability of input or test suites.
+While static analysis can be more thorough, dynamic testing is often preferred due to its time efficiency.
+However, dynamic testing heavily depends on the availability of appropriate inputs or test suites.
 
 %motivation
-Without complete test suites, the program cannot be properly tested and bugs may be omitted.
+Without complete test suites, the program cannot be properly tested and bugs may be omitted during testing.
 While valid input can be learned from software documentation or source code, when such information is not available or complete, other methods must be used.
@@
 Currently these are the only two options available and both approaches present limitations.
+Fuzzing done randomly with no inference attempted of the program's behavior cannot be effective as the input space is usually too large.
+Symbolic execution has traditionally been used with manual intervention to identify potential symbolic variables to begin exploration with, a luxury we do not have when analyzing closed binaries autonomously, in addition to scalability issues and limitations of decision procedures that can be used to solve the constraints.
 
-In this paper, we introduce a new method for generating valid user-input for closed binaries. 
+In this paper, we introduce a new method, called InputFinder, for generating valid user-input for closed binaries. 
+This new method could be useful in autonomous vulnerability scanning systems, or for reverse engineering unknown binaries.
 Our method is composed of two main components.
 The first component exploits side channels to build valid input for closed binaries.
-Specifically, as the program is given various input, \tool{} records the number of instructions retired during the program's execution for each input and compares the recorded value to learn the program's expected input.
+Specifically, as the program is given various input, \tool{} records the number of instructions retired during the program's execution for each input and uses a statistical measure to infer the program's expected input.
+The number of instructions retired differs from the number of instructions executed by a processing unit.
+Because of the out-of-order CPU pipeline, a CPU may execute more instructions than necessary (e.g. as a result of wrong branch prediction).
+The instructions retired represent the subset of instructions executed which leave the \textit{Retirement Unit} once execution has been deemed correct (i.e.
+the instructions which actually impact the program).
 The same side channel is used by \tool{} to also find the expected input size and to categorize the expected input based on type.
@@
 	\item We introduce a new approach to automatically generate input valid strings for unknown, closed-source binaries by leveraging hardware instruction counters.
-Our method relies on observing changes in the number of instructions executed by programs during their input validation process to learn valid input strings.
+Our method relies on observing changes in the number of instructions retired by programs during their input validation process to learn valid input strings.
 	\item We describe how our method can be used to not only find valid input strings, but also to determine the expect input size and classify input based on type (e.g.
@@
 Most modern microprocessors are equipped with a set of special-purpose registers used to count hardware events.
-Each register, or counter, can be programmed to measure specific events such as cache misses, floating point operations, and even instructions executed.
-Such hardware performance counters are often used by software developers to analyze their programs and improve on running time and efficiency.
+Each register, or counter, can be programmed to measure specific events such as cache misses, floating point operations, and even instructions retired.
+Such hardware performance counters are often used by software developers to analyze their programs and improve the running time and efficiency.
 However, in this paper we show how such performance counters can also be exploited as side channels.
 Utilities, such as Linux/Unix \textit{perf}, take advantage of such hardware counters in order to provide developers with useful information about a program's performance.
-\textit{perf} allows developers to specify the desired hardware event they wish to measure.
+The \textit{perf} utility allows developers to specify the desired hardware event they wish to measure.
 Once events have been selected, \textit{perf} interacts with the appropriate kernel modules to program the hardware counters accordingly.
-In this work, we use \textit{perf} to count the number of \textbf{user-land} instructions executed as a program is tested with various input.
-While other options are available for counting instructions executed, we choose \textit{perf} for its efficiency and minimal overhead.
-The number of \textit{total} instructions retired oscillates significantly for a binary from execution to execution.
-This phenomena reflects fluctuations in instructions executed by the kernel depending on the state of the machine due to cache-misses and other factors.
+In this work, we use \textit{perf} to count the number of \textbf{user-land} instructions retired as a program is tested with various input.
+\textbf{User-land} instructions are instructions of a program which run outside the operating system's kernel.
+While other options are available for counting instructions retired, we choose \textit{perf} for its efficiency and minimal overhead.
+The number of \textit{total} instructions retired varies significantly for a binary from execution to execution \cite{weaver2008can}.
+This phenomenon reflects fluctuations in instructions retired by the kernel depending on the state of the machine due to cache-misses and other factors.
 Since we rely on this count, such variations are problematic.
@@
 Such traces can be collected through a number of ways.
-One option includes the use of \textit{Pin} \cite{pintool}.
-\textit{Pin} is a framework for building customized binary analysis tools and can be used to design a tool which creates a transcript of the instructions executed by a program.
+One option includes the use of a tool such as \textit{Pin} \cite{pintool}.
+\textit{Pin} is a framework for building customized binary analysis tools and can be used to design a tool which creates a transcript of the instructions retired by a program.
 Once the traces are collected, they are analyze and mappings between instructions of different executions are produced.
@@
 Since black box fuzzing is completely random, its effectiveness is limited, especially with respect to coverage.
-On the other hand, in white box fuzzing, inputs are used to gather symbolic constraints which are in turn systematically negated and solved to produce more inputs.
-The approach introduced in this paper is similar to white box fuzzing in that information from the program's execution is used to craft valid input.
-White box fuzzing differs from black box fuzzing in that it uses the program's feedback to make inferences about valid input and use constraint solvers to generate more input.
+
+Symbolic execution is another method used for generating valid input.
+When a program analyzer symbolically executes a program, it does not use concrete values.
+Instead, the input is represented as a mathematical formula composed of symbols and equations resulting from the constraints (related to the input) gathered during the program's execution.
+Specifically, the tester first marks some memory locations as symbolic.
+This can also be done automatically by, for example, marking the buffer used in \textit{read()} system call as symbolic.
+However, this does not always produce good results, especially if the program reads one character at a time and then copies the read character into a second buffer.
+The program is then executed and the tester enters some input.
+Regardless of the value entered, the program replaces the input with a symbol.
+As the input flows through the program and encounters conditional statements, the symbolic value of the input changes to reflect the faced constraints.
+At the end of the execution, a constraint solver takes the formulat representing the symbolic value of the input and attempts to solve it in order to obtain valid program input.
+Symbolic execution has been proven effective by previous works \cite{symbolicexecution, exesymex}.
+Similar to white box fuzzing, symbolic execution has a fundamental limitation with respect to constraint solving.
+While symbolic execution engines perform well with linear constraints, non-linear constraints greatly impact efficiency.
+One of the most important works in this area is described in \cite{exesymex} which introduces EXE, a dynamic analysis tool capable of generating inputs that crash programs by running the programs with symbolic input.
+Unlike the EXE symbolic execution engine, the SAGE tool and SmartFuzz, the approach presented in our work is not hindered by constraint solvers.
+Additionally, compared to EXE, our method does not require any instrumentation or source code.
+\tool{} does not execute the program with symbolic input but rather it bootstraps itself by building up a set of valid inputs (from an initially empty set) by gathering information about the execution of a program.
+
+White box fuzzing leverages the symbolic execution traces of a given binary to produce more inputs.
+One of the many previous works is \cite{godefroid2008automated} which introduces SAGE, an automated white box fuzzer.
+White box fuzzing differs from black box fuzzing in that it uses the program's feedback to make inferences about valid input and uses constraint solvers to generate more input.
 White box fuzzing can attain better code coverage but can be slow and is crippled by innate limitations of constraint solvers.
+The only similarity between the approach introduced in this paper and white box fuzzing is that both approaches use information from the program's execution to craft valid input.
+However, our method does not use symbolic execution.
 
@@
 Having gathered static information, their dynamic analysis engine executes the program while monitoring dependencies between the input and the dynamic control flow path experienced by the program.
-Their work also takes advantage of a constraint solver that given a path of the program generates input which forces the program to follow the chosen path.
+Their work also takes advantage of a constraint solver that, given a path of the program, generates input which forces the program to follow the chosen path.
 Our work \tool{} differs from \cite{smartfuzzer} in several ways.
 One difference is that our work uses neither static analysis nor constraint solvers.
-Instead we rely only on information extracted from the execution of the program.
-
-Symbolic execution is another method used for generating valid input.
-In symbolic execution, an interpreter gathers constraints related to input throughout the program's execution, rather than using the provided input.
-At the end of the execution, a constraint solver takes the constraints gathered and outputs valid program input.
-Symbolic execution has been proven effective by previous works \cite{symbolicexecution, exesymex}.
-Similar to white box fuzzing, symbolic execution has a fundamental limitation with respect to constraint solving.
-While symbolic execution engines perform well with linear constraints, non-linear constraints greatly impact efficiency.
-One of the most important works in this area is described in \cite{exesymex} which introduces EXE, a dynamic analysis tool capable of generating inputs that crash programs by running the programs with symbolic input.
-Unlike symbolic execution, the approach presented in our work is not hindered by constraint solvers and does not require any instrumentation or source code.
-\tool{} does not execute the program with symbolic input but rather it uses concrete input and information about the execution of the program to craft valid input.
+Instead we rely only on information extracted from the execution of the uninstrumented program.
 
@@
 Instead, \tool{} infers the protocol through dynamic analysis by automatically finding input and using that to generate a protocol state machine.
-To date, there are no protocol reverse engineering techniques which assume no prior knowledge about the program's input.
+To date, we do not know of any protocol reverse engineering techniques which work without any prior knowledge about the program's input.
 
@@
 These changes reflect different execution paths of the program as a response to both valid and invalid input.
-Assuming most input is invalid, our method observes differences in the number of instructions executed to detect input which passes the validation mechanism.
+Assuming most input is invalid, our method observes differences in the number of instructions retired to detect input which passes the validation mechanism.
 
@@
 \includegraphics[height=2.5in,width=6.5in]{architecture.png}
+\caption{This figure depicts the architecture of the finding input process. There are three main parts: the execution engine, the valid character filter and the input string builder.}
 \label{fig:architecture}
-\caption{This figure depicts the architecture of the finding input process. There are three main parts: the execution engine, the valid character filter and the input string builder.}
 \end{figure*}
@@
 Figure \ref{fig:architecture} shows the overall architecture of the finding input component of \tool{}.
-The process is recursive and is composed of two steps.
 
-\begin{lstlisting}[caption={Pseudocode implementing the input finding process},label={findinputcode},escapeinside={@}{@}]
-string s = ""
-find_input(s, results) {
-    for char in printableChars:
-    #record number of instr executed per input
-    @\label{findinputcode_exec}@instrPerChar[char] = exec_bin_wperf(s+char)
+\begin{algorithm}
+\caption{Pseudocode implementing the input finding process}
+\label{findinputcode}
+\begin{algorithmic}[1]
+\State $s \leftarrow ``"$ \Comment{starting with an empty string}
+\Procedure{FindInput}{$s$}
+\For{i=0, \Call{size}{$printableChars$}} %\Comment{for all printable character}
+\State char = printableChars[$i$]
+\State insRetired = \Call{ExecBin}{$s+char$} \label{execute}%\Comment{execute binary with input $s+char$}
+\State insPerInput[$char$] = insRetired %\Comment{record number of instructions retired}
+\EndFor
+\State mode = \Call{CalcMode}{$insPerInput$} \label{mode}%\Comment{calculate mode across all executions}
+\For{i=0, \Call{size}{$printableChars$}} \label{filterbegin}%\Comment{for all printable character}
+\State char = printableChars[$i$]
+\State count = insPerInput[$char$]
+\If{mode-epsilon < count < mode+epsilon}
+\State ignore
+\Else
+\State validCharList += char
+\EndIf
+\EndFor \label{filterend}
+\If{no valid characters found}
+\State end of string reached
+\State return
+\Else
+\For{i=0, \Call{size}{$validCharList$}} \label{fork}
+\State validChar = validCharList[$i$]
+\State \Call{FindInput}{$s+validChar$}
+\EndFor
+\EndIf
+\EndProcedure
+\end{algorithmic}
+\end{algorithm}
 
-    #calculate mode for all instr counts	
-    @\label{findinputcode_mode}@mode = calculate_mode(instrPerChar)
-
-    #filtering stage: identify valid chars
-    for char in instrPerChar:
-      count = instrPerChar[char]
-      min = mode - epsilon
-      max = mode - epsilon
-      @\label{findinputcode_filter}@if not count in range(min, max):
-	validCharsList.append(char)
-
-    #base case
-    if len(validCharsList) == 0:
-       # reached end of input string
-       results.put(s)
-       return 0
-
-    #continue crafting input strings	
-    @\label{findinputcode_fork}@for valid_char in validCharsList:
-      find_input(s+validChar, results)
-}
-
-\end{lstlisting}
-
-The steps of the process for finding valid input are described in \autoref{findinputcode} and outlined below. Starting with an empty string, \textit{s}:
+% \algref{findinputcode}{line}
+The steps of the process for finding valid input are described in algorithm \ref{findinputcode} and outlined below. Starting with an empty string, \textit{s}:
 \begin{enumerate}
-\item \tool{} executes the program once for every printable character and records the number of user-land instructions executed by the program with the given input.
+\item \tool{} executes the program once for every printable character and records the number of user-land instructions retired by the program with the given input.
 The input is a string concatenation of the string \textit{s} and the current printable character.
-The Unix \textit{perf} utility is used for monitoring the number of instructions executed.
+The Unix \textit{perf} utility is used for monitoring the number of instructions retired.
 Note that \tool{} does not record the total number of instructions but rather user-land instructions.
-The reason for this, as explained in Section \ref{bg_hardwarereg}, is that the total number of instructions executed can vary greatly from execution to execution due to operating system factors (e.g.
+The reason for this, as explained in Section \ref{bg_hardwarereg}, is that the total number of instructions retired can vary greatly from execution to execution due to operating system factors (e.g.
 branch prediction, cache).
 The Expect tool is used for automating the program interaction.
-This step is shown in line \autoref{findinputcode_exec} of \autoref{findinputcode}.
+This step is shown in line \ref{execute} of algorithm \ref{findinputcode}
 \item Once all executions have been completed (one for each character), the mode number of instructions retired is computed across all executions.
 The mode is defined as the value that appears most often in the set of recordings.
-This is done in line \autoref{findinputcode_mode} of \autoref{findinputcode}.
+This is done in line \ref{mode} of algorithm \ref{findinputcode}.
 \item Next is the filtering stage which is responsible for identifying characters likely to be part of a valid input string at the current index.
-Under the assumption that most characters are not valid for the current index, \tool{} gathers all characters with the number of instructions executed outside the range of the mode +/- an epsilon value, as shown in line \autoref{findinputcode_filter} of \autoref{findinputcode}.
+Under the assumption that most characters are not valid for the current index, \tool{} gathers all characters with the number of instructions retired outside the range of the mode +/- an epsilon value, as shown in lines \ref{filterbegin}-\ref{filterend} of algorithm \ref{findinputcode}.
 These characters represent valid characters for the current index of the input string.
 The epsilon value accounts for small variations in the hardware counters.
-While the initial reaction is to expect a valid character to result in more instructions executed than a non-valid character, this is not always the case.
+While the initial reaction is to expect a valid character to result in more instructions retired than a non-valid character, this is not always the case.
 As discussed earlier, the input validation mechanism can be complex and can vary in behavior.
 The validation mechanism may verify input against all accepted input strings before denying the provided input.
-Such a scenario results in more instructions executed for invalid input than valid input.
-\item For each valid character \textit{c}, \tool{} forks, with every child process repeating steps 1-4, starting with $ c + s $ as the initial input string, where $+$ signifies concatenation (starting at line \autoref{findinputcode_fork} of \autoref{findinputcode}).
+Such a scenario results in more instructions retired for invalid input than valid input.
+\item For each valid character \textit{c}, \tool{} forks, with every child process repeating steps 1-4, starting with $ c + s $ as the initial input string, where $+$ signifies concatenation (starting at line \ref{fork} of algorithm \ref{findinputcode}).
 If no valid characters are found, the end of a valid input string has been reached and the string is recorded.
@@
 \includegraphics[height=2in,width=6.5in]{string_builder.png}
-\label{fig:string_builder}
-\caption{This diagram visualizes the input string being constructed throughout five consecutive time stages of the execution of \tool{}.
+\caption{This diagram visualizes the input string being constructed throughout five consecutive time stages of the execution of 
+\tool{}.
 Multiple outgoing arrows represent a process forking.
 This occurs when multiple valid characters are detected for a given index of the input string and, as a result, the finding input process forks to concurrently handle all of the discovered characters}
+\label{fig:string_builder}
 \end{figure*}
@@
 As seen in the figure, at time \textit{T0}, \tool{} begins with an empty string.
-At time \textit{T1}, \tool{} finds valid characters for index 0 of the input and is forked into \textit{n} child processes where \textit{n} is the number of valid characters found.
+At time \textit{T1}, \tool{} finds valid characters for index 0 of the input string and is forked into \textit{n} child processes where \textit{n} is the number of valid characters found.
 This continues until all child processes are done.
+% TODO: snippets of code of input validation code from our experiments (also for finding input size)
 
@@
 The same side channels can be used to detect the expected size of the program's input or even the size of the program's input buffer.
-Specifically, there is a change in the behavior of the program with respect to the number of instructions executed when the size of the input is larger than expected.
+Specifically, there is a change in the behavior of the program with respect to the number of instructions retired when the size of the input is larger than expected.
 Most programs are designed to read a predefined number of bytes from the source of input and store the read bytes in a buffer.
 A popular way of doing that is by using the \textit{fgets()} or \textit{scanf()} functions of the GNU C stdio library.
-As the input size grows, the number of instructions executed also increases linearly.
-This happens because additional instructions are executed to read more characters and store them into the buffer.
-However, when the input string is larger than the expected size the number of instructions executed no longer increases as a function of the input size.
+As the input size grows, the number of instructions retired also increases linearly.
+This happens because additional instructions are retired to read more characters and store them into the buffer.
+However, when the input string is larger than the expected size the number of instructions retired no longer increases as a function of the input size.
 This behavior can be observed and exploited to find the expected input size, which may often also be the size of the input buffer.
@@
 The character selected is used to build invalid input strings of various sizes.
-It is crucial to use invalid input to avoid noisy fluctuations in the number of instructions executed due to valid input processing.
-Initially, the program is executed with short length input strings (i.e. two and three characters).
-The number of user-land instructions executed during these two executions are recorded.
+It is crucial to use invalid input to avoid noisy fluctuations in the number of instructions retired due to valid input processing.
+Initially, the program is retired with short length input strings (i.e. two and three characters).
+The number of user-land instructions retired during these two executions are recorded.
 These two values may be equal or differ, depending on the program's logic.
 The two values may be equal if, for example, the program uses the system call \textit{read()} directly to read data from the input file descriptor.
-Since we are measuring the number of user-land instructions retired (not also instructions executed by the kernel), the instructions executed by the \textit{read} functions are not included in our counter so there will be no linear increase in the number of instructions executed as the input size is increased.
+Since we are measuring the number of user-land instructions retired (not also instructions retired by the kernel), the instructions retired by the \textit{read} functions are not included in our counter so there will be no linear increase in the number of instructions retired as the input size is increased.
 If the two values are equal, the program is executed with increasingly large input until a change in the number of executions is recorded due to the program logic.
 There may not always be a change (depending on the program logic when the input size exceeds the expected size), however, we note that this scenario is not very interesting since the expected size of the input may also be obtained by executing the program with \textit{strace}.
-If the values differ (as discussed earlier), the difference in the number of instructions executed is recorded and the program is executed, increasing the size of the input, until a change is detected.
+If the values differ (as discussed earlier), the difference in the number of instructions retired is recorded and the program is executed, increasing the size of the input, until a change is detected.
 Instead of increasing the input size linearly, the input size is increased exponentially until a change is detected.
 Once the change is detected, the input size is decreased to find the exact size at which the change happens.
-It is also possible that the number of instructions executed oscillates between two (or more values).
+It is also possible that the number of instructions retired oscillates between two (or more values).
 One example is a palindrome finder.
@@
 \subsection{Generating Protocol State Machine}
-% TODO
-% add diagram depicting process
 Thus far, we have described how \tool{} discovers valid input for a given binary autonomously.
@@
 
-\begin{lstlisting}[caption={Pseudocode implementing the protocol state machine generator},label={lst:psmcode}]
-void main(inputs) {
-  for len in len(inputs):
-    permutations = get_permutation(len)
-    results = []
-    for p in permutations:
-      exec(p, 0, results)
-    compare_inputs(results, inputs)
-}
+\begin{algorithm}
+\caption{Pseudocode implementing the protocol state machine generator}
+\label{psmcode}
+\begin{algorithmic}[1]
+\Procedure{main}{$inputs$}
+\For{i = 1, \Call{size}{$inputs$}}
+\State length = i
+\State permutations = \Call{GetPermutation}{$length$}
+\State results = $\emptyset$
+\For{j = 0, \Call{size}{$permutations$}}
+\State permutation = permutations[$j$]
+\State \Call{Exec}{$permutation, 0, results$}
+\EndFor
+\State \Call{Compare}{$results, inputs$}
+\EndFor
+\EndProcedure
 
-void exec(cmds, i, results) {  
-  i = 0
+\Procedure{Exec}{$cmds, i, results$}
+\If{ i == \Call{size}{$cmds$}}
+\State results += cmds
+\State return
+\EndIf
 
-  # base case
-  if i == len(cmds):
-    results.append(cmds)
-    return 0
+\State initCmds = cmds[0:i]
+\State cmd = cmds[i]
+\State args = \Call{FindArgs}{$initCmds, cmd, i$}
+%\Comment{find arguments for cmd, given initial commands initCmds}
 
-  #finds args for input string at index i
-  initCmds = cmds[0:i]
-  cmd = cmds[i]
-  args = find_args(initCmds, cmd, i)
+\For{j=0, \Call{size}{$args$}}
+\State cmds[$j$] = cmds[$j$] + arg
+\State \Call{Exec}{$cmds, i+1, results$}
+\EndFor
+\EndProcedure
 
-  for arg in args:
-    cmds[i] = cmds[i] + arg  #concatenation
-    exec(cmds, i+1, results)
-}
-
-void compare_inputs(results, inputs) {
-  #organize results
-  for cmds in results:
-    lastCmd = cmds[len(cmds)]
-
-    trace = get_pin_trace(cmds)
-
-    traceForLastCmd = split_trace(cmds)
-    traceVector = dfs_cfg(trace)
-
-    traces[lastCmd][cmds] = traceVector
-
-  #compare traces for same input/cmd
-  # from different permutations
-  for cmd in cmds:
-    for cmdTraces in traces[cmd]:
-      clusters = compare(cmdTraces)
-
-   #find outliers in clusters
-   filter(clusters)
-}
-
-\end{lstlisting}
+\Procedure{Compare}{$results, inputs$}
+\For{i=0, \Call{size}{$results$}}
+\State cmds = results[$i$]
+\State lastCmd = cmds[\Call{size}{$cmds$}-1]
+\State trace = \Call{ExecBinWithPin}{$cmds$}
+\State traceForLastCmd = \Call{SplitTrace}{$trace, cmds$}
+\State allTraces[$lastCmd$][$cmds$] = traceForLastCmd
+\EndFor
+\For{i=0, \Call{size}{$inputs$}}
+\State cmd = inputs[$i$]
+\For{j=0, \Call{size}{$allTraces[cmd]$}}
+\State cmdTraces = allTraces[$cmd$][$j$]
+\State clusters = \Call{Compare}{$cmdTraces$}
+\State goodPermutation = \Call{Filter}{$clusters$}
+\EndFor
+\EndFor
+\EndProcedure
+\end{algorithmic}
+\end{algorithm}
 
@@
 The architecture of this component is depicted in Figure \ref{fig:psm_arch}.
-The process is describe below and shown as pseudocode in Listing \ref{lst:psmcode}.
-First, all possible permutations of all sizes from 1 to the number of input strings discovered are computed.
+The process is described below and shown as pseudocode in algorithm \ref{psmcode}.
+First, all possible permutations of all sizes from 2 to the number of input strings discovered are computed (e.g.
+given input strings $a,b,c,d$, the permutations generated include [a,b] , [a,c] , [a,d] , [b,a] , [b,c] , [b,d] ...
+[a,b,c] , [a,b,d], [a,c,b] ...
+[a,b,c,d] , ...
+etc.).
 Every permutation is passed through the dynamic analysis engine.
@@
 Maintaining the original order of the inputs, the engine tests each string for the presence of expected arguments.
-For example, if the input string is \textit{SET} and the program accepts input of the format \textit{SET data}, the analysis tries to determine that `data' is a valid argument.
+For example, if the input string is \textit{SET} and the program accepts input of the format \textit{SET data}, the analysis tries to determine that `data' is a valid argument by applying the input finding method with \textit{SET} as the command and identifying additional characters after the $[SPACE]$ character.
 The arguments found are saved for later use.
@@
 Given a permutation $p$ of size $s$ let $\alpha(p_s)$ represent the execution trace of the command at index $s$ of the permutation (i.e. last command of the permutation).
-It is natural to assume that given the correct (or expected) order the of input strings $1$ through $(i-1)$ of the permutation, the input $p_s$ experiences a different execution trace (or control flow path) compared to an incorrect permutation (assuming $p_s$ is at the same index in the permutation).
-This information is used to determine, for every input string discovered, if there exist a permutations which affects its control flow path.
+Take for example two different permutations $p1$ and $p2$.
+Both permutations are of the same length, $i$.
+Both permutations are composed of the same input strings and have the same last input string (e.g. [a,b,c,d] vs. [b,a,c,d]).
+Assume permutation $p1$ represents the correct (or expected) order of the input strings.
+It is natural to assume that the input string $p_i$ (i.e.
+last input string in the permutation) experiences a different execution trace (or control flow path) as a result of permutation $p1$ than permutation $p2$.
+This information is used to determine, for every input string discovered, if there exist a permutation which affects its control flow path.
 If such a permutation is discovered, it is added to the protocol state machine.
@@
 This paper introduces two novel techniques for autonomously generating valid input strings for closed x86 binaries and for generating protocol state machine using the discovered input.
-For generating valid input strings, our method exploits the number of instructions executed to infer if a given character is valid for a given index of an input string.
+For generating valid input strings, our method exploits the number of instructions retired to infer if a given character is valid for a given index of an input string.
 This process allows us to recursively and incrementally craft valid inputs.

### 2015-05-17 paper/main.tex
final draft of the paper

@@
 \usepackage{algorithm}
+\usepackage{url}
+\usepackage{listings}
+\usepackage{color}
 
@@
 \title{InputFinder: Reverse Engineering Closed Binaries via Side Channels}
-\numberofauthors{2}
-\author{
-\alignauthor
-Bogdan Copos\\
-	\affaddr{University of California, Davis}\\
-	\email{bcopos@ucdavis.edu}
-\alignauthor
-Praveen Murthy\\
-	\affaddr{Fujitsu Laboratories of America}\\
-	\email{praveen.murthy@us.fujitsu.com}
-}
+%\numberofauthors{2}
+%\author{
+%\alignauthor
+%Bogdan Copos\\
+%	\affaddr{University of California, Davis}\\
+%	\email{bcopos@ucdavis.edu}
+%\alignauthor
+%Praveen Murthy\\
+%	\affaddr{Fujitsu Laboratories of America}\\
+%	\email{praveen.murthy@us.fujitsu.com}
+%}
 \maketitle
@@
 To assess its effectiveness, we have developed \tool{} and tested it against binaries from the DARPA Cyber Grand Challenge example set.
-Our evaluations show that ...
+Our evaluations show that our method is not only effective in finding input and determining whether a protocol is expected but can also find unexpected control flow paths.
 \end{abstract}
@@
 %motivation
-Without complete test suites, the program cannot be properly tested and bugs may be omitted during testing.
+Without complete test suites, the program cannot be properly tested and some bugs may be omitted during testing.
 While valid input can be learned from software documentation or source code, when such information is not available or complete, other methods must be used.
@@
 Fuzzing done randomly with no inference attempted of the program's behavior cannot be effective as the input space is usually too large.
-Symbolic execution has traditionally been used with manual intervention to identify potential symbolic variables to begin exploration with, a luxury we do not have when analyzing closed binaries autonomously, in addition to scalability issues and limitations of decision procedures that can be used to solve the constraints.
+Symbolic execution has traditionally been used with manual intervention to identify potential symbolic variables to begin exploration with, a luxury we do not have when analyzing closed binaries autonomously, in addition to scalability issues and limitations of decision procedures that are used to solve the constraints.
 
@@
 The first component exploits side channels to build valid input for closed binaries.
-Specifically, as the program is given various input, \tool{} records the number of instructions retired during the program's execution for each input and uses a statistical measure to infer the program's expected input.
+Specifically, as the program is given various inputs, \tool{} records the number of instructions \textbf{retired} during the program's execution for each input and uses a statistical measure to infer the program's expected input.
 The number of instructions retired differs from the number of instructions executed by a processing unit.
@@
 Our method relies on observing changes in the number of instructions retired by programs during their input validation process to learn valid input strings.
-	\item We describe how our method can be used to not only find valid input strings, but also to determine the expect input size and classify input based on type (e.g.
+	\item We describe how our method can be used not only to find valid input strings, but also to determine the expect input size and classify input based on type (e.g.
 alphabetic, alphanumeric, numeric, etc.).
@@
 The paper is organized as follows.
+Section \ref{prevwork} discusses related works and highlights differences between the related works and the work presented here.
 Section \ref{background} covers background information relating to concepts utilized in this work.
-Section \ref{prevwork} discusses related works and highlights differences between the related works and the work presented here.
 The details of our approach are explained in Section \ref{methodology}.
@@
 Then we introduce our method for building protocol state machines using the discovered input.
-Section \ref{results} describes evaluation of the work presented and the results.
+Section \ref{results} describes the evaluation of the work presented and the results.
 In Section \ref{futurework}, we share future plans for our work and explain cases which the proposed method does not handle.
 
+\section{Previous Work} \label{prevwork}
+Dynamic analysis is extensively studied in academia and is used widely in the industry by security specialists.
+Two of the main approaches to vulnerability testing are fuzzing and symbolic execution.
+
+Fuzzers are a popular choice amongst software testers.
+They are easy to use, easy to implement, and can produce surprising results \cite{millerfuzz, millerfuzzrevisited}.
+There are two classes of fuzzers: black box fuzzers and white box fuzzers.
+In black box fuzzing, the program is watched as it is tested with randomly generated input.
+If the program crashes, the input(s) are recorded and reported.
+No prior program specifications or knowledge of the input format is available in such cases.
+Since black box fuzzing is completely random, its effectiveness is limited, especially with respect to coverage.
+
+Symbolic execution is another method used for generating valid input.
+When a program analyzer symbolically executes a program, it does not use concrete values.
+Instead, the input is represented as a mathematical formula composed of symbols and equations resulting from the constraints (related to the input) gathered during the program's execution.
+Specifically, the tester first marks the input as symbolic.
+This can also be done automatically via taint analysis.
+The program is then executed by an interpreter that assumes symbolic values for those symbolic input locations.
+As the program executes, whenever operations are performed on those symbolic variables, a formula in terms of those symbols keeps track of the mathematical operations performed on the symbolic variables.
+The formula can also have constraints on those symbols based on outcomes of conditional branches encountered during the executions.
+At the end of execution, or during certain portions of the execution, the formula is examined by a constraint solver to determine whether it is satisfiable (meaning that a solution exists to make the formula true), and if it does, concrete values for all symbolic variables in the formula are computed.
+Those concrete values will then constitute a set of test cases that can execute the program along the set of paths (because of branches) encountered in the formula.
+Symbolic execution has been proven effective by previous works \cite{symbolicexecution, exesymex}.
+One of the most important works in this area is described in \cite{exesymex} which introduces EXE, a dynamic analysis tool capable of generating inputs that crash programs by running the programs with symbolic input.
+
+White box fuzzing leverages the symbolic execution traces of a given binary to produce more inputs.
+One of the many previous works is \cite{godefroid2008automated} which introduces SAGE, an automated white box fuzzer.
+White box fuzzing differs from black box fuzzing in that it uses the concrete inputs to gather constraints and then systematically negates constraint clauses to generate additional inputs that cover more branches in the program.
+White box fuzzing can attain better code coverage but can be slow.
+Symbolic execution (and consequently white box fuzzing) is crippled by innate limitations of constraint solvers.
+While symbolic execution engines perform well with linear constraints, non-linear constraints greatly impact efficiency.
+One of the works utilizing white box fuzzing is \cite{molnar2009dynamic} which introduces SmartFuzz, a dynamic test generation tool.
+SmartFuzz uses symbolic execution to learn about the program's input and passes that information to a fuzzer for test case generation.
+
+The main difference between our work and SmartFuzz is that our work does not use symbolic execution to gain insight into the format of the input.
+Similar to SmartFuzz, the output of our work can be used by a fuzzer to generate test cases.
+Unlike the EXE symbolic execution engine, the SAGE tool, and SmartFuzz, the approach presented in our work is not hindered by constraint solvers.
+Additionally, compared to EXE, our method does not require any instrumentation or source code.
+\tool{} does not execute the program with symbolic input but rather it bootstraps itself by building up a set of valid inputs (from an initially empty set) by gathering information about the execution of a program.
+
+Another related work is presented in \cite{smartfuzzer}, where the authors use a combination of static and dynamic analysis to build a smart fuzzer.
+Their work uses static analysis to collect preliminary information such as an approximation of the inter procedural control flow graph.
+Having gathered static information, their dynamic analysis engine executes the program while monitoring dependencies between the input and the dynamic control flow path experienced by the program.
+Their work also takes advantage of a constraint solver that, given a path of the program, generates input which forces the program to follow the chosen path.
+Our work \tool{} differs from their approach in several ways.
+One difference is that our work uses neither static analysis nor constraint solvers.
+Instead our input finding method relies only on information extracted from the execution of the un-instrumented program.
+
+There have also been numerous works studying and applying dynamic similarity.
+In \cite{blanketexec}, the authors propose blanket execution, a novel dynamic equivalence testing technique.
+Blanket execution is designed to collect information about the execution of functions, including side effects, which is then used for similarity testing.
+Another important work is \cite{differential} which introduces \textit{differential slicing} for the purpose of identifying flow and value differences across multiple executions of the same program for security applications.
+Both our approach and the previous approaches mentioned above use a dynamic approach to gather information about the program.
+However, our approach aims at finding differences in executions of a given program in order to make inferences about the underlying protocol expected by the program.
+This allows us to discover the orders of inputs which experience unique control flow paths.
+
+Another area of research relevant to this paper is protocol reverse engineering.
+An important result in this area is \cite{prospex} which introduces a technique for extracting protocol specifications from packet captures and automatically generating input which can be used by a stateful fuzzer.
+There are two notable differences between Prospex and our work.
+While the end result is similar, \tool{} assumes no prior knowledge about the tested program and hence it does not utilize packet captures for protocol specification extraction.
+Instead, \tool{} infers the protocol through dynamic analysis by automatically finding input and using that to generate a protocol state machine.
+Caballero et.
+al introduce Polyglot \cite{polyglot}, a tool that uses dynamic binary analysis to observe how programs process the received application data and uses that information to infer the protocol's message format.
+Their work is similar because of the use of dynamic binary analysis and the use of information from program traces to make inferences about the protocol format.
+However, our work does not try to determine the protocol message format but rather the high level protocol.
+Additionally, unlike Polyglot, our work does not use test cases or previously known input.
+To date, we do not know of any protocol reverse engineering techniques which work without any prior knowledge about the program's input.
+
 \section{Background} \label{background}
+
 \subsection{Side Channels} \label{bg_sidechannels}
@@
 
-\section{Previous Work} \label{prevwork}
-Dynamic analysis is extensively studied in academia and is used widely in the industry by security specialists.
-Two of the main approaches to vulnerability testing are fuzzing and symbolic execution.
-
-Fuzzers are a popular choice amongst software testers.
-They are easy to use, easy to implement, and can produce surprising results \cite{millerfuzz, millerfuzzrevisited}.
-There are two classes of fuzzers: black box fuzzers and white box fuzzers.
-In black box fuzzing, the program is watched as it is tested with randomly generated input.
-If the program crashes, the input(s) are recorded and reported.
-No prior program specifications or knowledge of the input format is available in such cases.
-Since black box fuzzing is completely random, its effectiveness is limited, especially with respect to coverage.
-
-Symbolic execution is another method used for generating valid input.
-When a program analyzer symbolically executes a program, it does not use concrete values.
-Instead, the input is represented as a mathematical formula composed of symbols and equations resulting from the constraints (related to the input) gathered during the program's execution.
-Specifically, the tester first marks some memory locations as symbolic.
-This can also be done automatically by, for example, marking the buffer used in \textit{read()} system call as symbolic.
-However, this does not always produce good results, especially if the program reads one character at a time and then copies the read character into a second buffer.
-The program is then executed and the tester enters some input.
-Regardless of the value entered, the program replaces the input with a symbol.
-As the input flows through the program and encounters conditional statements, the symbolic value of the input changes to reflect the faced constraints.
-At the end of the execution, a constraint solver takes the formulat representing the symbolic value of the input and attempts to solve it in order to obtain valid program input.
-Symbolic execution has been proven effective by previous works \cite{symbolicexecution, exesymex}.
-Similar to white box fuzzing, symbolic execution has a fundamental limitation with respect to constraint solving.
-While symbolic execution engines perform well with linear constraints, non-linear constraints greatly impact efficiency.
-One of the most important works in this area is described in \cite{exesymex} which introduces EXE, a dynamic analysis tool capable of generating inputs that crash programs by running the programs with symbolic input.
-Unlike the EXE symbolic execution engine, the SAGE tool and SmartFuzz, the approach presented in our work is not hindered by constraint solvers.
-Additionally, compared to EXE, our method does not require any instrumentation or source code.
-\tool{} does not execute the program with symbolic input but rather it bootstraps itself by building up a set of valid inputs (from an initially empty set) by gathering information about the execution of a program.
-
-White box fuzzing leverages the symbolic execution traces of a given binary to produce more inputs.
-One of the many previous works is \cite{godefroid2008automated} which introduces SAGE, an automated white box fuzzer.
-White box fuzzing differs from black box fuzzing in that it uses the program's feedback to make inferences about valid input and uses constraint solvers to generate more input.
-White box fuzzing can attain better code coverage but can be slow and is crippled by innate limitations of constraint solvers.
-The only similarity between the approach introduced in this paper and white box fuzzing is that both approaches use information from the program's execution to craft valid input.
-However, our method does not use symbolic execution.
-
-One of the works utilizing white box fuzzing is \cite{molnar2009dynamic} which introduces SmartFuzz, a dynamic test generation tool.
-SmartFuzz uses symbolic execution to learn about the program's input and passes that information to a fuzzer for test case generation.
-The main difference between our work and SmartFuzz is that our work does not use symbolic execution to gain insight into the format of the input.
-Similar to SmartFuzz, the output of our work can be used by a fuzzer to generate test cases.
-
-Another related work is presented in \cite{smartfuzzer}, where the authors use a combination of static and dynamic analysis to build a smart fuzzer.
-Their work uses static analysis to collect preliminary information such as an approximation of the inter procedural control flow graph.
-Having gathered static information, their dynamic analysis engine executes the program while monitoring dependencies between the input and the dynamic control flow path experienced by the program.
-Their work also takes advantage of a constraint solver that, given a path of the program, generates input which forces the program to follow the chosen path.
-Our work \tool{} differs from \cite{smartfuzzer} in several ways.
-One difference is that our work uses neither static analysis nor constraint solvers.
-Instead we rely only on information extracted from the execution of the uninstrumented program.
-
-There have also been numerous works studying and applying dynamic similarity.
-In \cite{blanketexec}, the authors propose blanket execution, a novel dynamic equivalence testing technique.
-Blanket execution is designed to collect information about the execution of functions, including side effects, which is then used for similarity testing.
-Another important work is \cite{differential} which introduces \textit{differential slicing} for the purpose of identifying flow and value differences across multiple executions of the same program for security applications.
-Both our approach and the previous approaches mentioned above use a dynamic approach to gather information about the program.
-However, our approach aims at finding differences in executions of a given program in order to make inferences about the underlying protocol expected by the program.
-This allows us to discover differences in inputs which experience different control flow paths.
-
-Another area of research relevant to this paper is protocol reverse engineering.
-An important result in this area is \cite{prospex} which introduces a technique for extracting protocol specifications from packet captures and automatically generating input which can be used by a stateful fuzzer.
-There are two notable difference between Prospex and our work.
-While the end result is similar, \tool{} assumes no prior knowledge about the tested program and hence it does not utilize packet captures for protocol specification extraction.
-Instead, \tool{} infers the protocol through dynamic analysis by automatically finding input and using that to generate a protocol state machine.
-To date, we do not know of any protocol reverse engineering techniques which work without any prior knowledge about the program's input.
-
 \section{Methodology} \label{methodology}
@@
 The first module is responsible for discovering input and properties of the expected input, including input size and the type of the input.
-The input is then be used by the second component to generate a protocol state machine.
+The input generated by the first module is then used by the second module to generate a protocol state machine.
 The protocol state machine can be used by third party tools such as stateful fuzzers for even better code coverage.
 \subsection{Finding Valid Input}
-The majority of software programs accept input, perform transformations on the input and output results.
+The majority of software programs accept input, perform transformations on the input, and output results.
 However, most programs do not accept completely random input.
@@
 Although validation mechanisms can be very sophisticated, they are often a combination of string comparisons and conditional statements.
+Code snippets of input validation from our tested binaries can be found in Appendix A.
 Our method exploits changes in the number of user-land instructions retired during multiple executions with varying inputs to make inferences about the program's validation mechanism.
@@
 \centering
-\includegraphics[height=2.5in,width=6.5in]{architecture.png}
-\caption{This figure depicts the architecture of the finding input process. There are three main parts: the execution engine, the valid character filter and the input string builder.}
+\includegraphics[width=\textwidth]{architecture.pdf}
+\caption{This figure depicts the architecture of the finding input process.
+There are three main parts: the input string builder, the performance counter (i.e.
+\textit{perf} utility), and the valid character filter}
 \label{fig:architecture}
@@
 Note that \tool{} does not record the total number of instructions but rather user-land instructions.
-The reason for this, as explained in Section \ref{bg_hardwarereg}, is that the total number of instructions retired can vary greatly from execution to execution due to operating system factors (e.g.
+The reason for this, as explained in Section \ref{bg_hardwarereg}, is in contrast to user-land instructions, that the total number of instructions retired can vary greatly from execution to execution for various reasons (e.g.
 branch prediction, cache).
 The Expect tool is used for automating the program interaction.
-This step is shown in line \ref{execute} of algorithm \ref{findinputcode}
+This step is shown in line \ref{execute} of algorithm \ref{findinputcode}.
 \item Once all executions have been completed (one for each character), the mode number of instructions retired is computed across all executions.
@@
 \centering
-\includegraphics[height=2in,width=6.5in]{string_builder.png}
-\caption{This diagram visualizes the input string being constructed throughout five consecutive time stages of the execution of 
+\includegraphics[width=\textwidth]{string_builder.pdf}
+\caption{This diagram visualizes the input strings being constructed throughout five consecutive time stages of the execution of 
 \tool{}.
 Multiple outgoing arrows represent a process forking.
-This occurs when multiple valid characters are detected for a given index of the input string and, as a result, the finding input process forks to concurrently handle all of the discovered characters}
+This occurs when multiple valid characters are detected for a given index of the input string and, as a result, the finding input process forks to concurrently handle all of the discovered characters.}
 \label{fig:string_builder}
@@
 This continues until all child processes are done.
-% TODO: snippets of code of input validation code from our experiments (also for finding input size)
 
@@
 As the input size grows, the number of instructions retired also increases linearly.
-This happens because additional instructions are retired to read more characters and store them into the buffer.
-However, when the input string is larger than the expected size the number of instructions retired no longer increases as a function of the input size.
+This happens because additional instructions are needed to read more characters from the input source.
+However, when the input string is larger than the expected size, the number of instructions retired no longer increases as a function of the input size.
 This behavior can be observed and exploited to find the expected input size, which may often also be the size of the input buffer.
@@
 It is crucial to use invalid input to avoid noisy fluctuations in the number of instructions retired due to valid input processing.
-Initially, the program is retired with short length input strings (i.e. two and three characters).
+Initially, the program is executed twice: once with an input string of two invalid characters and once with an input string of three invalid characters.
 The number of user-land instructions retired during these two executions are recorded.
-These two values may be equal or differ, depending on the program's logic.
+These two values may be equal or different, depending on the program's logic.
 The two values may be equal if, for example, the program uses the system call \textit{read()} directly to read data from the input file descriptor.
-Since we are measuring the number of user-land instructions retired (not also instructions retired by the kernel), the instructions retired by the \textit{read} functions are not included in our counter so there will be no linear increase in the number of instructions retired as the input size is increased.
+Since we are measuring the number of user-land instructions retired (not also instructions retired by the kernel), the instructions retired by the \textit{read()} function are not included in our counter so there will be no linear increase in the number of instructions retired as the input size is increased.
 If the two values are equal, the program is executed with increasingly large input until a change in the number of executions is recorded due to the program logic.
-There may not always be a change (depending on the program logic when the input size exceeds the expected size), however, we note that this scenario is not very interesting since the expected size of the input may also be obtained by executing the program with \textit{strace}.
 If the values differ (as discussed earlier), the difference in the number of instructions retired is recorded and the program is executed, increasing the size of the input, until a change is detected.
@@
 The same underlying concept can be applied to such cases.
+
+Determining the exact size is useful when performing symbolic execution on a buffer where one cannot make the size of the buffer symbolic (since frameworks such as \textit{Fuzzball} do not allow for that).
+In such cases, being able to determine the buffer size is necessary and useful to instruct the symbolic execution engine the exact size of the buffer that should be made symbolic.
+We detected 5 crash inputs in the tested binaries because of our finding input size approach.
 
@@
 Without first selecting a database and perhaps even a table, some commands such as \textit{INSERT} or \textit{SELECT} are useless and have no effect on any data.
-This is where the second component of \tool{} steps in.
-Specifically, this component identifies expected order (or permutation) of the discovered input strings.
+We develop the second component of \tool{} to address this issue.
+Specifically, this component identifies expected orderings (or permutations) of the discovered input strings, and constructs a protocol state machine.
 
@@
 \centering
-\includegraphics[width=0.5\textwidth]{psmarch}
+\includegraphics[width=0.53\textwidth]{psmarch.pdf}
 \label{fig:psm_arch}
@@
 
-Once input has been discovered, \tool{} constructs a protocol state machine.
 The architecture of this component is depicted in Figure \ref{fig:psm_arch}.
@@
 The dynamic analysis execution engine first splits the permutation into individual input strings.
-Maintaining the original order of the inputs, the engine tests each string for the presence of expected arguments.
-For example, if the input string is \textit{SET} and the program accepts input of the format \textit{SET data}, the analysis tries to determine that `data' is a valid argument by applying the input finding method with \textit{SET} as the command and identifying additional characters after the $[SPACE]$ character.
+Maintaining the original order of the inputs within the permutation, each string is tested for the presence of expected arguments.
+For example, if the input string is \textit{SET} and the program accepts input of the format \textit{SET data}, the analysis tries to determine that \textit{data} is a valid argument by applying the input finding method with \textit{SET} as the command and identifying additional characters after the $<SPACE>$ character.
 The arguments found are saved for later use.
-Once every input string of a given permutation has been tested for arguments, the inputs are passed through a instrumented binary for the collection of execution trace.
-The binary is instrumented with a \textit{Pin} tool, developed by us, which logs the instructions executed as a function of time.
+Once every input string of a given permutation has been tested for arguments, the inputs are passed through the instrumented version of the program binary for the collection of the execution trace.
+The binary is instrumented with a \textit{Pin} tool, developed by us, which logs the starting address of every basic block executed, as a function of time.
+A basic block is a set of instructions with a single entry and a single exit point.
 Once all permutations of a given size have been executed, the dynamic similarity test begins.
 This test splits each execution trace into parts, each part being associated with an input string.
+The splitting process is done by first executing the program without any input, examining the trace, and extracting a short sequence of basic blocks from the end of the trace.
+The sequence of basic blocks extracted is used as the delimiter during the splitting process and it represents the basic blocks executed by the program right before reading and processing the user input.
 We refer to the execution trace of the input string \textit{SET} as $\alpha(SET)$.
-For all permutations of the same size with the same last input string, the $\alpha(lastCmd)$ are clustered.
+For all permutations of the same size with the same last input string, the traces $\alpha(lastCmd)$ are clustered.
 Given a permutation $p$ of size $s$ let $\alpha(p_s)$ represent the execution trace of the command at index $s$ of the permutation (i.e. last command of the permutation).
@@
 last input string in the permutation) experiences a different execution trace (or control flow path) as a result of permutation $p1$ than permutation $p2$.
-This information is used to determine, for every input string discovered, if there exist a permutation which affects its control flow path.
+This information is used to determine, for every input string discovered, if there exists a permutation which affects its control flow path.
 If such a permutation is discovered, it is added to the protocol state machine.
@@
 \centering
-\includegraphics[width=0.5\textwidth]{protocoldiagram}
+\includegraphics[width=0.53\textwidth]{protocoldiagram.pdf}
 \label{fig:psm_example}
@@
 Since there is no information about the states and their meaning, the states are represented by circles with generic names (e.g. $S0$).
-The arrows represent input strings which allow for transition between states.
+The arrows represent input strings that enable the transition between states.
 
@@
 \subsection{Finding Input, Input Size and Classifying Input}
-\tool{} was tested on \numbinaries{} x86 binaries to measure its effectiveness from the DARPA Cyber Grand Challenge Example set \cite{darpacgc}.
-The results are summarized in Figure \ref{fig:findinputresults}.
-Analyzing the source code showed that out of the total \numbinaries{} binaries, 21 had predefined valid input strings.
-\tool{} found input strings for 14 of the 21 binaries and categorize the input correctly for 100\% of the 14 binaries.
-%TODO
-Our method was able to determine the input length for XXX of the binaries.
+In order to determine its effectiveness, \tool{} was tested on \numbinaries{} x86 binaries from the DARPA Cyber Grand Challenge Example set \cite{darpacgc}.
+The Cyber Grand Challenge (CGC) binaries are 32 bit x86 binaries in the CGC format.
+The only difference between the CGC format and the well known ELF format is in the header of the binaries and in the fact that dynamic linking is not supported.
+Additionally, they are designed to only be executed in the Cyber Grand Challenge platform and utilize different system calls than the UNIX system calls.
+The Cyber Grand Challenge organizers also provide the source code of the binaries which, with some effort, can be modified to use UNIX system calls and be built as ELF x86 binaries.
 
-\begin{figure}[h!]
-\caption{The results of our input finding method. `N/A' signifies that the binary did not have expect any predefined input strings but on the contrary accepted any input.}
+The results are summarized in Table \ref{binaryresults}.
+Analyzing the source code showed that out of the \numbinaries{} binaries, 21 had predefined valid input strings.
+\tool{} found input strings for 13 of the 21 binaries and categorized the input correctly for 100\% of the 13 binaries.
+The running time varied between a few minutes and an hour, depending on the number and length of the valid input strings discovered.
+Our method was able to determine the input length for 19 of the 24 binaries.
+Note that the our approach was able to find the expected input length even if it was not able to find valid input.
+Additionally, our method was able to determine crash inputs for 5 of the binaries.
+In all cases, the crash input caused a buffer overflow and caused the binary to exit unexpectedly.
+
+\begin{table*}
 \centering
-\includegraphics[width=0.5\textwidth]{findinputresults}
-\label{fig:findinputresults}
-\end{figure}
-%To further show the advantage of this method, we built a smart fuzzer which uses our method to determine input and protocol for a given binary.
-%We use the smart fuzzer and compare it to black box fuzzers symbolic execution engines with respect to code coverage over units of time.
-%Using such a comparison allows the reader to determine which method is more efficient while also performing well.
-
-% 7f... tic tac toe, asks for player names first so without skipping those steps it cannot find any valid input
+\begin{tabular}{|c|c|c|c|c|c|} \hline
+\textbf{Binary} & \textbf{Input} & \textbf{Number of Inputs} & \textbf{Input Size} & \textbf{Crash Input} & \textbf{Protocol State Machine}\\ \hline
+06459301 & yes (8 min) & 8 out of 8 & yes & no & -\\ \hline
+06b71301 & no & - & yes & no & -\\ \hline
+07a9a901 & - & - & yes & yes & - \\ \hline
+0b32aa01 & yes (2 min) & 1 out of 1 & yes & yes & - \\ \hline
+11dc8e01 & no & - & yes & no & - \\ \hline
+1877a601 & yes (11 min) & 6 out of 6 & yes & no & - \\ \hline
+250d1101 & yes (20 min) & 5 out of 5 & yes & no & yes \\ \hline
+2eca0101 & yes (42 min) & 7 out of 7 & no & no & - \\ \hline
+37e97201 & yes (38 min) & 6 out of 6 & yes & no & - \\ \hline
+3dcf1a01 & yes (6 min) & 6 out of 6 & yes & yes & - \\ \hline
+48b9cf01 & yes (50 min) & 9 out of 9 & yes & no & - \\ \hline
+65884701 & no & - & no & no & - \\ \hline
+701b7301 & - & - & yes & no & - \\ \hline
+7262d006 & yes (5 min) & 4 out of 5 & yes & no & - \\ \hline
+7fa39f01 & no & - & yes & yes & - \\ \hline
+b8993403 & yes (12 min) & 1 out of 1 & yes & no & - \\ \hline
+badd9e01 & no & - & yes & no & - \\ \hline
+caea9c01 & no & - & yes & no &  -\\ \hline
+cc366801 & yes (60 min) & 10 out of 10& no & no & - \\ \hline
+d476cd01 & no & - & yes & no & - \\ \hline
+df9df201 & no & - & yes & no & - \\ \hline
+e7cd3901 & - & - & no & yes & - \\ \hline
+f14eb101 & yes (53 min) & 11 out of 11 & no & no & yes \\ \hline
+f658d801 & yes (2 min) & 1 out of 4 & yes & no & - \\
+\hline\end{tabular}
+\caption{The table describes the results of our evaluation.
+The \textit{Input} column describes whether our approach was able to discover valid input strings.
+As previously mentioned, some binaries did not have any predefined input strings (hence such binaries are marked by the \textit{-} in this column).
+The \textit{Number of Inputs} column describes how many input strings our method discovered out of the total number of input strings defined by the binary.}
+\label{binaryresults}
+\end{table*}
 
 Additionally, our input finding method was tested on a binary used for technical interviews.
-The binary was designed as a reverse engineering challenge for interviewers.
-To pass the challenge, the interviewer has to crack the secret passcode to the program.
-The blog post describes the manual process used to crack the password \cite{interviewbinary}.
+The binary was designed as a reverse engineering challenge for interviewees.
+To pass the challenge, the interviewee has to crack the secret passcode to the program.
+A blog post describes the manual process used by one interviewee to crack the password \cite{interviewbinary}.
 The manual process is tedious and it includes methods such as disassembling the binary, stepping through the instructions using a debugger, and reverse engineering the logic of the program.
 As described in the blog post, the cracking of the binary can take several hours and requires advanced skills and years of background knowledge.
-\tool{} cracked the binary in a matter of minutes autonomously, proving itself as a useful tool not only finding valid input but also cracking passwords.
+\tool{} was able to crack the binary in a matter of minutes autonomously.
 
 \subsection{Protocol State Machine Generation}
-% TODO
 The evaluation of the protocol state machine generator was also done using the DARPA Cyber Grand Challenge binaries.
-However, out of the total of \numbinaries{} binaries, only [NUMBER] had an implicit protocol.
-Our method was able to generate a correct protocol state machine for [NUMBER]\% of these binaries.
-
-The reason for failure include ...
+However, out of the total of \numbinaries{} binaries, only 2 had a clear implicit protocol.
+Specifically, these 2 binaries requested authentication before allowing full functionality.
+Our method was able to generate a correct protocol state machine for both of these binaries.
+The protocol state machine generated was manually verified against the source code of the binaries.
 
 \subsection{Case Study}
-One interesting and noteworthy case study is a binary which implements a simple protocol that lets the user call functions applying root64 and parcour schemes.
-The binary is built such that after sending the \textit{HELLO} command, the binary will generate an authentication token which is echoed to the user and must be used for the \textit{AUTH} command.
+To further prove the usefulness of the method presented in this paper, we use one of the Cyber Grand Challenge binaries (i.e.
+250d1101) as a case study and for comparison with a black-box fuzzer and a symbolic execution engine.
+The binary implements a simple protocol that lets the user call functions applying root64 and parcour schemes.
+The binary is built such that after sending the \textit{HELLO} command, it will generate an authentication token which is echoed to the user and must be used for the \textit{AUTH} command.
 Without proper authentication, the \textit{SET} and \textit{CALL} commands have no effect.
-Before generating the protocol state machine, \tool{} analyzed the binary and discovered input strings.
-The inputs discovered are: $HELLO, AUTH, SET, CALL$ and $BYE$.
+Before generating the protocol state machine, \tool{} analyzed the binary and discovered the input strings $HELLO, AUTH, SET, CALL$, and $BYE$ in approximately 20 minutes.
 These inputs are the only inputs implemented by the binary's protocol.
@@
 This shows that the input finding approach in combination with the protocol state machine generator can help identify potential backdoors or unexpected control flow paths for a given program.
+
+When comparing our method with a black box fuzzer and a symbolic execution engine, we ran each technique for 30 minutes and gathered code coverage statistics.
+The black box fuzzer used is the open source \textit{zzuf} fuzzer and the symbolic execution engine used is that of the \textit{Fuzzball} \cite{zzuf, fuzzball}.
+While most popular for its symbolic execution, \textit{Fuzzball} is a tool for binary analysis, built on top of the BitBlaze Vine \cite{bitblaze, bitblazesite}, which can do a number of things including extracting various traces and computing coverage statistics.
+For simplicity reasons, we use the \textit{Fuzzball} tool to collect code coverage statistics.
+For compatibility reasons with the \textit{Fuzzball} tool, we translated the source code of the binary to use UNIX system calls and compiled it in the ELF format.
+The changes are minor and have no impact on the functionality of the program.
+
+The results of the comparison show that \textit{zzuf} covered a total of 8936 basic blocks and was unable to discover any valid input strings.
+At the same time, the \textit{Fuzzball} symbolic execution engine executed a total of 9922 and 10003 basic blocks (explained below) and was able to discover the valid input strings.
+The method presented in this paper executed 10055 basic blocks and not only discovered value input strings but also identified that a user can authenticate themselves with the command \textit{AUTH 0}.
+
+It should be noted that while the fuzzer and our method worked autonomously, the \textit{Fuzzball} symbolic executioner required some reverse engineering of the program to help setup.
+One of the important parameters needed for symbolic execution was the address (and size) of the memory desired to be treated as symbolic.
+The different values of the basic blocks executed by \textit{Fuzzball} are due to multiple executions of the engine with varying parameters.
+Overall, the results prove that the method presented in this paper is efficient with respect to both time and code coverage.
 
@@
 
+One may argue that there are other ways of finding the input strings of a binary.
+One option is the use of the Unix \textit{strings} utility, which finds all null terminated strings inside a binary.
+One problem with this approach is that the \textit{strings} utility may return a number of other strings within the binary, strings not used for input validation.
+Such strings can create a lot of noise and hinders the use of the \textit{strings} utility.
+Additionally, our approach can find dynamically generated input strings which the static \textit{strings} utility is not capable of.
+Another option is the use of the \textit{ltrace} tool which echoes library functions called and their arguments.
+\textit{ltrace} could be used to determine calls of the \textit{strcmp()} function, which can help identify valid input strings.
+However, a program can be implemented to dynamically detect if it is being traced by a utility such as \textit{ltrace}.
+We argue that a complex input validation process can make the \textit{ltrace} and \textit{strings} utlitities less efficient than our method.
+For example, one of the binaries used in the testing of our method, used a tree structure to represent valid input strings (see Appendix A, Example 2).
+Each node in the tree structure represented a valid character and its depth in the tree represented its index in the input string.
+Such an input checking method greatly impacts the efficiency of alternative approaches such as the \textit{strings} and the \textit{ltrace} utilities.
+The \textit{strings} utility would not be useful since the input strings are represented one character at a time.
+Regardless of whether the binary is designed to detect when it is being traced, using an \textit{ltrace}-based technique would also fail since the program compares characters and does not use a library function (such as \textit{strcmp()}).
+On the other hand, our approach successfully finds the valid input strings despite the complexity of the input validation mechanism.
+
 In the future, we wish to improve our current method.
-First, we would like to extend it to handle any type of input.
+First, we would like to extend to handle binary input.
 Furthermore, we would like to apply the input finding technique recursively.
 It is easy to imagine programs for which the second input varies based on the value of the first input.
-Also, as we've experienced during our evaluation, there exist programs which first request for a username, input which will not have any predefined valid values.
+Also, there obviously exist programs for which some inputs have no predetermined value (such as registering a username).
 It would be helpful to determine such cases, provide some input and continue the execution of the program, applying the input finding technique once again.
@@
 However, the protocol state machine does not currently handle situations in which there are no predefined arguments (e.g.
-given input \textit{CREATE} the arguments could be predefined such as \textit{dog} or random such as ``x'').
+given input \textit{CREATE} the arguments could be predefined such as \textit{dog} or random such as any number).
 Ultimately, we would like to generate some random arguments that remain constant for every input string in a given permutation.
@@
 As seen in Figure \ref{fig:psm_example}, the different authentication paths (use of \textit{HELLO} vs.
-omission of \textit{HELLO}) lead to different states when in reality the two states are identical.
+omission of \textit{HELLO}) lead to different states when in reality the two states (i.e.
+user is authenticated) are identical.
+Although we have not yet explored this, state minimization approaches for finite state machines may be applicable for this problem.
 Also, our method is based on the assumption that given a number of input strings, the majority of input string permutations are not valid and do not reflect the expected protocol.
 There may be cases where this assumption does not hold.
-In the future, we would like build onto our current approach for generating protocol state machines. 
-First, we want to be able to determine if two states are identical for a more accurate protocol state machine.
-Furthermore, we want to determine a better approach for identifying the valid sequence of input strings.
-One option would be to monitor other aspects of the program, particularly the \textit{.data}, \textit{.bss}, and the heap during the executions.
-Changes in these may reflect changes of the state of the program.
+A better option might be to monitor other aspects of the program, particularly the \textit{.data} and the \textit{.bss} sections and the heap during the executions and see how they react to various input string permutations.
+In the future, we would like to address the above mentioned concerns with our approach for generating protocol state machines.
 
 \section{Conclusions}
-This paper introduces two novel techniques for autonomously generating valid input strings for closed x86 binaries and for generating protocol state machine using the discovered input.
-For generating valid input strings, our method exploits the number of instructions retired to infer if a given character is valid for a given index of an input string.
+This paper introduces two novel techniques for autonomously generating valid input strings for closed x86 binaries and for generating a protocol state machine using the discovered input.
+For generating valid input strings, our method exploits the number of instructions retired to infer whether a given character is valid for a given index of an input string.
 This process allows us to recursively and incrementally craft valid inputs.
@@
 We show that our proposed methods are efficient and successful by testing them on \numbinaries{} DARPA binaries from the Cyber Grand Challenge \cite{darpacgc}. 
-Unlike previous works, our methods do not have any fundamental limitations (such as constraint solvers) and are robust enough to handle even binaries of other formats.
+Unlike previous works, our methods do not have any fundamental limitations (such as constraint solvers) and are robust enough to handle program binaries of other formats.
 
-We believe that the our techniques could be applied and used for a number of reasons including building test suites, discovering unwanted or unexpected control flow paths and even finding bugs (when used in combination with a fuzzer).
+We believe that the our techniques could be applied and used in a number of settings including the building of test suites, discovering unwanted or unexpected control flow paths, and  finding bugs (when used in combination with a fuzzer).
+
+%\section*{Acknowledgments}
+%The authors would like to thank Thuan Pham for his feedback and useful discussions.
 
@@
 
+\appendix
+\section{Input Validation Code Snippets}
+The following are two examples of input validation mechanisms from the DARPA Cyber Grand Challenge sample binaries.
+
+\subsection{Example 1}
+
+In this example, we can see that the program uses the \textit{strcmp()} function to determine if the user input is a valid command.
+Because of the nature of the \textit{strcmp()} function, when a character at a given index is valid (i.e.
+pertains to one of the commands), the function will then move to the next character of the input string and compare that with the next character of the command.
+In our work, we exploit the change in the number of instructions executed caused by this to find valid input strings.
+
+\definecolor{codegreen}{rgb}{0,0.6,0}
+\definecolor{codegray}{rgb}{0.5,0.5,0.5}
+\definecolor{codepurple}{rgb}{0.58,0,0.82}
+\definecolor{backcolour}{rgb}{0.95,0.95,0.92}
+
+\lstdefinestyle{mystyle}{
+    %backgroundcolor=\color{backcolour},   
+    commentstyle=\color{codegreen},
+    keywordstyle=\color{magenta},
+    numberstyle=\tiny\color{codegray},
+    stringstyle=\color{codepurple},
+    basicstyle=\footnotesize,
+    breakatwhitespace=false,         
+    breaklines=true,                 
+    captionpos=b,                    
+    keepspaces=true,                 
+    numbers=left,                    
+    numbersep=5pt,                  
+    showspaces=false,                
+    showstringspaces=false,
+    showtabs=false,                  
+    tabsize=2
+}
+
+\lstset{style=mystyle}
+
+\lstinputlisting[language=C]{inputval1.txt}
+
+\subsection{Example 2}
+This example comes from a program implementing a Federal Acquisition Regulation (FAR) section lookup service.
+FAR is a set of rules governing the federal government's purchasing process.
+Here the system call \textit{receive()} is used to read input directly from the specified file descriptor.
+The user input is parsed in the \textit{process()} function and the valid input commands are stored in a tree like structure created in the \textit{main()} function.
+Since the system call is used to read the user input, the number of user-land instructions executed will not increase as a function of the input size (as it would in the previous example).
+Our method is able to successfully detect both the valid input strings and the correct expected input size.
+On the other hand, alternative approaches such as the \textit{strings} utility and the \textit{ltrace} tool would fail.
+The \textit{strings} utility can only detect null terminated strings, not characters.
+An \textit{ltrace} based method would also fail since the program compares characters and does not use a library function for this task (such as \textit{strcmp()}).
+
+\lstinputlisting[language=C]{inputval3.txt}
+
 \end{document}

### 2015-11-30 paper/main.tex
final paper version and presentation for PPREW

@@
-\documentclass{acm_proc_article-sp}
+%\documentclass{acm_proc_article-sp}
+\documentclass{sig-alternate-05-2015}
 \usepackage{graphicx}
@@
 \begin{document}
-
-\title{InputFinder: Reverse Engineering Closed Binaries via Side Channels}
-%\numberofauthors{2}
-%\author{
-%\alignauthor
-%Bogdan Copos\\
-%	\affaddr{University of California, Davis}\\
-%	\email{bcopos@ucdavis.edu}
-%\alignauthor
-%Praveen Murthy\\
-%	\affaddr{Fujitsu Laboratories of America}\\
-%	\email{praveen.murthy@us.fujitsu.com}
-%}
+\CopyrightYear{2015} 
+\setcopyright{acmcopyright}
+\conferenceinfo{PPREW-5,}{December 08 2015, Los Angeles, CA, USA}
+\isbn{978-1-4503-3642-0/15/12}\acmPrice{\$15.00}
+\doi{http://dx.doi.org/10.1145/2843859.2843865}
+\title{InputFinder: Reverse Engineering Closed Binaries using Hardware Performance Counters}
+\numberofauthors{2}
+\author{
+\alignauthor
+Bogdan Copos\\
+	\affaddr{University of California, Davis}\\
+	\email{bcopos@ucdavis.edu}
+\alignauthor
+Praveen Murthy\\
+	\affaddr{Fujitsu Laboratories of America}\\
+	\email{praveen.murthy@us.fujitsu.com}
+}
 \maketitle
@@
 In this paper, we propose a new method for autonomously generating valid input and identifying protocols for closed x86 binaries.
-To assess its effectiveness, we have developed \tool{} and tested it against binaries from the DARPA Cyber Grand Challenge example set.
+The method presented can be used as a standalone tool or can be combined with other techniques for improved results.
+To assess its effectiveness, we test \tool{}, the implementation of our method, against binaries from the DARPA Cyber Grand Challenge example set.
 Our evaluations show that our method is not only effective in finding input and determining whether a protocol is expected but can also find unexpected control flow paths.
@@
 
-In this paper, we introduce a new method, called InputFinder, for generating valid user-input for closed binaries. 
-This new method could be useful in autonomous vulnerability scanning systems, or for reverse engineering unknown binaries.
+In this paper, we introduce a new method and its implementation, called InputFinder, for generating valid user-input for closed binaries. 
+This new method could be useful in autonomous vulnerability scanning systems, for reverse engineering unknown binaries, or as a complement to fuzzers or symbolic execution engines.
 Our method is composed of two main components.
-The first component exploits side channels to build valid input for closed binaries.
+The first component exploits hardware performance counters to build valid input for closed binaries.
 Specifically, as the program is given various inputs, \tool{} records the number of instructions \textbf{retired} during the program's execution for each input and uses a statistical measure to infer the program's expected input.
@@
 the instructions which actually impact the program).
-The same side channel is used by \tool{} to also find the expected input size and to categorize the expected input based on type.
+Since our input finding method works by leveraging information about how much work the processor performs during input validation, it does not depend on a particular platform, format (i.e.
+binary, source, etc), or method of obtaining input (e.g.
+stdin, arguments, network socket, file IO).
+However, there are some constraints.
+The underlying method does not work if the user input is transformed (e.g. hashed) before validation. 
+The implementation of our method also has some additional drawbacks.
+Our current implementation can only handle programs whose input is interpreted as a string (or a set of characters).
+The user input can have many components, however, there may not exist any internal dependencies between the components (e.g. value of one component defines the length of another component).
+If the input is composed of multiple components, the program must process the components sequentially in order for our method to work.
+The same hardware counters are used by \tool{} to also find the expected input size and to categorize the expected input based on type.
 The second component uses the discovered input to build a protocol state machine associated with the tested program.
@@
 
-There have also been numerous works studying and applying dynamic similarity.
-In \cite{blanketexec}, the authors propose blanket execution, a novel dynamic equivalence testing technique.
-Blanket execution is designed to collect information about the execution of functions, including side effects, which is then used for similarity testing.
-Another important work is \cite{differential} which introduces \textit{differential slicing} for the purpose of identifying flow and value differences across multiple executions of the same program for security applications.
-Both our approach and the previous approaches mentioned above use a dynamic approach to gather information about the program.
-However, our approach aims at finding differences in executions of a given program in order to make inferences about the underlying protocol expected by the program.
-This allows us to discover the orders of inputs which experience unique control flow paths.
+%There have also been numerous works studying and applying dynamic similarity.
+%In \cite{blanketexec}, the authors propose blanket execution, a novel dynamic equivalence testing technique.
+%Blanket execution is designed to collect information about the execution of functions, including side effects, which is then used for similarity testing.
+%Another important work is \cite{differential} which introduces \textit{differential slicing} for the purpose of identifying flow and value differences across multiple executions of the same program for security applications.
+%Both our approach and the previous approaches mentioned above use a dynamic approach to gather information about the program.
+%However, our approach aims at finding differences in executions of a given program in order to make inferences about the underlying protocol expected by the program.
+%This allows us to discover the orders of inputs which experience unique control flow paths.
 
@@
 
-\subsection{Side Channels} \label{bg_sidechannels}
-Side channels are streams of information that can be retrieved from the hardware of a device running a given program.
-This information can be used to gather knowledge regarding the program's internal components.
-Side channels have been widely explored by security researchers in order to discover weaknesses in programs \cite{weinbergside, schindler2002combined, genkin2014rsa, genkinstealing}.
-In cryptography, side channel attacks exploit such information channels to defeat crypto systems \cite{zhou2005side, black2002black, okeya2006side}.
-Recently, side channel attacks, such as OpenSSL's CREAM cache timing attack have made headlines \cite{creamssl}.
-Similar to CREAM and other side channel attacks, the technique discussed in this paper exploits such information channels to generate valid input for binaries.
+%\subsection{Side Channels} \label{bg_sidechannels}
+%Side channels are streams of information that can be retrieved from the hardware of a device running a given program.
+%This information can be used to gather knowledge regarding the program's internal components.
+%Side channels have been widely explored by security researchers in order to discover weaknesses in %programs \cite{weinbergside, schindler2002combined, genkin2014rsa, genkinstealing}.
+%In cryptography, side channel attacks exploit such information channels to defeat crypto systems \cite{zhou2005side, black2002black, okeya2006side}.
+%Recently, side channel attacks, such as OpenSSL's CREAM cache timing attack have made headlines \cite{creamssl}.
+%Similar to CREAM and other side channel attacks, the technique discussed in this paper exploits such %information channels to generate valid input for binaries.
 \subsection{Hardware counter registers} \label{bg_hardwarereg}
@@
 Such hardware performance counters are often used by software developers to analyze their programs and improve the running time and efficiency.
-However, in this paper we show how such performance counters can also be exploited as side channels.
+However, in this paper we show how such performance counters can also be used for reverse engineering.
 Utilities, such as Linux/Unix \textit{perf}, take advantage of such hardware counters in order to provide developers with useful information about a program's performance.
@@
 \textbf{User-land} instructions are instructions of a program which run outside the operating system's kernel.
-While other options are available for counting instructions retired, we choose \textit{perf} for its efficiency and minimal overhead.
+While other options are available for counting instructions, including program instrumentation, we choose \textit{perf} for its efficiency and minimal overhead.
+Experimental results comparing \textit{perf} and \textit{Pin} instrumentation show that instrumentation is at least 2 times slower.
 The number of \textit{total} instructions retired varies significantly for a binary from execution to execution \cite{weaver2008can}.
@@
 \textit{Pin} is a framework for building customized binary analysis tools and can be used to design a tool which creates a transcript of the instructions retired by a program.
-Once the traces are collected, they are analyze and mappings between instructions of different executions are produced.
+Once the traces are collected, they are analyzed and mappings between instructions of different executions are produced.
 These mappings can be used to understand the relationship between two executions (or two versions of a program).
@@
 \subsubsection{Finding Desired Input Size}
-The same side channels can be used to detect the expected size of the program's input or even the size of the program's input buffer.
-Specifically, there is a change in the behavior of the program with respect to the number of instructions retired when the size of the input is larger than expected.
-Most programs are designed to read a predefined number of bytes from the source of input and store the read bytes in a buffer.
-A popular way of doing that is by using the \textit{fgets()} or \textit{scanf()} functions of the GNU C stdio library.
-As the input size grows, the number of instructions retired also increases linearly.
+A similar approach can be used to detect the expected size of the program's input or even the size of the program's input buffer.
+Specifically, there may be a change in the behavior of the program with respect to the number of instructions retired when the size of the input is larger than expected.
+Most programs are designed to read a predefined number of bytes from the source of input and store the bytes in a buffer.
+However, the change in behavior as the input grows beyond the expected length can be observed and exploited. 
+Usually, as the input size grows, the number of instructions retired also changes (increases usually, except if optimizations are in place).
 This happens because additional instructions are needed to read more characters from the input source.
 However, when the input string is larger than the expected size, the number of instructions retired no longer increases as a function of the input size.
+It should be noted that the CGC binaries do not use the standard library functions.
+Instead, the CGC platform implements their own functions for handling user input, which use the CGC platform system call \textit{receive} to read one character at a time from the input source, in a loop, until a delimiter or a maximum size is reached.
+While the above described behavior is easily observed with CGC binaries, it also extends to programs using the standard GLIBC functions. 
+For example, when using the GLIBC functions such as \textit{getline} and \textit{fgets}, and the user input exceeds the expected length, the number of user-land instructions retired changes and remains constant as long as the input length exceeds the expected length.
 This behavior can be observed and exploited to find the expected input size, which may often also be the size of the input buffer.
+The same behavior is exerted for other similar functions
 Our method observes such differences in the number of user-land instructions retired as the program is given increasingly larger input.
@@
 To start, \tool{} finds the set of all invalid characters for index 0 of all input strings and picks one at random.
-The character selected is used to build invalid input strings of various sizes.
+The character selected is used to build invalid input strings of various lengths.
 It is crucial to use invalid input to avoid noisy fluctuations in the number of instructions retired due to valid input processing.
@@
 These two values may be equal or different, depending on the program's logic.
-The two values may be equal if, for example, the program uses the system call \textit{read()} directly to read data from the input file descriptor.
+
+The two values may be equal if, for example, the program uses the \textit{read()} system call directly to read data from the input file descriptor.
 Since we are measuring the number of user-land instructions retired (not also instructions retired by the kernel), the instructions retired by the \textit{read()} function are not included in our counter so there will be no linear increase in the number of instructions retired as the input size is increased.
-If the two values are equal, the program is executed with increasingly large input until a change in the number of executions is recorded due to the program logic.
+If the two values are equal, the program is executed with increasingly large input until a change in the number of executions is recorded.
+
 If the values differ (as discussed earlier), the difference in the number of instructions retired is recorded and the program is executed, increasing the size of the input, until a change is detected.
-Instead of increasing the input size linearly, the input size is increased exponentially until a change is detected.
+In most cases, the recorded difference represents the number of instructions needed to read an additional character.
+Imagine a program which reads one character at a time, in a loop, until the expected number of characters is read.
+The process stops when the difference in the number of instructions executed between two inputs of two consecutive lengths (i.e input X and input Y are of lengths $n$ and $n+1$) is different from the recorded difference.
+It is also possible that the number of instructions retired oscillates between two (or more values).
+This can happen if the program processes inputs of various lengths differently.
+One example is a palindrome finder, where the execution paths for input of an odd length and input of an even length differ with respect to the number of instructions.
+The same underlying concept can be applied to such cases, where the two (or more) oscillating values are recorded, and the input string is increased until a difference is observed.
+
+To speed up this process, instead of increasing the input size linearly, the input size is increased exponentially until a change is detected.
 Once the change is detected, the input size is decreased to find the exact size at which the change happens.
-It is also possible that the number of instructions retired oscillates between two (or more values).
-One example is a palindrome finder.
-The same underlying concept can be applied to such cases.
 
@@
 \subsection{Generating Protocol State Machine}
-Thus far, we have described how \tool{} discovers valid input for a given binary autonomously.
+Thus far we have described how \tool{} discovers valid input for a given binary autonomously.
 However, some binaries have an implicit protocol.
 Consider, for example, a database management program.
-Without first selecting a database and perhaps even a table, some commands such as \textit{INSERT} or \textit{SELECT} are useless and have no effect on any data.
+Without first selecting a database and perhaps even a table, some commands such as \textit{INSERT} or \textit{SELECT} (or their equivalent) have no effect.
 We develop the second component of \tool{} to address this issue.
@@
 The process is described below and shown as pseudocode in algorithm \ref{psmcode}.
-First, all possible permutations of all sizes from 2 to the number of input strings discovered are computed (e.g.
+The process can be split into three steps.
+\begin{itemize}
+\item First, all of the possible combinations of user inputs (given the discovered input strings) are generated.
+\item Then, the inputs are passed to an instrumented version of the binary, which outputs a single execution trace for each execution (one execution per permutation of input strings).
+\item Last, to determine the expect protocol, the execution traces are compared.
+\end{itemize}
+
+To generate all of the possible combinations of user inputs, all permutations of the discovered input strings are enumerated.
+The permutations range in size (i.e. number of strings) from 2 to $n$, where $n = | input\_strings\_set |$.
+However, enumerating all possible permutation of the discovered input strings may not be enough. 
+Consider cases where the user input is composed of two (or more) strings where the first string is a command and the rest are arguments.
+Now also imagine that the arguments of an input string depend on the previously entered string.
+To cover such cases, for each permutation, we test each input string for arguments (using the previously mentioned input finding technique).
+For example, for a given permutation $\{AUTH, SET\}$, our tool will first try to identify arguments for the \textit{AUTH} input string.
+Without finding the correct argument for the \textit{AUTH} command, the execution of \textit{SET} will not affect the program's state.
+Assume our approach was successful in identifying a single argument, \textit{0}, for the \textit{AUTH} string.
+Now, the tool will attempt to find arguments for \textit{SET} but with \textit{AUTH 0} as the first entered user input.
+Again, this ensures that our approach identifies arguments for strings which depend on previously entered strings.
+
+The resulting permutations of strings (and their arguments) are also passed through the instrumented binary in order to collect execution traces.
+The binary is instrumented with a \textit{Pin} tool, developed by us, which logs the starting address of every basic block executed, as a function of time.
+A basic block is a set of instructions with a single entry and a single exit point.
+Once all permutations of a given size have been executed, the dynamic similarity test analyzes the resulting execution traces.
+Each execution trace is split into sections, each sections associated with the program processing one of the input strings of the input permutation.
+These sections are delimited by a short sequence of specific basic blocks which are executed right before the reading and processing of the user input.
+To determine the \textit{correct} or \textit{expected} input string permutation, the execution trace section of the last command in a permutation is compared against the execution trace sections of that same command, in the same index of the permutation.
+Under the assumption that the execution of the last command will exhibit a different behavior given the correct prior user inputs, our approach identifies which input string permutation causes the execution trace section of the last command to differ from other execution sections of the same command, in the same index of the permutation.
+If such a permutation is discovered, it is added to the protocol state machine.
+
+%First, all possible permutations of all sizes from 2 to the number of input strings discovered are computed (e.g.
 given input strings $a,b,c,d$, the permutations generated include [a,b] , [a,c] , [a,d] , [b,a] , [b,c] , [b,d] ...
@@
 etc.).
-Every permutation is passed through the dynamic analysis engine.
-The dynamic analysis execution engine first splits the permutation into individual input strings.
-Maintaining the original order of the inputs within the permutation, each string is tested for the presence of expected arguments.
-For example, if the input string is \textit{SET} and the program accepts input of the format \textit{SET data}, the analysis tries to determine that \textit{data} is a valid argument by applying the input finding method with \textit{SET} as the command and identifying additional characters after the $<SPACE>$ character.
-The arguments found are saved for later use.
-Once every input string of a given permutation has been tested for arguments, the inputs are passed through the instrumented version of the program binary for the collection of the execution trace.
-The binary is instrumented with a \textit{Pin} tool, developed by us, which logs the starting address of every basic block executed, as a function of time.
-A basic block is a set of instructions with a single entry and a single exit point.
-Once all permutations of a given size have been executed, the dynamic similarity test begins.
-This test splits each execution trace into parts, each part being associated with an input string.
-The splitting process is done by first executing the program without any input, examining the trace, and extracting a short sequence of basic blocks from the end of the trace.
-The sequence of basic blocks extracted is used as the delimiter during the splitting process and it represents the basic blocks executed by the program right before reading and processing the user input.
-We refer to the execution trace of the input string \textit{SET} as $\alpha(SET)$.
-For all permutations of the same size with the same last input string, the traces $\alpha(lastCmd)$ are clustered.
-Given a permutation $p$ of size $s$ let $\alpha(p_s)$ represent the execution trace of the command at index $s$ of the permutation (i.e. last command of the permutation).
-Take for example two different permutations $p1$ and $p2$.
-Both permutations are of the same length, $i$.
-Both permutations are composed of the same input strings and have the same last input string (e.g. [a,b,c,d] vs. [b,a,c,d]).
-Assume permutation $p1$ represents the correct (or expected) order of the input strings.
-It is natural to assume that the input string $p_i$ (i.e.
+%Every permutation is passed through the dynamic analysis engine.
+%The dynamic analysis execution engine first splits the permutation into individual input strings.
+%Maintaining the original order of the inputs within the permutation, each string is tested for the presence of expected arguments.
+%For example, if the input string is \textit{SET} and the program accepts input of the format \textit{SET data}, the analysis tries to determine that \textit{data} is a valid argument by applying the input finding method with \textit{SET} as the command and identifying additional characters after the $<SPACE>$ character.
+%The arguments found are saved for later use.
+%Once every input string of a given permutation has been tested for arguments, the inputs are passed through the instrumented version of the program binary for the collection of the execution trace.
+%The binary is instrumented with a \textit{Pin} tool, developed by us, which logs the starting address of every basic block executed, as a function of time.
+%A basic block is a set of instructions with a single entry and a single exit point.
+%Once all permutations of a given size have been executed, the dynamic similarity test begins.
+%This test splits each execution trace into parts, each part being associated with an input string.
+%The splitting process is done by first executing the program without any input, examining the trace, and extracting a short sequence of basic blocks from the end of the trace.
+%The sequence of basic blocks extracted is used as the delimiter during the splitting process and it represents the basic blocks executed by the program right before reading and processing the user input.
+%We refer to the execution trace of the input string \textit{SET} as $\alpha(SET)$.
+%For all permutations of the same size with the same last input string, the traces $\alpha(lastCmd)$ are clustered.
+%Given a permutation $p$ of size $s$ let $\alpha(p_s)$ represent the execution trace of the command at index $s$ of the permutation (i.e. last command of the permutation).
+%Take for example two different permutations $p1$ and $p2$.
+%Both permutations are of the same length, $i$.
+%Both permutations are composed of the same input strings and have the same last input string (e.g. [a,b,c,d] vs. [b,a,c,d]).
+%Assume permutation $p1$ represents the correct (or expected) order of the input strings.
+%It is natural to assume that the input string $p_i$ (i.e.
 last input string in the permutation) experiences a different execution trace (or control flow path) as a result of permutation $p1$ than permutation $p2$.
-This information is used to determine, for every input string discovered, if there exists a permutation which affects its control flow path.
-If such a permutation is discovered, it is added to the protocol state machine.
+%This information is used to determine, for every input string discovered, if there exists a permutation %which affects its control flow path.
+%If such a permutation is discovered, it is added to the protocol state machine.
 
@@
 The Cyber Grand Challenge organizers also provide the source code of the binaries which, with some effort, can be modified to use UNIX system calls and be built as ELF x86 binaries.
+At the time of this writing, there were only 24 CGC binaries available, all of which were used for the analysis of our approach.
 
@@
 Our method was able to determine the input length for 19 of the 24 binaries.
-Note that the our approach was able to find the expected input length even if it was not able to find valid input.
+Note that our approach was able to find the expected input length even if it was not able to find valid input.
 Additionally, our method was able to determine crash inputs for 5 of the binaries.
@@
 The black box fuzzer used is the open source \textit{zzuf} fuzzer and the symbolic execution engine used is that of the \textit{Fuzzball} \cite{zzuf, fuzzball}.
-While most popular for its symbolic execution, \textit{Fuzzball} is a tool for binary analysis, built on top of the BitBlaze Vine \cite{bitblaze, bitblazesite}, which can do a number of things including extracting various traces and computing coverage statistics.
+While most popular for its symbolic execution, \textit{Fuzzball} is a tool for binary analysis, built on top of the BitBlaze Vine \cite{bitblaze}, which can do a number of things including extracting various traces and computing coverage statistics.
 For simplicity reasons, we use the \textit{Fuzzball} tool to collect code coverage statistics.
@@
 While the method presented in this paper has many advantages, it also has a few limitations.
-For certain programs that allow any input, such as an echo server, our method is not any more efficient than a black box fuzzer.
-Additionally, our method cannot handle cases in which input is composed of multiple inter-dependent fields, such as when the value of a field determines the length of another field and when binaries validate input fields out of order.
-This is one reason why our method failed to detect input for some of the binaries tested.
+The input finding implementation presented in this paper is only capable of finding input for programs which interpret the user input as a string (or set of characters).
+The method presented cannot handle cases in which input is composed of multiple inter-dependent fields. 
+As an example, the method presented fails in cases where the input is composed of multiple fields and the value of a field determines the length of another field or in cases where binaries validate input fields out of order.
+Such failures were observed in the evaluation of \tool{}. 
 Another reason of failure is binaries which accept non printable input (e.g. hex characters, binary, etc.).
 
-One may argue that there are other ways of finding the input strings of a binary.
-One option is the use of the Unix \textit{strings} utility, which finds all null terminated strings inside a binary.
-One problem with this approach is that the \textit{strings} utility may return a number of other strings within the binary, strings not used for input validation.
-Such strings can create a lot of noise and hinders the use of the \textit{strings} utility.
-Additionally, our approach can find dynamically generated input strings which the static \textit{strings} utility is not capable of.
+Another drawback of the presented method is its runtime complexity.
+For any given binary, to find valid input strings (without valid arguments), the runtime is $O\{n*m\}$, where $n$ is the length of the \textbf{longest} valid input string of the valid input string set and $m$ is the total number of valid input strings accepted by the program.
+The process can be parallelized but it is limited by the number of hardware performance counter registers.
+
+In a few cases, experiments have resulted in false positives, especially when experiments were run consecutively without pauses in between.
+Closer analysis showed that these false positives occur because of inaccurate performance counters, a phenomena we cannot yet explain.
+To account for this, an experiment can be re-run.
+
+The method proposed in this paper is not the only way of finding input strings for a closed binary.
+Similar results may be obtained by instrumenting the binary (using \textit{Pin} or even a debugger such as the \textit{GNU GDB}) to count the number of instructions executed.
+However, there is a significant difference in the overhead and runtime between instrumentation and our approach.
+This is one of the reasons why for the protocol generation part, our \textit{Pin} tool recorded basic block executed and \textbf{not} individual instructions.
+In the future, we wish to explore hardware performance counters and see if they can also be used for the protocol generation.
+Another option for finding input strings of a binary is to use the Unix \textit{strings} utility, which finds all null terminated strings inside a binary.
+One problem is that the \textit{strings} utility may return a number of other strings within the binary, strings not used for input validation, resulting in high false positives.
+Additionally, \textit{strings} is restricted to only finding static strings, whereas our approach can find dynamically generated input strings.
 Another option is the use of the \textit{ltrace} tool which echoes library functions called and their arguments.
@@
 Although we have not yet explored this, state minimization approaches for finite state machines may be applicable for this problem.
+
 Also, our method is based on the assumption that given a number of input strings, the majority of input string permutations are not valid and do not reflect the expected protocol.
@@
 A better option might be to monitor other aspects of the program, particularly the \textit{.data} and the \textit{.bss} sections and the heap during the executions and see how they react to various input string permutations.
+Also, we believe hardware performance counters (e.g. branch instructions) may be used to learn more about the program's state and whether a given input caused the program to change states.
 In the future, we would like to address the above mentioned concerns with our approach for generating protocol state machines.
@@
 
-%\section*{Acknowledgments}
-%The authors would like to thank Thuan Pham for his feedback and useful discussions.
+\section{Acknowledgments}
+The authors would like to thank Thuan Pham for his feedback and useful discussions.
 

### 2015-11-30 pprew presentation/pprew-slides.tex
final paper version and presentation for PPREW

\documentclass[pdf]{beamer}
\usepackage{graphicx}
\usepackage{pgfpages}
%\setbeameroption{show notes on second screen=right}
\mode<presentation>{}
\title{InputFinder: Reverse Engineering Closed Binaries using Hardware Performance Counters}
\author[shortname]{Bogdan Copos \inst{1} \and Praveen Murthy \inst{2}}
\institute[shortinst]{\inst{1} University of California, Davis \and \inst{2} Fujitsu Laboratories of America}

\begin{document}

% TITLE
\begin{frame}
	\titlepage
\end{frame}

% INTRODUCTION
\begin{frame}{Introduction}
\begin{itemize}
\item Most program analysis techniques necessitate test suites
\item Test suite availability and completeness issues
\item Analysis dependencies:
	\begin{itemize}
	\item platform
	\item architecture
	\item binary format
	\item source code availability
	\item ...
	\end{itemize}
\end{itemize}
\end{frame}

% BACKGROUND
%% BLACK FUZZING
\begin{frame}{Background}
\framesubtitle{\textit{Fuzzing}}
\textit{Black-box fuzzing} 
\begin{itemize}
\item Randomly generated input
\item Advantages
	\begin{itemize}
	\item easy to implement
	\item previous success
	\end{itemize}
\item Disadvantages
	\begin{itemize}
	\item poor code coverage
	\item slow
	\end{itemize}
\end{itemize}
\vspace{\baselineskip}
\end{frame}
%% WHITE FUZZING
\begin{frame}{Background}
\framesubtitle{\textit{Fuzzing}}
\textit{White-box fuzzing} 
\begin{itemize}
\item Systematically generated input
\item Advantages
	\begin{itemize}
	\item improved effectiveness
	\end{itemize}
\item Disadvantages
	\begin{itemize}
	\item limitations of constraint sovler (if symbolic execution is used)
	\item source code dependencies (if instrumentation is used, i.e. afl-fuzz)
	\end{itemize}
\end{itemize}
\vspace{\baselineskip}
\end{frame}

%% SYMBOLIC EXECUTION
\begin{frame}{Background}
\framesubtitle{\textit{Symbolic Execution}}
\begin{itemize}
\item Execution of program with symbols as input
\item Build path conditions %a condition on the input symbols such that if a path is feasible its path-condition is satisfiable
\item Constraint-solver limitations
	\begin{itemize}
	\item non-linear constraints
	\item scalability (path explosion)
	\end{itemize}
\item \textit{Concolic execution}
\end{itemize}
\end{frame}

% PROBLEM STATEMENT=
\begin{frame}{Problem Statement}
\textcolor{red}{Platform, architecture, format, source-code} \textbf{independent} method for generating input for programs.
\end{frame}

%EXPLANATION
\begin{frame}{}
Programs will normally
\begin{itemize}
\item read input
\pause
\item \textcolor{red}{validate input}
\pause
\item process
\pause
\item change state (possibly)
\pause
\item provide output (possibly)
\end{itemize}
\end{frame}

\begin{frame}{Validation}
The processor may perform \textcolor{red}{additional} work during input validation.

\vspace{\baselineskip}

E.g.: \textit{strcmp()} stops at first character mismatch.
\pause

\vspace{\baselineskip}

\textbf{Use this observation to identify valid input}

\end{frame}
\note{The amount of work done by the processor during the validation process depends on the validity of the input. For example, when using GLIBC's strcmp function, the more valid characters the input has, the more work the processor will perform. This information can be used to identify when the program is given valid input. For the purpose of this work, we defined valid input as input that passes the validation mechanism. Valid input may be benign but it may also cause the program to crash.}

% PROPOSE APPROACH
\begin{frame}{Proposed Approach}
Leverage information exposed by the hardware [performance counters] during the program's execution
\vspace{\baselineskip}
\begin{enumerate}
\item Analyze the number of instructions retired to incrementally build input strings
\item Use discovered input and dynamically generated traces to identify the expected protocol
\end{enumerate}
\end{frame}

% ASSUMPTIONS
\begin{frame}{}
Assumptions:
\begin{itemize}
\item Instructions retired vary based on the validity of the input.
\pause
\item Validation occurs in user-space
\pause
\item Most printable characters are not valid at a given input string index.
\pause
\item Input is validated in order.
\pause
\item No inter dependencies between fields of a single input.
\pause
\item Input is not altered prior to validation.
\end{itemize}
\end{frame}
\note{For our approach to be successful, we make several assumptions. One assumption is that, as mentioned before, the number of instructions retired varies depending on the input and its validity. The more valid characters an input has, the more work is done by the processor during validation. We also assume that validation occurs in user-space and we can measure those instructions. Another assumption is that for a given program, for any of its input strings, the majority of the printable characters will not be valid at a given index. Currently our approach cannot handle cases where the input is not validated in order, where the input is composed of multiple fields that are inter dependent, or where the input undergoes transformations, such as hashing or encryption, before being validated. }

% METHODOLOGY
%\begin{frame}{Methodology}
%\end{frame}

% FINDING INPUT
\begin{frame}{Finding Input}
\begin{enumerate}
\item Start with an empty input string
\pause
\item Count the number of user-land instruction retired as the program is given as input each character from the set of printable characters
\pause
\item Calculate the mode number of user-land instructions 
\pause
\item Valid characters for the current index of the input are characters for which the number of user-land instructions retired is not within the range of mode +/- threshold
\item Append one of the discovered valid characters to the current input string and repeat steps 2-5 (until no more valid characters are identified).
\end{enumerate}
\end{frame}

\begin{frame}{}
\begin{figure}
\includegraphics[scale=0.90]{../paper/architecture.pdf}
\end{figure}
\end{frame}

\begin{frame}{Protocol State Machine Generation}
To discover the expected protocol:
\begin{enumerate}
\item Generate every permutations of input strings
\item Discover arguments for input strings
\item Test permutations with instrumented binary
\item Compare execution traces
\end{enumerate}
\end{frame}
\note{To find the underlying expected protocol implemented by a given program, we first start with the input strings generated by the input finding technique described earlier. We generate all permutations of those strings, of all sizes from 2 to the number of input strings. So if we have 10 input strings, our largest permutation will be composed of 10 strings. Next, we take these permutations and while maintaining the order of the input strings, we test the strings for arguments using our input finding approach. This allows us to find cases where a given command is affected by the arguments of the previous command. One example of this is authentication. Once done, we test all permutations using an instrument binary, starting with the shortest permutations. The binary is instrumented using a PIN tool which records an execution trace. We then compare the execution traces for executions for which the last input is the same (but previous input strings differ). We break up the execution trace by input command. For example, if an execution trace reflects 3 input strings. After splitting the trace file, we will have 3 parts, one corresponding to each command. Assuming most combinations of input strings will be incorrect with respect to the expected protocol, we eliminate all execution traces where the last part is identical. More simply put, if the execution trace for the last command isn't different, nothing special happened so the permutation must be incorrect. However, if compared to other traces, the last part differs for a given permutation, the execution of the last command took a different control path and we identify that as being the correct permutation. It should be noted that similarly to our input finding approach, there is also an implicit "majority" rule, where we assume most permutations of a given length are incorrect and do not reflect the expected protocol.}

\begin{frame}{Execution Traces}
Instrument binary using PIN to record every basic block (defined by its starting address) executed in order of execution
\begin{figure}
\begin{center}
\includegraphics[scale=0.5]{../paper/traceexample.pdf}
\end{center}
\end{figure}
\end{frame}

\begin{frame}{Example}
Valid input strings: HELLO, AUTH, SET, CALL, BYE

Permutations: [ [ HELLO, AUTH ], [ HELLO, CALL ], ... [HELLO, AUTH, SET, CALL, BYE] ]

Arguments: check arguments for HELLO; send HELLO, check arguments for AUTH; send HELLO, check arguments for CALL ... send HELLO, AUTH, SET, CALL, check arguments for BYE

\end{frame}

% EVALUATION
\begin{frame}{Evaluation}
\begin{itemize}
\item 24 DARPA Cyber Grand Challenge (CGC) binaries
\item 1 cyber security interview challenge (password cracking program) 
\end{itemize}
\end{frame}

\begin{frame}{DARPA Cyber Grand Challenge}
\begin{itemize}
\item Architecture (32 bit, x86)
\item Binary format
\item System calls
	\begin{itemize}
	\item \textit{allocate()}
	\item \textit{deallocate()}
	\item \textit{terminate()}
	\item \textit{fdwait()}
	\item \textit{random()}
	\item \textit{receive()}
	\item \textit{transmit()}
	\end{itemize}
\item No standard library
\end{itemize}
\end{frame}

% RESULTS
\begin{frame}{Results}
\begin{table}
\scalebox{0.55}{
\begin{tabular}{|c|c|c|c|c|c|} \hline
\textbf{Binary} & \textbf{Input} & \textbf{Number of Inputs} & \textbf{Input Size} & \textbf{Crash Input} & \textbf{Protocol State Machine}\\ \hline
06459301 & yes (8 min) & 8 out of 8 & yes & no & -\\ \hline
06b71301 & no & - & yes & no & -\\ \hline
07a9a901 & - & - & yes & yes & - \\ \hline
0b32aa01 & yes (2 min) & 1 out of 1 & yes & yes & - \\ \hline
11dc8e01 & no & - & yes & no & - \\ \hline
1877a601 & yes (11 min) & 6 out of 6 & yes & no & - \\ \hline
250d1101 & yes (20 min) & 5 out of 5 & yes & no & yes \\ \hline
2eca0101 & yes (42 min) & 7 out of 7 & no & no & - \\ \hline
37e97201 & yes (38 min) & 6 out of 6 & yes & no & - \\ \hline
3dcf1a01 & yes (6 min) & 6 out of 6 & yes & yes & - \\ \hline
48b9cf01 & yes (50 min) & 9 out of 9 & yes & no & - \\ \hline
65884701 & no & - & no & no & - \\ \hline
701b7301 & - & - & yes & no & - \\ \hline
7262d006 & yes (5 min) & 4 out of 5 & yes & no & - \\ \hline
7fa39f01 & no & - & yes & yes & - \\ \hline
b8993403 & yes (12 min) & 1 out of 1 & yes & no & - \\ \hline
badd9e01 & no & - & yes & no & - \\ \hline
caea9c01 & no & - & yes & no &  -\\ \hline
cc366801 & yes (60 min) & 10 out of 10& no & no & - \\ \hline
d476cd01 & no & - & yes & no & - \\ \hline
df9df201 & no & - & yes & no & - \\ \hline
e7cd3901 & - & - & no & yes & - \\ \hline
f14eb101 & yes (53 min) & 11 out of 11 & no & no & yes \\ \hline
f658d801 & yes (2 min) & 1 out of 4 & yes & no & - \\ \hline
interview challenge & yes (2 min) & 1 out of 1 & yes & - & - \\ \hline
\end{tabular}}
\end{table}
\end{frame}
\note{The table describes the results of our evaluation. The \textit{Input} column describes whether our approach was able to discover valid input strings. As previously mentioned, some binaries did not have any predefined input strings (hence such binaries are marked by the \textit{-} in this column). The \textit{Number of Inputs} column describes how many input strings our method discovered out of the total number of input strings defined by the binary.}

\begin{frame}{Results}
\begin{itemize}
\item CGC
	\begin{itemize}
	\item Binaries with well defined input: 21 out of 24
	\item Found some input strings 13 out of 21 (62\%)
	\item Found all input strings 11 out of 13 (85\%)
	\end{itemize}
\item Interview challenge password cracked
\end{itemize}
\end{frame}

\begin{frame}{Case Study}
\framesubtitle{\textit{250d1101}}

Binary implements a simple protocol that allows a user perform \textit{root64} encoding/decoding and utilize \textit{parcour} schemes.

\begin{itemize}
\item 5/5 input strings discovered
\item Correct protocol identified, including hidden authentication backdoor
\end{itemize}
\begin{figure}
\includegraphics[scale=0.35]{../paper/protocoldiagram.pdf}
\end{figure}
\end{frame}
\note{The user must first authenticate themselves before being able to perform any actions. The authentication process begins with the user sending the HELLO command to which the program responds with a randomly generated authentication token. The user can use that token to authenticate themselves using the provided token. Once authenticated, the user can use the SET and CALL commands to ... and the BYE command to terminate the program.}

% LIMITATIONS
%\begin{frame}{Limitations}
%\end{frame}

% FUTURE WORK
\begin{frame}{Future Work}
\begin{itemize}
\item Eliminate the "majority" constraint
\item Use hardware counters for protocol state machine generation
\end{itemize}
\end{frame}
\note{One of the biggest weaknesses}

% TODO:
% - citations (challenge interview)
% - write notes for slides
% 	- mention performance of hardware counters

\end{document}

### 2015-11-30 paper/piechart.py

@@ -1,20 +1,0 @@
-from pylab import *
-
-# make a square figure and axes
-figure(1, figsize=(5,5))
-#ax = axes([0.1, 0.1, 0.8, 0.8])
-
-# The slices will be ordered and plotted counter-clockwise.
-labels = 'YES', 'N/A', 'NO'
-fracs = [58.34,  12.5, 29.16]
-explode=(0, 0, 0)
-
-pie(fracs, explode=explode, labels=labels,
-                autopct='%1.1f%%', shadow=False, startangle=90)
-                # The default startangle is 0, which would start
-                # the Frogs slice on the x-axis.  With startangle=90,
-                # everything is rotated counter-clockwise by 90 degrees,
-                # so the plotting starts on the positive y-axis.
-
-
-show()

### 2015-12-04 pprew presentation/pprew-slides.tex
final presentation version

@@
 \usepackage{pgfpages}
-%\setbeameroption{show notes on second screen=right}
+%\setbeameroption{show only notes}
 \mode<presentation>{}
@@
 \institute[shortinst]{\inst{1} University of California, Davis \and \inst{2} Fujitsu Laboratories of America}
+\date{}
 
@@
 \begin{itemize}
-\item Most program analysis techniques necessitate test suites
+\item Dynamic program analysis techniques necessitate test suites
 \item Test suite availability and completeness issues
@@
 \end{frame}
+
 %% WHITE FUZZING
@@
 	\begin{itemize}
-	\item limitations of constraint sovler (if symbolic execution is used)
+	\item limitations of constraint solver (if symbolic execution is used)
 	\item source code dependencies (if instrumentation is used, i.e. afl-fuzz)
@@
 \begin{itemize}
-\item Execution of program with symbols as input
-\item Build path conditions %a condition on the input symbols such that if a path is feasible its path-condition is satisfiable
+\item Execution of a program with symbols as input
+\item Build path conditions
+\item Advantages
+	\begin{itemize}
+	\item code coverage
+	\item good results
+	\end{itemize} 
+\item Disadvantages
+	\begin{itemize}
 	\item Constraint-solver limitations
@@
 		\end{itemize}
+	\end{itemize}
 \item \textit{Concolic execution}
@@
 \end{frame}
+\note{A condition on the input symbols such that if a path is feasible its path-condition is satisfiable}
 
-% PROBLEM STATEMENT=
+\begin{frame}{DARPA Cyber Grand Challenge}
+
+\textbf{Mission}: Design and build autonomous security systems which are capable of finding and patching vulnerabilities in unknown, closed source binaries.
+
+\end{frame}
+\note{This work started while working on a system for the DARPA Cyber Grand Challenge.
+The mission was to build fully autonomous systems capable of finding and patching vulnerabilities vulnerabilities in unknown, stripped, closed source binaries.}
+
+\begin{frame}{DARPA Cyber Grand Challenge}
+\begin{itemize}
+\item Architecture (32 bit, x86)
+\item Binary format
+\item System calls
+	\begin{itemize}
+	\item \textit{allocate()}
+	\item \textit{deallocate()}
+	\item \textit{terminate()}
+	\item \textit{fdwait()}
+	\item \textit{random()}
+	\item \textit{receive()}
+	\item \textit{transmit()}
+	\end{itemize}
+\item No standard libraries 
+\end{itemize}
+\end{frame}
+\note{The binary format was similar to ELF although there were some differences.
+For example, no dynamic linking was allowed.
+Also, the binaries were x86 32 bit but did not use the UNIX system calls.
+Instead there were 7 system calls available.
+This also means that the binaries did not use the GNU standard libraries.}
+
+% PROBLEM STATEMENT
 \begin{frame}{Problem Statement}
@@
 \begin{frame}{Validation}
-The processor may perform \textcolor{red}{additional} work during input validation.
+The processor performs \textcolor{red}{additional} work during input validation.
+\pause
+\vspace{\baselineskip}
 
+The amount of work varies based on the validity of the input.
+
+\pause
 \vspace{\baselineskip}
@@
 \end{frame}
+\note{Our approach leverages information exposed by the hardware performance counters during the program's execution.
+Similar results can be obtained by using instrumentation, however, experimental results show that instrumentation is significantly slower.
+The only overhead of using hardware counters is in reading the values from the counters.
+There are two main components.
+The first analyzes the number of instructions retired to incrementally build input strings.
+Second, using the discovered input and execution traces, we identify the expected protocol.}
 
@@
 \begin{itemize}
-\item Instructions retired vary based on the validity of the input.
+\item Number of instructions retired vary based on the validity of the input.
 \pause
@@
 \pause
-\item Input is validated in order.
+\item Input is interpreted as a string and validated in order.
 \pause
@@
 \end{frame}
-\note{For our approach to be successful, we make several assumptions.
+\note{\tiny While our method can be a standalone tool, it is not meant to completely replace symbolic execution engines or fuzzers.
+We believe its strengths complement the weaknesses of other methods.
+That said, our approach does make some assumptions that must hold.
+For our approach to be successful, we make several assumptions.
 One assumption is that, as mentioned before, the number of instructions retired varies depending on the input and its validity.
@@
 Another assumption is that for a given program, for any of its input strings, the majority of the printable characters will not be valid at a given index.
-Currently our approach cannot handle cases where the input is not validated in order, where the input is composed of multiple fields that are inter dependent, or where the input undergoes transformations, such as hashing or encryption, before being validated.
-}
+Currently our approach cannot handle cases where the input is not validated in order, where the input is composed of multiple fields that are inter dependent, or where the input undergoes transformations, such as hashing or encryption, before being validated.}
 
@@
 \item Valid characters for the current index of the input are characters for which the number of user-land instructions retired is not within the range of mode +/- threshold
+\pause
 \item Append one of the discovered valid characters to the current input string and repeat steps 2-5 (until no more valid characters are identified).
@@
 \end{frame}
+\note{\tiny First we start with an empty input string.
+We do not assume any prior knowledge of the binary.
+The input finding process begins by testing the binary, with printable characters, one at a time, as the input.
+Using the UNIX perf utility, we count the number of user land instructions retired during each execution.
+We record these counter values and at the end, calculate the mode.
+This is where the "majority" rule comes into play.
+We assume that valid characters are characters for which the number of instructions retired is outside the range of mode +/- some threshold.
+Essentially, we assume that most printable characters are not valid, therefore will take the same execution path.
+On the other hand, valid characters will take different paths (compared to non-valid characters).
+The threshold is there because hardware performance counters are not perfect and from our experiments, the number of user land instructions retired can vary by 1 or 2 instructions for identical executions.
+To build entire valid input strings, we continue steps 2 through 5 until no more valid characters are identified.}
 
+% FINDING INPUT ARCHITECTURE
 \begin{frame}{}
@@
 \end{frame}
+
+%FINDING INPUT SIZE
+\begin{frame}{Finding Input Size}
+\textbf{Idea}: Work done by processor to read input is a function of the input size and the input buffer size
+\begin{enumerate}
+\item Send three short inputs (e.g. 3,4,5 characters long)
+\item Record delta (difference in number of instructions retired between consecutive length inputs)
+\item Increase input size exponentially and send three inputs of consecutive lengths (e.g. 100,101,102 characters long)
+\item Compare current deltas with expected delta
+\item Repeat steps 3-4 until change is observed
+\end{enumerate}
+
+\end{frame}
+\note{}
+
+%CLASSIFYING INPUT
+\begin{frame}{Classifying Input}
+Determine type of accepted characters even when the program does not have predefined input/commands.
+
+\pause
+\vspace{\baselineskip}
+
+Help fuzzers achieve better results, faster.
+
+\end{frame}
+\note{Another benefit of our method is that we can help classify the type of input a program is willing to accept.
+For example, consider programs which do not have a predefined set of commands or protocol.
+However, they only accept certain type of printable characters.
+Our approach is able to determine that by finding the set of acceptable characters and doing simple analysis on them.
+This helps fuzzers achieve better coverage faster.}
 
@@
 \end{frame}
-\note{To find the underlying expected protocol implemented by a given program, we first start with the input strings generated by the input finding technique described earlier.
+\note{\fontsize{4pt}{0}\selectfont To find the underlying expected protocol implemented by a given program, we first start with the input strings generated by the input finding technique described earlier.
 We generate all permutations of those strings, of all sizes from 2 to the number of input strings.
@@
 \begin{frame}{Example}
-Valid input strings: HELLO, AUTH, SET, CALL, BYE
+\textbf{Valid input strings}: HELLO, AUTH, SET, CALL, BYE
 
-Permutations: [ [ HELLO, AUTH ], [ HELLO, CALL ], ... [HELLO, AUTH, SET, CALL, BYE] ]
+\pause
+\vspace{\baselineskip}
 
-Arguments: check arguments for HELLO; send HELLO, check arguments for AUTH; send HELLO, check arguments for CALL ... send HELLO, AUTH, SET, CALL, check arguments for BYE
+\textbf{Permutations}: [ [ HELLO, AUTH ], [ HELLO, CALL ], ... [HELLO, AUTH, SET, CALL, BYE] ]
+
+\pause
+\vspace{\baselineskip}
+
+\textbf{Arguments}: check arguments for HELLO; send HELLO, check arguments for AUTH; send HELLO, check arguments for CALL ...
+
+\pause
+\vspace{\baselineskip}
+
+\textbf{Compare traces}: [HELLO, AUTH] vs. [CALL, AUTH] vs. [SET, AUTH], vs. [BYE, AUTH] ...
 
@@
 \item 1 cyber security interview challenge (password cracking program) 
-\end{itemize}
-\end{frame}
-
-\begin{frame}{DARPA Cyber Grand Challenge}
-\begin{itemize}
-\item Architecture (32 bit, x86)
-\item Binary format
-\item System calls
-	\begin{itemize}
-	\item \textit{allocate()}
-	\item \textit{deallocate()}
-	\item \textit{terminate()}
-	\item \textit{fdwait()}
-	\item \textit{random()}
-	\item \textit{receive()}
-	\item \textit{transmit()}
-	\end{itemize}
-\item No standard library
 \end{itemize}
@@
 
-Binary implements a simple protocol that allows a user perform \textit{root64} encoding/decoding and utilize \textit{parcour} schemes.
+Binary implements a simple protocol which allows the user to perform base 64 encoding/decoding and utilize an obfuscation scheme.
+
+\pause
+\vspace{\baselineskip}
+
+The protocol dictates that the user must first authenticate before being able to execute any commands.
+
+\vspace{\baselineskip}
+\pause
 
@@
 \begin{itemize}
-\item Eliminate the "majority" constraint
-\item Use hardware counters for protocol state machine generation
+\item Eliminate the ``majority" constraint
+
+\vspace{\baselineskip}
+\pause
+
+\item Use hardware counters for protocol state machine generation (vs. PIN instrumentation)
 \end{itemize}
 \end{frame}
-\note{One of the biggest weaknesses}
+\note{One of the biggest weaknesses of our approach is the majority constraint.
+We'd like to not have to assume that most characters are invalid for a given index of a string.
+Similarly we'd like to avoid assuming that most permutations of input commands are incorrect.
+We plan on exploring other events which can be measured using hardware counters to eliminate these constraints.
+Also, we'd like to use hardware counters for protocol state machine generation, to avoid the overhead of PIN instrumentation.}
+
+% THANK YOU
+\begin{frame}
+\begin{center}
+Thank you!
+
+\vspace{1in}
+
+Questions?
+\end{center}
+\end{frame}
 
@@
 % - citations (challenge interview)
-% - write notes for slides
-% 	- mention performance of hardware counters
 
